### 2024-11-29 study1/analysis/1_cleaning.qmd
Update 1_cleaning.qmd
@AnafNeves @EmmaLBenn @maisiebennett added another outlier detection check + streamlined the code a bit + added some TODOs

---
title: "Interoception Scale - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
---

## Data Preparation

```{r}
#| warning: false
#| code-fold: false
#| error: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)

df <- read.csv("../data/rawdata_participants.csv")
```

## Feedback

### Experiment Enjoyment

```{r}
#| warning: false

#| error: false

p_feed <- estimate_density(df$Experiment_Enjoyment) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_bar(data=df, aes(x = Experiment_Enjoyment, fill = Condition), bins=28) +
  scale_fill_manual(values = c("Dimensions"= "#ff944d","Domains" = "#ca0020", "Random"="#ff80d5")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Experiment Enjoyment and Condition", y = "Number of Participants", fill = " ", x = "Ratings of enjoyment on a scale 1-5", subtitle = "Distribution of Participants' Experiment Enjoyment and Condition") +
  facet_wrap(~Condition) +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
  )
p_feed

```

**TODO: Add feedback comments table**

## Exclusions

```{r}
outliers <- list()
```

### Attention Checks

```{r}
#| warning: false
#| error: false

dfchecks <- df |>
  dplyr::mutate(
    # "I always know that I am attentively doing a study"
    A1 = ifelse(Sexual_State_A == 6 | Sexual_State_A == 5, 0, 1),
    # "Even if I am anxious, I should now answer all the way to the left"
    A2 = ifelse(Anxious_SkinThermo_A == 0, 0, 1),
    # "I often experience sensations, and I will answer zero to this question"
    A3 = ifelse(Nociception_ColonBladder_A == 0, 0, 1),
    # "In general, I am very sensitive and attentive to the questions I am currently answering"
    A4 = ifelse(Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5, 0, 1),
    # "I often pay attention to the answers I am giving"
    A5 = ifelse(Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5, 0, 1),
    # "I can always accurately answer to the left on this question to show that I am reading it"
    A6 = ifelse(Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1, 0, 1),
    # "I can always accurately perceive that to this question I should answer the lowest option"
    A7 = ifelse(Accuracy_Genital_A == 0, 0, 1),
    # "Sometimes I notice that I need to answer all the way to the right"
    A8 = ifelse(Confusion_ColonBladder_A == 6, 0, 1),
    .keep = "none"
  ) 
dfchecks$Total <- rowSums(dfchecks)

dfchecks |>
  mutate(Total = as.factor(paste0(Total, "/8"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

```{r}
outliers$attentionchecks <- df$Participant[dfchecks$Total >= 1]
```

We removed `r length(outliers$attentionchecks)` (`insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for having failed at least 1 attention check (out of 8).

### Mahalanobis Distance

```{r}
# Compute distance
dfoutlier <- performance::check_outliers(select(df, contains("_Q")), 
                                         method=c("lof"),
                                         threshold=list(lof = 0.0001)) |> 
  as.data.frame() |> 
  mutate(Participant = fct_reorder(df$Participant, Distance_LOF),
         Outlier_AttentionCheck = ifelse(Participant %in% outliers$attentionchecks, 1, 0),
         Outlier = ifelse(Outlier_LOF == 1, "Distance Outlier", "Valid"),
         Outlier = ifelse(Outlier_AttentionCheck == 1, "Failed Attention Check", Outlier))

# parameters::parameters(glm(Outlier_AttentionCheck ~ Distance_LOF, data = dfoutlier, family = "binomial"))

outliers$distance <- as.character(dfoutlier[dfoutlier$Outlier_LOF == 1, "Participant"])
outliers$distance <- outliers$distance[!outliers$distance %in% outliers$attentionchecks]

dfoutlier |> 
  ggplot(aes(x=Participant, y=Distance_LOF)) +
  geom_bar(aes(fill=Outlier), stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

**TODO: Add a little paragraph explaining what it is etc.**
We removed `r length(outliers$distance)` (`insight::format_percent(length(outliers$distance) / nrow(df))`) participants based on multivariate distance.

```{r}
df <- filter(df, !Participant %in% outliers$attentionchecks)
df <- filter(df, !Participant %in% outliers$distance)
```

## Final Sample

::: panel-tabset

## Summary

-The final sample includes `r report::report_participants(df, age = "Age", sex = NA, gender = "Gender")`.

## Age and Gender

```{r p_demographics_age_gender}
#| warning: false

#| error: false

p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=28) +
  geom_vline(xintercept = mean(df$Age), color = "#ca0020", linewidth=1.5) +
  scale_fill_manual(values = c("Male"= "#64B", "Female"= "#8EB69B", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age and Gender", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age and gender") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),)

p_age
```

## Education

```{r p_demographics_education}
#| warning: false

#| error: false

# Did not add education disciplines

p_edu <- df |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

p_edu

# Student
table(df$Student) |>
  as.data.frame() |>
  rename(Student = Var1, Frequency = Freq) |>
  knitr::kable()
```

## Ethinicity

```{r p_demographics_ethnicity}
#| warning: false

#| error: false

p_eth <- df |>
  ggplot(aes(x = Ethnicity)) +
  geom_bar(aes(fill = Ethnicity)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Ethnicity", y = "Number of Participants", subtitle = "Participants per Ethnicity") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.6), angle = 45, hjust = 1),
    axis.title.x = element_blank())

p_eth

```

## Country of Residence

```{r p_demographics_country}
#| warning: false

#| error: false

p_cou <- df |>
 ggplot(aes(x = Country)) +
 geom_bar(aes(fill = Country)) +
 scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
 scale_fill_viridis_d(guide = "none") +
 labs(title = "Country of Residence", y = "Number of Participants", subtitle = "Participants per Country they are currently living") +
 theme_modern(axis.title.space = 15) +
 theme(
   plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
   plot.subtitle = element_text(size = rel(1.2), vjust = 7),
   axis.text.y = element_text(size = rel(1.1)),
   axis.text.x = element_text(size = rel(0.7), angle = 45, hjust = 1),
   axis.title.x = element_blank())

p_cou
```

```{r p_demographics_map}
#| warning: false

#| error: false

p_map <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_map
```
:::

```{r p_summary}
#| warning: false

#| error: false

p_map 
p_age 
p_edu
```

## Save

```{r}
#| warning: false

#| error: false

write.csv(df, "../data/data_participants.csv", row.names = FALSE)
```