### 2024-11-30 study1/analysis/1_cleaning.qmd
add TODO in cleaning
@AnafNeves @EmmaLBenn @maisiebennett

---
title: "Interoception Scale - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
bibliography: references.bib
---

## Data Preparation

```{r}
#| warning: false
#| code-fold: false
#| error: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)

df <- read.csv("../data/rawdata_participants.csv")
```

## Feedback

### Experiment Enjoyment

```{r}
df |> 
  summarise(n = n(), .by=c("Condition", "Experiment_Enjoyment")) |>
  filter(!is.na(Experiment_Enjoyment)) |> 
  mutate(n = n / sum(n),
         Experiment_Enjoyment = fct_rev(as.factor(Experiment_Enjoyment)),
         .by="Condition") |>
  ggplot(aes(y = Condition, x = n, fill = Experiment_Enjoyment)) +
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual(values=c("green", "yellowgreen",  "yellow", "orange", "red")) +
  scale_x_continuous(expand=c(0, 0))  +
  labs(title="Experiment Enjoyment",
       subtitle="Proportion of participants by condition") + 
  guides(fill = guide_legend(reverse=TRUE)) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "top",
        legend.title = element_blank())

lm(Experiment_Enjoyment ~ Condition, data = df) |> 
  modelbased::estimate_contrasts(p_adjust="none", contrast = "Condition") |> 
  display()
```

### Comments

```{r}
data.frame(Comments = df$Experiment_Feedback) |> 
  filter(!tolower(Comments) %in% c(NA, "no", "none", "no thank you", "no not at all", "no thanks", "no comment", "nil", "n.a", "n/a", "na", "not really", "thank you", "nothing", "no not at the moment", "none for now", "thank you!", "thanks", "not really.")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) 
```

## Exclusions

```{r}
outliers <- list()
```

### Attention Checks

```{r}
dfchecks <- df |>
  dplyr::mutate(
    # "I always know that I am attentively doing a study"
    A1 = ifelse(Sexual_State_A == 6 | Sexual_State_A == 5, 0, 1),
    # "Even if I am anxious, I should now answer all the way to the left"
    A2 = ifelse(Anxious_SkinThermo_A == 0, 0, 1),
    # "I often experience sensations, and I will answer zero to this question"
    A3 = ifelse(Nociception_ColonBladder_A == 0, 0, 1),
    # "In general, I am very sensitive and attentive to the questions I am currently answering"
    A4 = ifelse(Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5, 0, 1),
    # "I often pay attention to the answers I am giving"
    A5 = ifelse(Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5, 0, 1),
    # "I can always accurately answer to the left on this question to show that I am reading it"
    A6 = ifelse(Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1, 0, 1),
    # "I can always accurately perceive that to this question I should answer the lowest option"
    A7 = ifelse(Accuracy_Genital_A == 0, 0, 1),
    # "Sometimes I notice that I need to answer all the way to the right"
    A8 = ifelse(Confusion_ColonBladder_A == 6, 0, 1),
    .keep = "none"
  ) 
dfchecks$Total <- rowSums(dfchecks)

dfchecks |>
  mutate(Total = as.factor(paste0(Total, "/8"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

```{r}
outliers$attentionchecks <- df$Participant[dfchecks$Total >= 1]
```

We removed `r length(outliers$attentionchecks)` (`r insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for having failed at least 1 attention check (out of 8).

### Multivariate Distance

```{r}
# Compute distance
dfoutlier <- performance::check_outliers(select(df, contains("_Q")), 
                                         method=c("lof"),
                                         threshold=list(lof = 0.0001)) |> 
  as.data.frame() |> 
  mutate(Participant = fct_reorder(df$Participant, Distance_LOF),
         Outlier_AttentionCheck = ifelse(Participant %in% outliers$attentionchecks, 1, 0),
         Outlier = ifelse(Outlier_LOF == 1, "Distance Outlier", "Valid"),
         Outlier = ifelse(Outlier_AttentionCheck == 1, "Failed Attention Check", Outlier))

outliers$distance <- as.character(dfoutlier[dfoutlier$Outlier_LOF == 1, "Participant"])
outliers$distance <- outliers$distance[!outliers$distance %in% outliers$attentionchecks]

dfoutlier |> 
  ggplot(aes(x=Participant, y=Distance_LOF)) +
  geom_bar(aes(fill=Outlier), stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

Mahalanobis distance [@demaesschalck2000] is commonly used to detect multivariate outliers, which are data points with unusual combinations of scores on two or more variables [@filzmoser2016]. This distance is calculated relative to the centroid—a central point where the means of all variables intersect—with larger distances indicating greater deviation from the center.

This measure takes into account correlations within the dataset, which is especially useful when trying to refine and validate a questionnaire with overlapping items.

We removed `r length(outliers$distance)` (`r insight::format_percent(length(outliers$distance) / nrow(df))`) participants based on multivariate distance.

### Predictors

```{r}
dfoutlier <- full_join(dfoutlier, select(df, Participant, Experiment_Duration), by = join_by(Participant))

parameters::parameters(glm(Outlier_AttentionCheck ~ Distance_LOF, data = dfoutlier, family = "binomial"))
parameters::parameters(glm(Outlier_AttentionCheck ~ poly(Experiment_Duration, 2), data = dfoutlier, family = "binomial"))
parameters::parameters(glm(Distance_LOF ~ poly(Experiment_Duration, 2), data = dfoutlier))
```

```{r}
r_full <- cor(select(df, contains("_Q")))
r_cleaned <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$distance)) |> 
  select(contains("_Q")) |> 
  cor()

data.frame(Full=r_full[lower.tri(r_full)], Clean=r_cleaned[lower.tri(r_cleaned)]) |> 
  pivot_longer(everything()) |> 
  ggplot(aes(x=value)) +
  geom_histogram(aes(fill=name, color=name), bins=60, alpha=0.3, position="identity") +
  theme_minimal()
```

```{r}
#| code-fold: false

df <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$distance))
```

## Final Sample

::: panel-tabset

### Summary

- The final sample includes `r report::report_participants(df, age = "Age", sex = NA, gender = "Gender")`.

**TODO**
- The age plot code should be streamlined (there is no need to compute estimate_density)
- Education: add in the preprocessing a manual reassignment of some of the "Other" manual inputs to our categories (and put Other if not in the categories)
- We can merge the edu and student plot by having student be a "color" aesthethic 
- Make Pie chart for ethnicity (see doggonogo for example). 
- Ethnicity: manually check the others and either assign them to the correct category or put them in "Other"

### Age and Gender

```{r p_demographics_age_gender}
p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=28, position="identity", alpha=0.5) +
  geom_vline(xintercept = mean(df$Age), color = "#ca0020", linewidth=1.5) +
  scale_fill_manual(values = c("Male"= "#64B", "Female"= "#8EB69B", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age and Gender", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age and gender") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),)

p_age
```

### Education

```{r p_demographics_education}
#| warning: false

#| error: false

# Did not add education disciplines

p_edu <- df |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

p_edu

# Student
table(df$Student) |>
  as.data.frame() |>
  rename(Student = Var1, Frequency = Freq) |>
  knitr::kable()
```

### Ethinicity

```{r p_demographics_ethnicity}
#| warning: false

#| error: false

p_eth <- df |>
  ggplot(aes(x = Ethnicity)) +
  geom_bar(aes(fill = Ethnicity)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Ethnicity", y = "Number of Participants", subtitle = "Participants per Ethnicity") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.6), angle = 45, hjust = 1),
    axis.title.x = element_blank())

p_eth

```

### Country of Residence

```{r p_demographics_map}
p_map <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_map
```

```{r}
sort(table(df$Country)) |> 
  as.data.frame() |> 
  gt::gt()
```

:::

```{r p_summary}
p_map 
p_age 
p_edu
```

## Save

```{r}
write.csv(df, "../data/data_participants.csv", row.names = FALSE)
```