### 2025-02-14 study2/analysis/1_cleaning.qmd
update

---
title: "Interoception Scale (Study 2) - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
---

## Data Preparation

```{r}
#| warning: false
#| code-fold: false
#| error: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)

df <- read.csv("../data/rawdata_participants.csv")
```

## Feedback

### Experiment Enjoyment

```{r}
df |> 
  summarise(n = n(), .by=c("Experiment_Enjoyment")) |>
  filter(!is.na(Experiment_Enjoyment)) |> 
  mutate(n = n / sum(n),
         Experiment_Enjoyment = fct_rev(as.factor(Experiment_Enjoyment))) |>
  ggplot(aes(y = n, x = 1, fill = Experiment_Enjoyment)) +
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual(values=c("green", "yellowgreen",  "yellow", "orange", "red")) +
  coord_flip() +
  scale_x_continuous(expand=c(0, 0))  +
  scale_y_continuous(labels = scales::percent)  +
  labs(title="Experiment Enjoyment",
       subtitle="Proportion of participants by condition") + 
  guides(fill = guide_legend(reverse=TRUE)) +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "top",
    legend.title = element_blank()) 
```

### Comments

```{r}
data.frame(Comments = df$Experiment_Feedback) |> 
  filter(!tolower(Comments) %in% c(NA, "no", "none", "no thank you", "no not at all", "no thanks", "no comment", "nil", "n.a", "n/a", "na", "not really", "thank you", "nothing", "no not at the moment", "none for now", "thank you!", "thanks", "not really.")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) 
```

## Exclusions

```{r}
outliers <- list()
```

### Attention Checks

```{r}
dfchecks <- df |>
  dplyr::mutate(
    # "I can always accurately answer to the extreme left on this question to show that I am reading it"
    A1 = ifelse(MINT_AttentionCheck_1 == 0, 0, 1),
    # "I notice that I am being asked to respond all the way to the right"
    A2 = ifelse(MAIA_AttentionCheck_1 == 6, 0, 1),
    # "I can always accurately choose the lowest option"
    A3 = ifelse(IAS_AttentionCheck_1 == 1, 0, 1),
    # "Respond all the way to the right."
    A4 = ifelse(BodyAwareness_AttentionCheck_1 == 5, 0, 1),
    # "I am able to respond all the way to the left"
    A5 = ifelse(TAS_AttentionCheck_1 == 1, 0, 1),
    # "On the whole, I know I must press the highest option"
    A6 = ifelse(PI18_AttentionCheck_1 == 5, 0, 1),
    # "I feel that to show I'm being attentive I will press the lowest option"
    A7 = ifelse(CEFSA_AttentionCheck_1 == 0, 0, 1),
    .keep = "none"
  ) 
dfchecks$Total <- rowSums(dfchecks)

dfchecks |>
  mutate(Total = as.factor(paste0(Total, "/8"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

```{r}
outliers$attentionchecks <- df$Participant[dfchecks$Total >= 1]
```

We removed `r length(outliers$attentionchecks)` (`r insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for having failed at least 1 attention check (out of 8).

### Experiment Duration

```{r}
dfchecks$Duration <- df$Experiment_Duration
dfchecks$Outlier <- ifelse(dfchecks$Total >= 1, 1, 0)
dfchecks <- filter(dfchecks, Duration < 45)

m <- mgcv::gam(Outlier ~ s(Duration), data = dfchecks, family = "binomial")

estimate_relation(m, length=50) |> 
  ggplot(aes(x = Duration, y = Predicted)) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2) +
  geom_line() +
  geom_vline(xintercept=10, linetype="dashed", color="red") +
  theme_minimal() +
  ggside::geom_xsidedensity(data=mutate(dfchecks, 
                                        Outlier=ifelse(Outlier==1, "Failed attention check", "Valid")), 
                            aes(fill=Outlier), alpha=0.3) +
  ggside::theme_ggside_void() +
  labs(title = "Experiment Duration",
       subtitle = "Predicted probability of failing attention checks by duration",
       x = "Experiment Duration (minutes)",
       y = "Probability of failing attention checks") 
```

```{r}
outliers$duration <- as.character(df[df$Experiment_Duration < 10, "Participant"])
outliers$duration <- outliers$duration[!outliers$duration %in% outliers$attentionchecks]
```

We removed `r length(outliers$duration)` (`r insight::format_percent(length(outliers$duration) / nrow(df))`) participants for having completed the experiment in less than 5 minutes.

### Multivariate Distance

```{r}
# Compute distance
dfoutlier <- performance::check_outliers(select(df, contains("MAIA_")), 
                                         method=c("optics"),
                                         threshold=list(optics=2.5, optics_xi=0.03)) |> 
  as.data.frame() |> 
  mutate(Participant = fct_reorder(df$Participant, Distance_OPTICS),
         Outlier_AttentionCheck = ifelse(Participant %in% outliers$attentionchecks, 1, 0),
         Outlier_Duration = ifelse(Participant %in% outliers$duration, 1, 0),
         Outlier = ifelse(Outlier_AttentionCheck == 1, "Failed Attention Checks", "Passed"),
         Outlier = ifelse(Outlier == "Passed" & Outlier_Duration == 1, "Duration", Outlier))

outliers$distance <- as.character(dfoutlier[dfoutlier$Distance_OPTICS < 3, "Participant"])
outliers$distance <- outliers$distance[!outliers$distance %in% c(outliers$attentionchecks, outliers$duration)]

p1 <- dfoutlier |> 
  ggplot(aes(x=Participant, y=Distance_OPTICS)) +
  geom_bar(aes(fill=Outlier), stat="identity") +
  geom_hline(yintercept = 3, linetype="dashed", color="red") +
  labs(fill = "Status") +
  scale_fill_manual(values=c("Duration" = "purple", "Failed Attention Checks"="red", "Passed"="grey")) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.major.x = element_blank())

m <- mgcv::gam(Outlier_AttentionCheck ~ s(Distance_OPTICS), data = dfoutlier, family = "binomial")
# parameters::parameters(m)

p2 <- estimate_relation(m, length=30) |> 
  ggplot(aes(x=Distance_OPTICS, y=Predicted)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha=0.2) +
  geom_line() +
  geom_vline(xintercept=3, linetype="dashed", color="red") +
  theme_minimal()

p1 | p2
```

We removed `r length(outliers$distance)` (`r insight::format_percent(length(outliers$distance) / nrow(df))`) participants based on multivariate distance.

```{r}
#| code-fold: false

df <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$duration, outliers$distance))
```

## Dimension Computation

```{r}
#| code-fold: true

compute_and_remove <- function(df, name="BodyAwareness", pattern=name, method="mean") {
  items <- select(df, starts_with(pattern), -contains("AttentionCheck"))
  df <- df[!names(df) %in% names(items)]
  if(method == "mean") {
    df[[name]] <- rowMeans(items, na.rm=TRUE)
  } else {
    df[[name]] <- rowSums(items, na.rm=TRUE)
  }
  df
}
```

```{r}
#| code-fold: false

df <- compute_and_remove(df, name="IAS", method="sum")

df <- compute_and_remove(df, name="BPQ", pattern="BodyAwareness", method="mean")

df[grepl("^MAIA_.*_R$", names(df))] <- 6 - df[grepl("^MAIA_.*_R$", names(df))]  # Reverse
df <- compute_and_remove(df, name="MAIA_AttentionRegulation", method="mean")
df <- compute_and_remove(df, name="MAIA_BodyListening", method="mean")
df <- compute_and_remove(df, name="MAIA_EmotionalAwareness", method="mean")
df <- compute_and_remove(df, name="MAIA_NotDistracting", method="mean")
df <- compute_and_remove(df, name="MAIA_Noticing", method="mean")
df <- compute_and_remove(df, name="MAIA_NotWorrying", method="mean")
df <- compute_and_remove(df, name="MAIA_SelfRegulation", method="mean")
df <- compute_and_remove(df, name="MAIA_Trusting", method="mean")
```

```{r}
#| code-fold: false

df <- compute_and_remove(df, name="PHQ4_Anxiety", method="sum")
df <- compute_and_remove(df, name="PHQ4_Depression", method="sum")

# TODO: check if any reversed items in the following questionnaires
df <- compute_and_remove(df, name="TAS_EOT", method="mean")
df <- compute_and_remove(df, name="TAS_DDF", method="mean")
df <- compute_and_remove(df, name="TAS_DIF", method="mean")

df <- compute_and_remove(df, name="ERS_Arousal", method="mean")
df <- compute_and_remove(df, name="ERS_Sensitivity", method="mean")
df <- compute_and_remove(df, name="ERS_Persistence", method="mean")

df <- compute_and_remove(df, name="CERQ_Acceptance", method="mean")
df <- compute_and_remove(df, name="CERQ_PositiveRefocusing", method="mean")
df <- compute_and_remove(df, name="CERQ_PositiveReappraisal", method="mean")
df <- compute_and_remove(df, name="CERQ_SelfBlame", method="mean")
df <- compute_and_remove(df, name="CERQ_OtherBlame", method="mean")
df <- compute_and_remove(df, name="CERQ_Rumination", method="mean")
df <- compute_and_remove(df, name="CERQ_Catastrophizing", method="mean")
df <- compute_and_remove(df, name="CERQ_RefocusPlanning", method="mean")
df <- compute_and_remove(df, name="CERQ_Perspective", method="mean")

df <- compute_and_remove(df, name="CEFSA_Body", method="mean")
df <- compute_and_remove(df, name="CEFSA_Agency", method="mean")
df <- compute_and_remove(df, name="CEFSA_Emotion", method="mean")
df <- compute_and_remove(df, name="CEFSA_Familiarity", method="mean")
df <- compute_and_remove(df, name="CEFSA_Connection", method="mean")
df <- compute_and_remove(df, name="CEFSA_Reality", method="mean")
df <- compute_and_remove(df, name="CEFSA_Self", method="mean")

df[grepl("^PI18.*_R$", names(df))] <- 6 - df[grepl("^PI18.*_R$", names(df))]  # Reverse
df$PI_Alive <- rowMeans(df[grepl("^PI18.*A_", names(df))], na.rm=TRUE)
df <- compute_and_remove(df, name="PI_Safe", pattern="PI18_GS", method="mean")
df <- compute_and_remove(df, name="PI_Enticing", pattern="PI18_GE", method="mean")
df$PI_Good <- rowMeans(select(df, PI_Safe, PI_Enticing), na.rm=TRUE)
df <- select(df, -starts_with("PI18_"))
```

## Final Sample

::: panel-tabset

### Summary

The final sample includes `r report::report_participants(df, age = "Age", sex = NA, gender = "Gender")`.

Among these, `r sum(is.na(df$MINT_Awareness_RelA_29))` have missing data for the MINT due to a technical error.

### Age and Gender

```{r p_demographics_age_gender}
p_age <- df |>
  ggplot(aes(x = Age, fill = Gender)) + 
  geom_histogram(data=df, aes(x = Age, fill=Gender), binwidth = 2) +
  geom_vline(xintercept = mean(df$Age), color = "red", linewidth=1.5, linetype="dashed") +
  scale_fill_viridis_d() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(20, max(df$Age), by = 10 )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

p_age
```

### Education

```{r p_demographics_education}
#| warning: false
#| error: false

p_edu <- df |>
  mutate(Student = ifelse(is.na(Student), FALSE, Student),
         Education = fct_relevel(Education, "High school", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Student)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level", fill = "Is currently a student") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

p_edu
```

### Ethnicity

```{r p_demographics_ethnicity}
#| warning: false
#| error: false

p_eth <- df |>
  filter(!is.na(Ethnicity)) |> 
  ggplot(aes(x = "", fill = Ethnicity)) +
  geom_bar() +
  coord_polar("y") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )

p_eth
```

### Country

```{r p_demographics_map}
p_map <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_map
```

```{r}
sort(table(df$Country)) |> 
  as.data.frame() |>
  rename(Country = Var1) |> 
  arrange(desc(Freq)) |> 
  gt::gt()
```

### BMI

```{r}
p_bmi <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=BMI)) +
  annotate("rect", xmin=10, xmax=18.5, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  annotate("rect", xmin=18.5, xmax=25, ymin=0, ymax=Inf, fill="green", alpha=0.25) +
  annotate("rect", xmin=25, xmax=30, ymin=0, ymax=Inf, fill="yellow", alpha=0.25) +
  annotate("rect", xmin=30, xmax=40, ymin=0, ymax=Inf, fill="orange", alpha=0.25) +
  annotate("rect", xmin=40, xmax=Inf, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  geom_density(aes(color=Gender, fill=Gender), alpha=0.5, linewidth=2) +
  labs(title = "BMI distribution") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank())

p_bmi_a <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=Age, y=BMI, color=Gender)) +
  geom_point(size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender), alpha=0.3, method = 'loess', formula = 'y ~ x') +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none")

p_bmi_b <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=Physical_Active, y=BMI)) +
  geom_jitter(aes(color=Gender), width=0.1, size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender, color=Gender), alpha=0.3, method = 'lm', formula = 'y ~ x') +
  labs(x = "Physically Active", y = "") +
  scale_x_continuous(breaks=0:7) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank()) +
  ggside::geom_xsidehistogram(aes(fill=Gender), bins = 30, position="dodge") +
  ggside::theme_ggside_void()

df$Physical_Workout[df$Physical_Workout > 25] <- NA

p_bmi_c <- df |> 
  filter(!is.na(BMI) & !is.na(Physical_Workout), Gender != "Other") |> 
  ggplot(aes(x=Physical_Workout, y=BMI)) +
  geom_point(aes(color=Gender), size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender, color=Gender), alpha=0.3, method = 'lm', formula = 'y ~ x') +
  labs(x = "Hours of Active Workout", y = "") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none") +
  ggside::geom_xsidehistogram(aes(fill=Gender), bins = 30, position="dodge") +
  ggside::theme_ggside_void()

p_bmi / (p_bmi_a | p_bmi_b | p_bmi_c)
```

### Mental Health

```{r}
df$Disorders_Psychiatric_Mood <- ifelse(str_detect(df$Disorders_Psychiatric, "GAD|Bipolar"), TRUE, FALSE)
df$Disorders_Psychiatric_MoodTreatment <- ifelse(
  df$Disorders_Psychiatric_Mood & !is.na(df$Disorders_PsychiatricTreatment) & str_detect(df$Disorders_PsychiatricTreatment, "Mood|Antidepressant|Anxiolytic|Psychotherapy"), TRUE, FALSE)

df$Disorders_Psychiatric_Anxiety <- ifelse(str_detect(df$Disorders_Psychiatric, "Panic|GAD|Social Phobia|Phobia"), TRUE, FALSE)
df$Disorders_Psychiatric_AnxietyTreatment <- ifelse(
  df$Disorders_Psychiatric_Anxiety & !is.na(df$Disorders_PsychiatricTreatment) & str_detect(df$Disorders_PsychiatricTreatment, "Anxiolytic|Psychotherapy|Mindfulness"), TRUE, FALSE)

df$Disorders_Psychiatric_Eating <- ifelse(str_detect(df$Disorders_Psychiatric, "Eating"), TRUE, FALSE)
df$Disorders_Psychiatric_Addiction <- ifelse(str_detect(df$Disorders_Psychiatric, "Addiction"), TRUE, FALSE)
df$Disorders_Psychiatric_Borderline <- ifelse(str_detect(df$Disorders_Psychiatric, "BPD"), TRUE, FALSE)

df$Disorders_Psychiatric_Autism <- ifelse(str_detect(df$Disorders_Psychiatric, "ASD"), TRUE, FALSE)
df$Disorders_Psychiatric_ADHD <- ifelse(str_detect(df$Disorders_Psychiatric, "ADHD"), TRUE, FALSE)

# df$Disorders_Psychiatric_Other <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)

select(df, Participant, Gender, 
             Disorders_Psychiatric_Mood, Disorders_Psychiatric_MoodTreatment,
             Disorders_Psychiatric_Anxiety, Disorders_Psychiatric_AnxietyTreatment,
             Disorders_Psychiatric_Eating, Disorders_Psychiatric_Addiction, Disorders_Psychiatric_Borderline,
             Disorders_Psychiatric_Autism, Disorders_Psychiatric_ADHD) |> 
  pivot_longer(cols = starts_with("Disorders_"), names_to = "Disorder", values_to = "Value") |> 
  mutate(Disorder = str_remove_all(Disorder, fixed("Disorders_Psychiatric_")),
         Disorder = str_replace(Disorder, "MoodTreatment", "Mood Disorder (with treatment)"),
         Disorder = str_replace(Disorder, "AnxietyTreatment", "Anxiety (with treatment)"),
         Disorder = str_replace(Disorder, "Mood$", "Mood Disorder")) |> 
  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Disorder")) |> 
  mutate(N_tot = sum(N), .by="Disorder") |>
  mutate(Disorder = fct_reorder(Disorder, desc(N_tot)),
         Treatment = ifelse(str_detect(Disorder,  "Treatment"), "With treatment", "Without treatment")) |>
  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

### Somatic

```{r}
somatic <- data.frame(
  Hypermobility = ifelse(str_detect(df$Disorders_Somatic, "Hypermobility Syndrome"), TRUE, FALSE),
  Fibromyalgia = ifelse(str_detect(df$Disorders_Somatic, "Fibromyalgia"), TRUE, FALSE),
  ChronicFatigue = ifelse(str_detect(df$Disorders_Somatic, "Chronic Fatigue Syndrome"), TRUE, FALSE),
  ChronicPain = ifelse(str_detect(df$Disorders_Somatic, "Chronic Pain Syndrome"), TRUE, FALSE),
  BackPain = ifelse(str_detect(df$Disorders_Somatic, "Back Pain"), TRUE, FALSE),
  MuscleTension = ifelse(str_detect(df$Disorders_Somatic, "Muscle Tension"), TRUE, FALSE),
  SkinRashes = ifelse(str_detect(df$Disorders_Somatic, "Skin Rashes"), TRUE, FALSE),
  Eczema = ifelse(str_detect(df$Disorders_Somatic, "Eczema"), TRUE, FALSE),
  Psoriasis = ifelse(str_detect(df$Disorders_Somatic, "Psoriasis"), TRUE, FALSE),
   # Dermatological = ifelse(str_detect(df$Disorders_Somatic, "Other Dermatological"), TRUE, FALSE),
  # Musculoskeletal = ifelse(str_detect(df$Disorders_Somatic, "Other Musculoskeletal"), TRUE, FALSE),
  Sjogrens = ifelse(str_detect(df$Disorders_Somatic, "Sjogren"), TRUE, FALSE),
  ChestPain = ifelse(str_detect(df$Disorders_Somatic, "Chest Pain"), TRUE, FALSE),
  Arrhythmia = ifelse(str_detect(df$Disorders_Somatic, "Cardiac Arrhythmia"), TRUE, FALSE),
  Hypertension = ifelse(str_detect(df$Disorders_Somatic, "Hypertension"), TRUE, FALSE),
  Hypotension = ifelse(str_detect(df$Disorders_Somatic, "Hypotension"), TRUE, FALSE),
  IBS = ifelse(str_detect(df$Disorders_Somatic, "IBS"), TRUE, FALSE),
  GERD = ifelse(str_detect(df$Disorders_Somatic, "GERD"), TRUE, FALSE),
  Crohn = ifelse(str_detect(df$Disorders_Somatic, "Crohn"), TRUE, FALSE),
  UlcerativeColitis = ifelse(str_detect(df$Disorders_Somatic, "Ulcerative Colitis"), TRUE, FALSE),
  Celiac = ifelse(str_detect(df$Disorders_Somatic, "Celiac Disease"), TRUE, FALSE),
  Gluten = ifelse(str_detect(df$Disorders_Somatic, "Gluten"), TRUE, FALSE),
  Lactose = ifelse(str_detect(df$Disorders_Somatic, "Lactose"), TRUE, FALSE),
  ShortBreath = ifelse(str_detect(df$Disorders_Somatic, "Shortness of Breath"), TRUE, FALSE),
  Asthma = ifelse(str_detect(df$Disorders_Somatic, "Asthma"), TRUE, FALSE),
  COPD = ifelse(str_detect(df$Disorders_Somatic, "COPD"), TRUE, FALSE),
  SleepApnea = ifelse(str_detect(df$Disorders_Somatic, "Sleep Apnea"), TRUE, FALSE),
  Bronchitis = ifelse(str_detect(df$Disorders_Somatic, "Chronic Bronchitis"), TRUE, FALSE),
  Nausea = ifelse(str_detect(df$Disorders_Somatic, "Nausea/Vomiting"), TRUE, FALSE),
  Dizziness = ifelse(str_detect(df$Disorders_Somatic, "Dizziness/Lightheadedness"), TRUE, FALSE),
  Migraine = ifelse(str_detect(df$Disorders_Somatic, "Migraine"), TRUE, FALSE),
  Neuropathy = ifelse(str_detect(df$Disorders_Somatic, "Neuropathy"), TRUE, FALSE),
  # Epilepsy = ifelse(str_detect(df$Disorders_Somatic, "Epilepsy"), TRUE, FALSE),
  MS = ifelse(str_detect(df$Disorders_Somatic, "Multiple Sclerosis"), TRUE, FALSE),
  FrequentUrination = ifelse(str_detect(df$Disorders_Somatic, "Frequent Urination"), TRUE, FALSE),
  Endometriosis = ifelse(str_detect(df$Disorders_Somatic, "Endometriosis"), TRUE, FALSE),
  Cystitis = ifelse(str_detect(df$Disorders_Somatic, "Interstitial Cystitis"), TRUE, FALSE),
  PelvicPain = ifelse(str_detect(df$Disorders_Somatic, "Chronic Pelvic Pain Syndrome"), TRUE, FALSE)
)

somatic |> 
  select(where(~n_distinct(.) > 1)) |> 
  correlation::correlation(redundant = TRUE, p_adjust="none") |>
  correlation::cor_sort() |> 
  mutate(label = ifelse(p < .001, "*", ""),
         label = ifelse(r == 1, "", label)) |> 
  ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
  geom_tile() +
  geom_text(aes(label=label), size=3) +
  scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
  labs(title = "Co-occurence Matrix of Somatic Disorders") +
  theme_minimal()  +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_blank())

# pca <- as.data.frame(prcomp(somatic)$rotation[, 1:5])
# pca$PC <- colnames(pca)[max.col(pca, ties.method='first')]
# 
# pca |> 
#   rownames_to_column(var="Disorder") |> 
#   pivot_longer(-c(Disorder, PC)) |> 
#   filter(PC == name) |> 
#   arrange(name, desc(value)) |> 
#   ggplot(aes(x=value, y=Disorder, fill=name)) +
#   geom_bar(stat = "identity") +
#   facet_wrap(~name, scale="free_y")
```

```{r}
#| fig-keep: "first"

# somatic |> 
#   select(where(~n_distinct(.) > 1)) |> 
#   EGAnet::bootEGA(
#     corr="pearson",
#     EGA.type = "EGA.fit",
#     model = "glasso", 
#     algorithm = "leiden",
#     seed=42)
```

```{r}
# df$Disorders_Somatic_Number <- ifelse(df$Disorders_Somatic == "", 0, str_count(df$Disorders_Somatic, ";")+1)

# 1. Autonomic Dysfunction & Anxiety-Somatic Symptoms
# Mechanism: Dysregulated autonomic nervous system (e.g., panic disorder, dysautonomia, hyperventilation).
df$Disorders_Somatic_Autonomic_N <- somatic |> 
  select(Dizziness, ChestPain, ShortBreath, Arrhythmia, Nausea, Hypotension, Hypertension, MuscleTension) |> 
  rowSums()
df$Disorders_Somatic_Autonomic <- ifelse(df$Disorders_Somatic_Autonomic_N >= 1, TRUE, FALSE)

# 2. Functional Somatic Syndromes (Pain, Fatigue, GI Symptoms, and Anxiety)
# Mechanism: Central sensitization, stress-induced HPA axis dysregulation, visceral hypersensitivity.
df$Disorders_Somatic_Functional_N <- somatic |> 
  select(Fibromyalgia, Neuropathy, Migraine, ChronicFatigue, ChronicPain, IBS, BackPain, PelvicPain, FrequentUrination) |> 
  rowSums()
df$Disorders_Somatic_Functional <- ifelse(df$Disorders_Somatic_Functional_N >= 1, TRUE, FALSE)

# 3. Hypervigilance, Mast Cell Dysfunction & Multi-System Reactivity
# Mechanism: Mast cell activation, histamine intolerance, neuroinflammation, heightened interoception.
df$Disorders_Somatic_Hyperreactivity_N <- somatic |> 
  select(Migraine, SkinRashes, Nausea, FrequentUrination, Dizziness, MuscleTension, Endometriosis) |> 
  rowSums()
df$Disorders_Somatic_Hyperreactivity <- ifelse(df$Disorders_Somatic_Hyperreactivity >= 1, TRUE, FALSE)

# 4. Immune-Inflammatory & Gut-Brain Axis Dysregulation
# Conditions: Celiac disease, gluten sensitivity, lactose intolerance, MS, neuropathy, eczema, asthma, IBS.
# Mechanism: Leaky gut, immune activation, microbiome imbalance, chronic low-grade inflammation.
df$Disorders_Somatic_Immune_N <- somatic |> 
  select(Celiac, Gluten, Lactose, MS, Eczema, Asthma, IBS) |> 
  rowSums()
df$Disorders_Somatic_Immune <- ifelse(df$Disorders_Somatic_Immune_N >= 1, TRUE, FALSE)

# 5. Cardiopulmonary & Stress-Linked Disorders
# Mechanism: Vascular and respiratory dysregulation, autonomic instability, stress-exacerbated cardiovascular symptoms.
df$Disorders_Somatic_Cardio_N <- somatic |> 
  select(Arrhythmia, COPD, ChestPain, ShortBreath, Hypertension, SleepApnea) |> 
  rowSums()
df$Disorders_Somatic_Cardio <- ifelse(df$Disorders_Somatic_Cardio_N >= 1, TRUE, FALSE)

# 6. Connective Tissue & Dysautonomia-Linked Disorders
# Conditions: Hypermobility, Sjogren’s syndrome, hypotension, chronic fatigue, dizziness.
# Mechanism: Connective tissue laxity affecting autonomic function (e.g., Ehlers-Danlos syndrome, POTS).
df$Disorders_Somatic_Connective_N <- somatic |> 
  select(Hypermobility, Hypotension, Sjogrens, Hypotension, ChronicFatigue, Dizziness) |> 
  rowSums()
df$Disorders_Somatic_Connective <- ifelse(df$Disorders_Somatic_Connective_N >= 1, TRUE, FALSE)

p2 <- df |> 
  select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
  select(ends_with("_N"), Gender, Participant) |>
  pivot_longer(starts_with("Disorders_Somatic_")) |> 
  summarize(n = n(), .by=c("name", "value", "Gender")) |> 
  mutate(name = str_remove(name, "Disorders_Somatic_"),
         name = str_remove(name, "_N")) |> 
  ggplot(aes(x = value, y=n, fill=Gender)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Number of Issues") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~name)
p2
# 
# p2 <- df |> 
#   select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
#   select(-ends_with("_N"), Gender, Participant) |>
#   pivot_longer(starts_with("Disorders_Somatic_")) |> 
#   summarize(n = n(), .by=c("name", "value", "Gender")) |> 
#   filter(value == TRUE) |> 
#   ggplot(aes(x = name, y=n, fill=Gender)) +
#   geom_bar(stat="identity") +
#   scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
#   scale_y_continuous(expand = c(0, 0)) +
#   labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Type") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
# p2
```

### Wearables

```{r}
#| fig-height: 12

p1 <- df |> 
  ggplot(aes(x = Wearables_Number)) +
  geom_bar() +
  scale_x_continuous(breaks = c(0:max(df$Wearables_Number)), expand=c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Number of Wearable Devices", y = "Number of Participants", x="Number Owned") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal()

p2 <- df |> 
  select(contains("Wearables"), -contains("Importance"), -Wearables_Number) |> 
  pivot_longer(everything()) |> 
  summarize(n = n(), .by=c("name", "value")) |> 
  mutate(value = fct_relevel(value, "Not owning", "Never", "Very rarely", "A few times per week", "Once a day", "A few times per day", "A few times per hour"),
         name = str_remove(name, "Wearables_")) |> 
  ggplot(aes(x=value, y=n, fill=name)) +
  geom_bar(stat = "identity") +
  labs(x = "Usage frequency", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

get_databubble <- function(df, var="Heart") {
  x <- paste0("Wearables_", var)
  y <- paste0("Wearables_", var, "Importance")
  dat <- df[df[[x]] != "Not owning", ]
  dat <- summarize(dat, n = n(), .by = all_of(c(x, y)))
  lvl <- c("Never", "Very rarely", "A few times per week", "Once a day", "A few times per day", "A few times per hour")
  dat[[x]] <- factor(dat[[x]], levels=lvl[lvl %in% unique(dat[[x]])])
  names(dat) <- c("Usage", "Importance", "n")
  dat$Device <- var
  dat
}

p3 <- rbind(
  get_databubble(df, "Heart"),
  get_databubble(df, "Sleep"),
  get_databubble(df, "Steps"),
  get_databubble(df, "Weight"),
  get_databubble(df, "CalorieIntake"),
  get_databubble(df, "CaloriesBurnt")
) |> 
  filter(!is.na(Importance)) |> 
  ggplot(aes(x=Usage, y=Importance)) +
  geom_jitter(aes(size = n, color = Device), width = 0.2, height = 0.2, alpha = 0.5) +
  geom_smooth(aes(x=as.numeric(Usage), color = Device), method = "lm", se = FALSE, formula = 'y ~ x') +
  labs(x = "Usage frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

p1 / p2 / p3
```

:::

## Save

```{r}
df |> 
  select(-contains("AttentionCheck")) |> 
  write.csv("../data/data_participants.csv", row.names = FALSE)
```