### 2025-02-14 study2/analysis/2_analysis.qmd

---
title: "InteroceptionScale (study 2) - Data Analysis"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(EGAnet)
library(tidygraph)
library(ggraph)

set.seed(42)
```

```{r}
#| code-fold: false

df <- read.csv("../data/data_participants.csv") |> 
  filter(!is.na(MINT_Awareness_ExAc_35))
```

## Structure Validation

```{r}
#| code-fold: false

items <- select(df, starts_with("MINT_"))
names(items) <- str_remove(names(items), "MINT_")

colors <- c(MINT="#00BCD4", MINTi="#00838F", IAS="#F44336", MAIA="#FF9800", MAIAi="#EF6C00", BPQ="#795548")
```

```{r}
labels <- list(
  MINT_Deficit_Urin_1 = "I sometimes feel like I need to urinate or defecate but when I go to the bathroom I produce less than I expected",
  MINT_Deficit_Urin_2 = "I often feel the need to urinate even when my bladder is not full",
  MINT_Deficit_Urin_3 = "Sometimes I am not sure whether I need to go to the toilet or not (to urinate or defecate)",
  MINT_Deficit_CaCo_4 = "Sometimes my breathing becomes erratic or shallow and I often don't know why",
  MINT_Deficit_CaCo_5 = "I often feel like I can't get enough oxygen by breathing normally",
  MINT_Deficit_CaCo_6 = "Sometimes my heart starts racing and I often don't know why",
  MINT_Deficit_CaNo_7 = "I often only notice how I am breathing when it becomes loud",
  MINT_Deficit_CaNo_8 = "I only notice my heart when it is thumping in my chest",
  MINT_Deficit_CaNo_9 = "I often only notice how I am breathing when my breathing becomes shallow or irregular",
  MINT_Deficit_Olfa_10 = "I often check the smell of my armpits",
  MINT_Deficit_Olfa_11 = "I often check the smell of my own breath",
  MINT_Deficit_Olfa_12 = "I often check the smell of my farts",
  MINT_Deficit_Sati_13 = "I don't always feel the need to eat until I am really hungry",
  MINT_Deficit_Sati_14 = "Sometimes I don't realise I was hungry until I ate something",
  MINT_Deficit_Sati_15 = "I don't always feel the need to drink until I am really thirsty",
  # MINT_Awareness_SexA_16 = "I always feel in my body if I am sexually aroused",
  # MINT_Awareness_SexA_17 = "I can always tell that I am sexually aroused from the way I feel inside",
  # MINT_Awareness_SexA_18 = "I always know when I am sexually aroused",
  MINT_Awareness_SexS_19 = "During sex or masturbation, I often feel very strong sensations coming from my genital areas",
  MINT_Awareness_SexS_20 = "When I am sexually aroused, I often notice specific sensations in my genital area (e.g., tingling, warmth, wetness, stiffness, pulsations)",
  MINT_Awareness_SexS_21 = "My genital organs are very sensitive to pleasant stimulations",
  MINT_Awareness_SexO_22 = "In general, I am very sensitive to changes in my genital organs",
  MINT_Awareness_SexO_23 = "I can notice even very subtle changes in the state of my genital organs",
  MINT_Awareness_SexO_24 = "I am always very aware of the state of my genital organs, even when I am calm",
  MINT_Awareness_UrSe_25 = "In general, I am very aware of the sensations that are happening when I am urinating",
  MINT_Awareness_UrSe_26 = "In general, I am very aware of the sensations that are happening when I am defecating",
  MINT_Awareness_UrSe_27 = "I often experience a pleasant sensation when relieving myself when urinating or defecating)",
  MINT_Awareness_RelA_28 = "I always know when I am relaxed",
  MINT_Awareness_RelA_29 = "I always feel in my body if I am relaxed",
  MINT_Awareness_RelA_30 = "My body is always in the same specific state when I am relaxed",
  MINT_Awareness_StaS_31 = "Being relaxed is a very different bodily feeling compared to other states (e.g., feeling anxious, sexually aroused or after exercise)",
  MINT_Awareness_StaS_32 = "Being sexually aroused is a very different bodily feeling compared to other states (e.g., feeling anxious, relaxed, or after physical exercise)",
  MINT_Awareness_StaS_33 = "Being anxious is a very different bodily feeling compared to other states (e.g., feeling sexually aroused, relaxed or after exercise)",
  MINT_Awareness_ExAc_34 = "I can always accurately feel when I am about to burp",
  MINT_Awareness_ExAc_35 = "I can always accurately feel when I am about to fart",
  MINT_Awareness_ExAc_36 = "I can always accurately feel when I am about to sneeze",
  MINT_Sensitivity_Card_37 = "In general, I am very sensitive to changes in my heart rate",
  MINT_Sensitivity_Card_38 = "I can notice even very subtle changes in the way my heart beats",
  MINT_Sensitivity_Card_39 = "I often notice changes in my heart rate",
  MINT_Sensitivity_Resp_40 = "I can notice even very subtle changes in my breathing",
  MINT_Sensitivity_Resp_41 = "I am always very aware of how I am breathing, even when I am calm",
  MINT_Sensitivity_Resp_42 = "In general, I am very sensitive to changes in my breathing",
  # MINT_Sensitivity_Sign_43 = "When something important is happening in my life, I can feel immediately feel changes in my heart rate",
  # MINT_Sensitivity_Sign_44 = "When something important is happening in my life, I can immediately feel changes in my breathing",
  # MINT_Sensitivity_Sign_45 = "When something important is happening in my life, I can feel it in my body",
  MINT_Sensitivity_Gast_46 = "I can notice even very subtle changes in what my stomach is doing",
  MINT_Sensitivity_Gast_47 = "In general, I am very sensitive to what my stomach is doing",
  MINT_Sensitivity_Gast_48 = "I am always very aware of what my stomach is doing, even when I am calm",
  MINT_Sensitivity_Derm_49 = "In general, my skin is very sensitive",
  MINT_Sensitivity_Derm_50 = "My skin is susceptible to itchy fabrics and materials",
  MINT_Sensitivity_Derm_51 = "I can notice even very subtle stimulations to my skin (e.g., very light touches)"
  # MINT_Sensitivity_SexC_52 = "When I am sexually aroused, I often feel changes in the way my heart beats (e.g., faster or stronger)",
  # MINT_Sensitivity_SexC_53 = "When I am sexually aroused, I often feel changes in my breathing (e.g., faster, shallower, or less regular)",
  # MINT_Sensitivity_SexC_54 = "When I am sexually aroused, I often feel changes in my temperature (e.g., feeling warm or cold)"
)

cleanlabels <- function(x, qname=FALSE) {
  is_factor <- FALSE
  if(is.factor(x)) {
    is_factor <- TRUE
    vec <- x
    x <- levels(x)
  }

  x <- x |> 
    str_replace("MINT_", ifelse(qname, "MINT - ", "")) |> 
    str_replace("MAIA_", ifelse(qname, "MAIA - ", "")) |> 
    str_replace("_", " (") 
  x <- ifelse(str_detect(x, "\\("), paste0(x, ")"), x)

  if(is_factor) {
    levels(vec) <- x
    return(vec)
  }
  x
}
```

### Cluster Stability

#### Full

```{r}
#| eval: false

ega1 <- bootEGA(
  items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3, ncores = 4)

save(ega1, file="models/ega1.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega1.RData")

itemstability <- EGAnet::itemStability(ega1, IS.plot=FALSE)
plot(itemstability)
```

#### Final

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

items <- items |> 
  select(-contains("UrSe"))
```

```{r}
#| eval: false

ega2 <- bootEGA(
  items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3, ncores = 4)

save(ega2, file="models/ega2.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega2.RData")

# EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega2, IS.plot=FALSE)
plot(itemstability)
```

#### Item Table

```{r}
#| message: false

make_loadingtable <- function(x) {
  t <- as.data.frame(x) |>
    datawizard::data_addprefix("C")
  t$Cluster <- colnames(t)[max.col(t, ties.method='first')]

  t |>
    rownames_to_column(var="Item") |>
    rowwise() |>
    mutate(Max = max(c_across(-c(Item, Cluster)))) |>
    arrange(Cluster, desc(Max)) |>
    as.data.frame()
}

dimensions <- list(
  C08 = "CaCo",       # Cardiorespiratory Confusion
  C01 = "UrIn",       # Urointestinal Inaccuracy
  C02 = "Derm",       # Dermatoception
  C10 = "CaNo",       # Cardiorespiratory Noticing
  C13 = "Sati",       # Satiety Noticing

  C14 = "SexS",       # Sexual Arousal Sensitivity
  C07 = "StaS",       # State Specificity
  C09 = "SexO",       # Sexual Organs Sensitivity
  C03 = "ExAc",       # Expulsion Accuracy
  C06 = "RelA",       # Relaxation Awareness

  C05 = "Resp",       # Respiroception
  C11 = "Card",       # Cardioception
  C12 = "Gast",       # Gastroception
  C04 = "Olfa"       # Olfactory Compensation

  # C04 = "SexA",       # Sexual Arousal Awareness
  # C08 = "UrSe",       # Urointestinal Sensitivity
  # C12 = "Sign",       # Signalling
  # C07 = "SexC"        # Sexual Arousal Changes  
)

higher <- make_loadingtable(t(net.loads(ega2$EGA$higher_order, loading.method="revised")$std))
cluster_order <- names(higher)[!names(higher) %in% c("Item", "Cluster", "Max")]

lower <- make_loadingtable(net.loads(ega2$EGA$lower_order, loading.method="revised")$std) |>
  select(names(higher)) |>
  mutate(Cluster = fct_relevel(Cluster, cluster_order)) |>
  arrange(Cluster, desc(Max))

higher$Item <- paste0("M", higher$Item)
# higher$Label <- c("Interoceptive Deficit", "Interoceptive Awareness", "Interoceptive Sensitivity")
higher$Label <- ""
# higher$ID  <- ""
lower <- mutate(lower, Label = labels[paste0("MINT_", Item)], ID=Item, Item = 1:n())

rbind(higher, select(lower, -ID)) |>
# rbind(higher, lower) |>
  datawizard::data_relocate("Label", after = "Item") |>
  datawizard::data_rename(names(dimensions), dimensions) |>
  select(-Cluster, -Max) |> 
  mutate(`|` = "") |> 
  gt::gt() |>
  gt::tab_row_group(
    label = "Items",
    rows = 1:(nrow(lower) + nrow(higher))
  ) |>
  gt::tab_row_group(
    label = "Metaclusters",
    rows = 1:nrow(higher)
  ) |>
  gt::data_color(columns=-Item,
                 method = "numeric",
                 palette = c("red", "white", "green"),
                 domain = c(-1, 1)) |>
  gt::tab_style(
    style = gt::cell_text(size="small", style="italic"),
    locations = gt::cells_body(columns="Label", rows=c(4:45))
  ) |>
  gt::tab_style(
    style = list(gt::cell_text(weight="bold"),
                 gt::cell_fill(color = "#F5F5F5")),
    locations = list(
      gt::cells_row_groups(groups = "Items"),
      gt::cells_row_groups(groups = "Metaclusters")
    )
  ) |>
  gt::fmt_number(columns=-Item, decimals=2) |>
  gt::tab_header(
    title = gt::md("**Item Loadings**"),
    subtitle = "Node centrality"
  ) 
```

### Score Extraction

```{r}
df$MINT_Deficit_CaCo <- rowMeans(select(df, starts_with("MINT_Deficit_CaCo_")), na.rm = TRUE)
df$MINT_Deficit_Urin <- rowMeans(select(df, starts_with("MINT_Deficit_Urin_")), na.rm = TRUE)
# df$MINT_Deficit <- rowMeans(select(df, MINT_Deficit_CaCo, MINT_Deficit_Urin))

df$MINT_Sensitivity_Derm <- rowMeans(select(df, starts_with("MINT_Sensitivity_Derm_")), na.rm = TRUE)
df$MINT_Deficit_Sati <- rowMeans(select(df, starts_with("MINT_Deficit_Sati")), na.rm = TRUE)
df$MINT_Deficit_CaNo <- rowMeans(select(df, starts_with("MINT_Deficit_CaNo")), na.rm = TRUE)
# df$MINT_LackSensitivity <- rowMeans(select(df, MINT_Sensitivity_Derm, MINT_Deficit_Sati, MINT_Deficit_CaNo))

df$MINT_Awareness_SexS <- rowMeans(select(df, starts_with("MINT_Awareness_SexS")), na.rm = TRUE)
df$MINT_Awareness_StaS <- rowMeans(select(df, starts_with("MINT_Awareness_StaS")), na.rm = TRUE)
df$MINT_Awareness_SexO <- rowMeans(select(df, starts_with("MINT_Awareness_SexO")), na.rm = TRUE)
df$MINT_Awareness_ExaC <- rowMeans(select(df, starts_with("MINT_Awareness_ExaC")), na.rm = TRUE)
df$MINT_Awareness_RelA <- rowMeans(select(df, starts_with("MINT_Awareness_RelA")), na.rm = TRUE)
df$MINT_Awareness <- rowMeans(select(df, MINT_Awareness_SexS, MINT_Awareness_StaS, MINT_Awareness_SexO, MINT_Awareness_ExaC, MINT_Awareness_RelA))

df$MINT_Interoception_Card <- rowMeans(select(df, starts_with("MINT_Sensitivity_Card")), na.rm = TRUE)
df$MINT_Interoception_Resp <- rowMeans(select(df, starts_with("MINT_Sensitivity_Resp")), na.rm = TRUE)
df$MINT_Interoception_Gast <- rowMeans(select(df, starts_with("MINT_Sensitivity_Gast")), na.rm = TRUE)
df$MINT_Interoception <- rowMeans(select(df, MINT_Interoception_Card, MINT_Interoception_Resp, MINT_Interoception_Gast))

df$MINT_Deficit_Olfa <- rowMeans(select(df, starts_with("MINT_Deficit_Olfa")), na.rm = TRUE)
# df$MINT_Awareness_UrSe <- rowMeans(select(df, starts_with("MINT_Awareness_UrSe")), na.rm = TRUE)

# df$MINT_Awareness <- rowMeans(select(df, starts_with("MINT_Awareness"), -matches("[[:digit:]]")))

df <- df[names(df)[!grepl("MINT_.*[0-9]$", names(df))]]

minti <- data.frame(
  # MINT_Deficit = df$MINT_Deficit,
  MINT_Awareness = df$MINT_Awareness,
  MINT_Interoception = df$MINT_Interoception
)
```

```{r}
#| warning: false

make_cor <- function(df1, df2=NULL) {
  df1 |> 
    correlation::correlation(data2=df2, p_adjust="none", redundant=ifelse(is.null(df2), TRUE, FALSE)) |> 
    correlation::cor_sort() |>
    mutate(label = ifelse(p < .001, paste0(insight::format_value(r)), ""),
           Parameter1 = cleanlabels(Parameter1, qname=FALSE),
           Parameter2 = cleanlabels(Parameter2, qname=TRUE)) |> 
    ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
    geom_tile() +
    geom_text(aes(label=label), size=3) +
    scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust=1),
          legend.position = "none")
}

make_cor(select(df, starts_with("MINT"))) +
  labs(x="", y="", title="Correlation between MINT dimensions") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, 
                                   face=c("bold", "plain", "plain", "plain", "bold", "bold", "plain", "plain",
                                          "plain", "plain", "plain", "plain", "plain", "plain", "plain", "plain",
                                          "plain", "plain")))
```

## Convegent Validity

::: panel-tabset

### Interoception

```{r}
#| warning: false

make_cor <- function(df1, df2) {
  df1 |> 
    correlation::correlation(df2, p_adjust="none") |> 
    correlation::cor_sort() |>
    mutate(label = ifelse(p < .001, paste0(insight::format_value(r)), ""),
           Parameter1 = cleanlabels(Parameter1, qname=FALSE),
           Parameter2 = cleanlabels(Parameter2, qname=TRUE)) |> 
    ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
    geom_tile() +
    geom_text(aes(label=label), size=3) +
    scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust=1),
          legend.position = "none")
}

make_cor(select(df, starts_with("MINT"), -matches("[[:digit:]]")),
         select(df, starts_with("MAIA"), IAS, BPQ)) +
  labs(x="MINT", y="", title="Correlations between the MINT and other Interoception Questionnaires") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, 
                                   face=c("bold", "plain", "plain", "plain", "bold", "bold", "plain", "plain",
                                          "plain", "plain", "plain", "plain", "plain", "plain", "plain", "plain",
                                          "plain", "plain")))
```

```{r}
select(df, starts_with("MINT"), starts_with("MAIA"), IAS, BPQ) |> 
  parameters::factor_analysis(n=4, rotation="oblimin") |> 
  plot()
```

```{r}
# library(umap)
# 
# items_intero <- select(df, starts_with("MINT"), -matches("[[:digit:]]"), starts_with("MAIA"), IAS, BPQ)
# rez <- umap::umap(datawizard::data_transpose(items_intero[complete.cases(items_intero),]))
# 
# rez
# as.data.frame(rez$layout) |> 
#   rownames_to_column(var="Variable") |> 
#   separate(Variable, into=c("Questionnaire", "Dimension", "Sub"), fill="right") |> 
#   ggplot(aes(x=V1, y=V2, color=Questionnaire)) +
#   geom_point(size=3) +
#   scale_color_manual(values=colors)
```

### Alexithymia

```{r}
compare_models <- function(outcome="TAS_DIF") {

  # Individual predictors
  preds <- c("BPQ", "IAS", 
             names(select(df, starts_with("MINT"), -matches("[[:digit:]]"))),
             names(select(df, starts_with("MAIA"))))

  models <- list()
  for(p in preds) {
    models[[p]] <- lm(as.formula(paste0(outcome, " ~ ", p)), data=df)
  }
  perf <- compare_performance(models) |> 
    arrange(desc(R2))

  best_mint <- perf$Name[str_detect(perf$Name, "MINT_")][1]
  best_maia <- perf$Name[str_detect(perf$Name, "MAIA")][1]

  models <- list()
  df_pred <- data.frame()
  for(p in c(best_mint, best_maia, "BPQ", "IAS")) {
    models[[p]] <- lm(as.formula(paste0(outcome, " ~ ", p)), data=df)
    pred <- modelbased::estimate_relation(models[[p]], by=p, length=30) 
    names(pred)[1] <- "Value"
    pred$Predictor <- p
    df_pred <- rbind(df_pred, pred)
  }

  test <- test_performance(models)

  cols <- colors
  cols[[best_mint]] <- cols["MINT"]
  cols[[best_maia]] <- cols["MAIA"]

  p <- df_pred |>
    mutate(Questionnaire = case_when(
      str_detect(Predictor, "MINT") ~ "MINT", 
      str_detect(Predictor, "MAIA") ~ "MAIA",
      .default = Predictor)) |> 
    ggplot(aes(x=Value, y=Predicted)) +
    geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=Predictor), alpha=0.3) +
    geom_line(aes(color=Predictor)) +
    scale_fill_manual(values=cols) +
    scale_color_manual(values=cols) +
    facet_wrap(~Predictor, scales="free_x") +
    labs(y=outcome, x="Interoception Index")

  # Full Questionnaires
  m_mint <- lm(as.formula(paste0(outcome, " ~ .")),
              data = select(df, all_of(outcome), starts_with("MINT"), -matches("[[:digit:]]$"))) |> 
    step(direction = "both", trace=0)
  m_minti <- lm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(minti, df[outcome])) |> 
    step(direction = "both", trace=0)
  m_ias <- lm(as.formula(paste0(outcome, " ~ IAS")),
              data = df)
  m_maia <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), starts_with("MAIA"))) |> 
    step(direction = "both", trace=0)
  m_maiai <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), 
                              starts_with("MAIA"), 
                              -contains("NotWorrying"),
                              -contains("NotDistracting"))) |> 
    step(direction = "both", trace=0)
  m_bpq <- lm(as.formula(paste0(outcome, " ~ BPQ")),
                data = df)

  test_all <- test_performance(m_mint, m_minti, m_ias, m_maia, m_maiai, m_bpq)
  test_all$Predictors <- c(
    paste0(cleanlabels(find_predictors(m_mint)$conditional), collapse = ", "),
    paste0(cleanlabels(find_predictors(m_minti)$conditional), collapse = ", "),
    "", 
    paste0(cleanlabels(find_predictors(m_maia)$conditional), collapse = ", "),
    paste0(cleanlabels(find_predictors(m_maiai)$conditional), collapse = ", "),
    "")

  list(test=test,  test_all=test_all, p=p)
}

rez1 <- compare_models("TAS_DIF")
rez1$test
rez1$p
rez1$test_all

rez2 <- compare_models("TAS_DDF")
rez2$test
rez2$p
rez2$test_all

rez3 <- compare_models("TAS_EOT")
rez3$test
rez3$p
rez3$test_all
```

### Mood

```{r}
rez <- compare_models("LifeSatisfaction")
rez$test
rez$p
rez$test_all

rez <- compare_models("PHQ4_Anxiety")
rez$test
rez$p
rez$test_all

rez <- compare_models("PHQ4_Depression")
rez$test
rez$p
rez$test_all
```

### Emotions

```{r}
rez <- compare_models("ERS_Arousal")
rez$test
rez$p
rez$test_all

rez <- compare_models("ERS_Sensitivity")
rez$test
rez$p
rez$test_all

rez <- compare_models("ERS_Persistence")
rez$test
rez$p
rez$test_all
```

:::

## Predictive

### Mental Health

::: panel-tabset

#### Mood Disorders

```{r}
compare_predictive <- function(outcome="Disorders_Psychiatric_Mood") {
  m_mint <- glm(as.formula(paste0(outcome, " ~ .")),
              data = select(df, all_of(outcome), starts_with("MINT"), -matches("[[:digit:]]$")),
              family = "binomial") |> 
  step(direction = "both", trace=0)

  m_minti <- glm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(minti, df[outcome]),
              family = "binomial") |> 
    step(direction = "both", trace=0)

  m_ias <- glm(as.formula(paste0(outcome, " ~ IAS")),
              data = df,
              family = "binomial")

  m_maia <- glm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), starts_with("MAIA")),
                family = "binomial") |> 
    step(direction = "both", trace=0)

  m_maiai <- glm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), 
                              starts_with("MAIA"), 
                              -contains("NotWorrying"),
                              -contains("NotDistracting")),
                family = "binomial") |> 
    step(direction = "both", trace=0)

  m_bpq <- glm(as.formula(paste0(outcome, " ~ BPQ")),
                data = df,
                family = "binomial")

  test <- test_performance(m_mint, m_minti, m_ias, m_maia, m_maiai, m_bpq)

  get_auc <- function(m) {
    roc <- as.data.frame(performance_roc(m))
    auc <- bayestestR::area_under_curve(roc$Specificity, roc$Sensitivity)
    paste0("AUC = ", insight::format_percent(auc))
  }

  p <- rbind(
    mutate(as.data.frame(performance_roc(m_mint)), Predictor = "MINT"),
    mutate(as.data.frame(performance_roc(m_minti)), Predictor = "MINTi"),
    mutate(as.data.frame(performance_roc(m_ias)), Predictor = "IAS"),
    mutate(as.data.frame(performance_roc(m_maia)), Predictor = "MAIA"),
    mutate(as.data.frame(performance_roc(m_maiai)), Predictor = "MAIAi"),
    mutate(as.data.frame(performance_roc(m_bpq)), Predictor = "BPQ")
  ) |> 
      ggplot(aes(x=Specificity)) +
      geom_abline(intercept=0, slope=1, color="gray") +
      geom_line(aes(y=Sensitivity, color=Predictor), linewidth=2, alpha=0.9) +
      geom_label(data=data.frame(
        label = c(get_auc(m_mint), get_auc(m_minti), get_auc(m_ias), get_auc(m_maia), get_auc(m_maiai), get_auc(m_bpq)),
        Predictor = c("MINT", "MINTi", "IAS", "MAIA", "MAIAi", "BPQ"),
        x = 0.25,
        y = c(0.65, 0.60, 0.55, 0.45, 0.35, 0.20)), 
        aes(x=x, y=y, label=label), color=colors) +
      labs(x= "1 - Specificity (False Positive Rate)", y="Sensitivity (True Positive Rate)",
           color="Questionnaire Version") +
      scale_color_manual(values=colors) +
      scale_x_continuous(labels=scales::percent) +
      scale_y_continuous(labels=scales::percent) +
      theme_minimal()

  list(m_mint=m_mint, m_ias=m_ias, m_maia=m_maia, m_maiai=m_maiai, test=test, p=p)
}
```

```{r}
rez <- compare_predictive("Disorders_Psychiatric_Mood")
rez$p
```

```{r}
rez <- compare_predictive("Disorders_Psychiatric_MoodTreatment")
rez$p
```

#### Anxiety

```{r}
rez <- compare_predictive("Disorders_Psychiatric_Anxiety")
rez$p
```

```{r}
rez <- compare_predictive("Disorders_Psychiatric_AnxietyTreatment")
rez$p
```

#### Eating

```{r}
rez <- compare_predictive("Disorders_Psychiatric_Eating")
rez$p
```

#### Addiction

```{r}
rez <- compare_predictive("Disorders_Psychiatric_Addiction")
rez$p
```

#### ADHD

```{r}
rez <- compare_predictive("Disorders_Psychiatric_ADHD")
rez$p
```

#### Autism

```{r}
rez <- compare_predictive("Disorders_Psychiatric_Autism")
rez$p
```

#### Borderline

```{r}
# rez <- compare_predictive("Disorders_Psychiatric_Borderline")
# rez$p
```

::::

### Somatic Issues

::: panel-tabset

#### Autonomic

```{r}
rez <- compare_predictive("Disorders_Somatic_Autonomic")
rez$p
```

#### Functional

```{r}
rez <- compare_predictive("Disorders_Somatic_Functional")
rez$p
```

#### Hyperreactivity

```{r}
rez <- compare_predictive("Disorders_Somatic_Hyperreactivity")
rez$p
```

#### Immune

```{r}
rez <- compare_predictive("Disorders_Somatic_Immune")
rez$p
```

#### Cardio

```{r}
rez <- compare_predictive("Disorders_Somatic_Cardio")
rez$p
```

#### Connective

```{r}
rez <- compare_predictive("Disorders_Somatic_Connective")
rez$p
```

:::

### Biofeedback Usage

```{r}
model <- glm(as.formula(paste0("Wearables_Number ~ .")),
              data = select(df, Wearables_Number, starts_with("MINT"), -matches("[[:digit:]]")),
              family = "poisson") |> 
  step(direction = "both", trace=0)

parameters::parameters(model)

model <- glm(as.formula(paste0("Wearables_Heart ~ .")),
              data = df |>
              mutate(Wearables_Heart = Wearables_Heart != "Not owning") |> 
              select(Wearables_Heart, starts_with("MINT"), -matches("[[:digit:]]")),
              family = "binomial") |> 
  step(direction = "both", trace=0, k = log(nrow(df)))
summary(model)

model <- lm(as.formula(paste0("Wearables_HeartImportance ~ .")),
              data = df |>
              filter(!is.na(Wearables_HeartImportance)) |> 
              select(Wearables_HeartImportance, starts_with("MINT"), -matches("[[:digit:]]"))) |> 
  step(direction = "both", trace=0, k = log(nrow(filter(df, !is.na(Wearables_HeartImportance)))))

parameters::parameters(model)
```

## Exploratory

### CEFSA

```{r}
make_cor(select(df, starts_with("MINT"), -matches("[[:digit:]]")), 
         select(df, starts_with("CEFSA")))

make_cor(select(df, starts_with("MAIA"), IAS, BPQ), 
         select(df, starts_with("CEFSA")))
```

### Emotion Regulation

```{r}
make_cor(select(df, starts_with("MINT"), -matches("[[:digit:]]")), 
         select(df, starts_with("CERQ")))

make_cor(select(df, starts_with("MAIA"), IAS, BPQ), 
         select(df, starts_with("CERQ")))
```

### Primals

```{r}
make_cor(select(df, starts_with("MINT"), -matches("[[:digit:]]")), 
         select(df, starts_with("PI")))

make_cor(select(df, starts_with("MAIA"), IAS, BPQ), 
         select(df, starts_with("PI")))
```