### 2025-03-09 study2sona/analysis/analysis.qmd

---
title: "Data Quality - Data Analysis"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(EGAnet)
library(tidygraph)
library(ggraph)

set.seed(42)
```

```{r}
#| code-fold: false

df <- read.csv("../../study2/data/rawdata_participants.csv") |> 
  mutate(Condition = fct_relevel(Condition, "Prolific", "Attention Checks", "Quiz", "Control"))
```

```{r}

```

## Attention Checks

```{r}
dfchecks <- df |>
  dplyr::mutate(
    # "I can always accurately answer to the extreme left on this question to show that I am reading it"
    A1 = ifelse(MINT_AttentionCheck_1 == 0, 0, 1),
    # "I notice that I am being asked to respond all the way to the right"
    A2 = ifelse(MAIA_AttentionCheck_1 == 6, 0, 1),
    # "I can always accurately choose the lowest option"
    A3 = ifelse(IAS_AttentionCheck_1 == 1, 0, 1),
    # "Respond all the way to the right."
    A4 = ifelse(BodyAwareness_AttentionCheck_1 == 5, 0, 1),
    # "I am able to respond all the way to the left"
    A5 = ifelse(TAS_AttentionCheck_1 == 1, 0, 1),
    # "On the whole, I know I must press the highest option"
    A6 = ifelse(PI18_AttentionCheck_1 == 5, 0, 1),
    # "I feel that to show I'm being attentive I will press the lowest option"
    A7 = ifelse(CEFSA_AttentionCheck_1 == 0, 0, 1),
    .keep = "none"
  ) 
dfchecks$TotalFailed <- rowSums(dfchecks)
df$TotalFailed <- dfchecks$TotalFailed
df$Failed <- ifelse(dfchecks$TotalFailed >= 1, 1, 0)

dfchecks |>
  mutate(Total = as.factor(paste0(TotalFailed, "/8"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

```{r}
df |> 
  summarize(Failed = sum(Failed), N = n(), p = Failed / N, .by="Condition") |> 
  gt::gt()
```

### Failed at least one

```{r}
m1 <- glm(Failed ~ Prolific, 
          data=mutate(df, Prolific = ifelse(Condition == "Prolific", "Prolific", "SONA")), 
          family="binomial")
# m1 <- brms::brm(Failed ~ Condition, data=df, family="poisson", backend="cmdstanr")

modelbased::estimate_means(m1, by="Prolific", keep_iterations=10) |> 
  plot()

modelbased::estimate_contrasts(m1)
```

```{r}
m1 <- glm(Failed ~ Condition, data=df, family="binomial")
# m1 <- brms::brm(Failed ~ Condition, data=df, family="poisson", backend="cmdstanr")

modelbased::estimate_means(m1, by="Condition", keep_iterations=10) |> 
  plot()

modelbased::estimate_contrasts(m1)
```

### Number

```{r}
m1 <- glm(TotalFailed ~ Condition, data=df, family="poisson")
m2 <- glmmTMB::glmmTMB(TotalFailed ~ Condition,
                      data=df,
                      ziformula=~Condition,
                      family="poisson")
m3 <- glmmTMB::glmmTMB(TotalFailed ~ Condition,
                      data=df,
                      ziformula=~Condition,
                      family=glmmTMB::truncated_nbinom2)

compare_performance(m1, m2, m3)
test_performance(m1, m2, m3)

modelbased::estimate_means(m1, by="Condition") |> 
  plot()
```