### 2025-07-01 study2/analysis/1_cleaning.qmd

---
title: "Interoception Scale (Study 2) - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
---

## Data Preparation

```{r}
#| warning: false
#| code-fold: false
#| error: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)

df <- read.csv("../data/rawdata_participants.csv")
```

## Recruitment History

```{r}
#| message: false

# Consecutive count of participants per day (as area)
df |>
  mutate(Date = as.Date(Experiment_StartDate, format = "%Y-%m-%d %H:%M:%S")) |> 
  summarize(N = n(), .by=c("Date", "Recruitment")) |> 
  complete(Date, Recruitment, fill = list(N = 0)) |> 
  mutate(N = cumsum(N), .by="Recruitment") |>
  ggplot(aes(x = Date, y = N)) +
  geom_area(aes(fill=Recruitment)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = "Recruitment History",
    x = "Date",
    y = "Total Number of Participants"
  ) +
  see::theme_modern()
```

```{r}
#| message: false

# Table
summarize(df, N = n(), .by=c("Recruitment")) |> 
  arrange(desc(N)) |> 
  gt::gt() |> 
  gt::grand_summary_rows(columns = "N", fns = Total ~ sum(.)) |> 
  gt::opt_stylize(style = 2, color = "gray") |> 
  gt::tab_header("Number of participants per recruitment source") 
```

## Feedback

### Experiment Enjoyment

```{r}
df |> 
  summarise(n = n(), .by=c("Experiment_Enjoyment")) |>
  filter(!is.na(Experiment_Enjoyment)) |> 
  mutate(n = n / sum(n),
         Experiment_Enjoyment = fct_rev(as.factor(Experiment_Enjoyment))) |>
  ggplot(aes(y = n, x = 1, fill = Experiment_Enjoyment)) +
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual(values=c("green", "yellowgreen",  "yellow", "orange", "red")) +
  coord_flip() +
  scale_x_continuous(expand=c(0, 0))  +
  scale_y_continuous(labels = scales::percent)  +
  labs(title="Experiment Enjoyment",
       subtitle="Proportion of participants by condition") + 
  guides(fill = guide_legend(reverse=TRUE)) +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "top",
    legend.title = element_blank()) 
```

### Comments

```{r}
data.frame(Comments = df$Experiment_Feedback) |> 
  filter(!tolower(Comments) %in% c(NA, "no", "none", "no thank you", "no not at all", "no thanks", "no comment", "nil", "n.a", "n/a", "na", "not really", "thank you", "nothing", "no not at the moment", "none for now", "thank you!", "thanks", "not really.")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) 
```

## Exclusions

```{r}
outliers <- list()
```

### Attention Checks

```{r}
dfchecks <- df |>
  dplyr::mutate(
    # "I can always accurately answer to the extreme left on this question to show that I am reading it"
    A1 = ifelse(MINT_AttentionCheck_1 == 0, 0, 1),
    # "I notice that I am being asked to respond all the way to the right"
    A2 = ifelse(MAIA_AttentionCheck_1 == 6, 0, 1),
    # "I can always accurately choose the lowest option"
    A3 = ifelse(IAS_AttentionCheck_1 == 1, 0, 1),
    # "Respond all the way to the right."
    A4 = ifelse(BodyAwareness_AttentionCheck_1 == 5, 0, 1),
    # "I am able to respond all the way to the left"
    A5 = ifelse(TAS_AttentionCheck_1 == 1, 0, 1),
    # "On the whole, I know I must press the highest option"
    A6 = ifelse(PI18_AttentionCheck_1 == 5, 0, 1),
    # "I feel that to show I'm being attentive I will press the lowest option"
    A7 = ifelse(CEFSA_AttentionCheck_1 == 0, 0, 1),
    .keep = "none"
  ) 
dfchecks$Total <- rowSums(dfchecks)

dfchecks |>
  mutate(Total = as.factor(paste0(Total, "/8"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

```{r}
outliers$attentionchecks <- df$Participant[dfchecks$Total >= 1]
```

We removed `r length(outliers$attentionchecks)` (`r insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for having failed at least 1 attention check (out of 8).

### Experiment Duration

```{r}
dfchecks$Duration <- df$Experiment_Duration
dfchecks$Outlier <- ifelse(dfchecks$Total >= 1, 1, 0)
dfchecks <- filter(dfchecks, Duration < 45)

m <- mgcv::gam(Outlier ~ s(Duration), data = dfchecks, family = "binomial")

estimate_relation(m, length=50) |> 
  ggplot(aes(x = Duration, y = Predicted)) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2) +
  geom_line() +
  geom_vline(xintercept=10, linetype="dashed", color="red") +
  theme_minimal() +
  ggside::geom_xsidedensity(data=mutate(dfchecks, 
                                        Outlier=ifelse(Outlier==1, "Failed attention check", "Valid")), 
                            aes(fill=Outlier), alpha=0.3) +
  ggside::theme_ggside_void() +
  labs(title = "Experiment Duration",
       subtitle = "Predicted probability of failing attention checks by duration",
       x = "Experiment Duration (minutes)",
       y = "Probability of failing attention checks") 
```

```{r}
outliers$duration <- as.character(df[df$Experiment_Duration < 10, "Participant"])
outliers$duration <- outliers$duration[!outliers$duration %in% outliers$attentionchecks]
```

We removed `r length(outliers$duration)` (`r insight::format_percent(length(outliers$duration) / nrow(df))`) participants for having completed the experiment in less than 5 minutes.

### Multivariate Distance

```{r}
# Compute distance
dfoutlier <- performance::check_outliers(select(df, contains("MAIA_")), 
                                         method=c("optics"),
                                         threshold=list(optics=2.5, optics_xi=0.01)) |> 
  as.data.frame() |> 
  mutate(Participant = fct_reorder(df$Participant, Distance_OPTICS),
         Outlier_AttentionCheck = ifelse(Participant %in% outliers$attentionchecks, 1, 0),
         Outlier_Duration = ifelse(Participant %in% outliers$duration, 1, 0),
         Outlier = ifelse(Outlier_AttentionCheck == 1, "Failed Attention Checks", "Passed"),
         Outlier = ifelse(Outlier == "Passed" & Outlier_Duration == 1, "Duration", Outlier))

outliers$distance <- as.character(dfoutlier[dfoutlier$Distance_OPTICS < 3.3, "Participant"])
outliers$distance <- outliers$distance[!outliers$distance %in% c(outliers$attentionchecks, outliers$duration)]

p1 <- dfoutlier |> 
  ggplot(aes(x=Participant, y=Distance_OPTICS)) +
  geom_bar(aes(fill=Outlier), stat="identity") +
  geom_hline(yintercept = 3.3, linetype="dashed", color="red") +
  labs(fill = "Status") +
  scale_fill_manual(values=c("Duration" = "purple", "Failed Attention Checks"="red", "Passed"="grey")) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.major.x = element_blank())

m <- mgcv::gam(Outlier_AttentionCheck ~ s(Distance_OPTICS), data = dfoutlier, family = "binomial")
# parameters::parameters(m)

p2 <- estimate_relation(m, length=30) |> 
  ggplot(aes(x=Distance_OPTICS, y=Predicted)) +
  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha=0.2) +
  geom_line() +
  geom_vline(xintercept=3.3, linetype="dashed", color="red") +
  geom_hline(yintercept=1/3, linetype="dashed", color="black") +
  theme_minimal()

p1 | p2
```

We removed `r length(outliers$distance)` (`r insight::format_percent(length(outliers$distance) / nrow(df))`) participants based on multivariate distance.

```{r}
#| code-fold: false

df <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$duration, outliers$distance))
```

## Dimension Computation

```{r}
#| code-fold: true

compute_and_remove <- function(df, name="BodyAwareness", pattern=name, method="mean") {
  items <- select(df, starts_with(pattern), -contains("AttentionCheck"))
  df <- df[!names(df) %in% names(items)]
  if(method == "mean") {
    df[[name]] <- rowMeans(items, na.rm=TRUE)
  } else {
    df[[name]] <- rowSums(items, na.rm=TRUE)
  }
  df
}
```

```{r}
#| code-fold: false

df <- compute_and_remove(df, name="IAS", method="sum")

df <- compute_and_remove(df, name="BPQ", pattern="BodyAwareness", method="mean")

df[grepl("^MAIA_.*_R$", names(df))] <- 6 - df[grepl("^MAIA_.*_R$", names(df))]  # Reverse
df <- compute_and_remove(df, name="MAIA_AttentionRegulation", method="mean")
df <- compute_and_remove(df, name="MAIA_BodyListening", method="mean")
df <- compute_and_remove(df, name="MAIA_EmotionalAwareness", method="mean")
df <- compute_and_remove(df, name="MAIA_NotDistracting", method="mean")
df <- compute_and_remove(df, name="MAIA_Noticing", method="mean")
df <- compute_and_remove(df, name="MAIA_NotWorrying", method="mean")
df <- compute_and_remove(df, name="MAIA_SelfRegulation", method="mean")
df <- compute_and_remove(df, name="MAIA_Trusting", method="mean")
```

```{r}
#| code-fold: false

df <- compute_and_remove(df, name="PHQ4_Anxiety", method="sum")
df <- compute_and_remove(df, name="PHQ4_Depression", method="sum")

# TODO: check if any reversed items in the following questionnaires
df <- compute_and_remove(df, name="TAS_EOT", method="mean")
df <- compute_and_remove(df, name="TAS_DDF", method="mean")
df <- compute_and_remove(df, name="TAS_DIF", method="mean")

df <- compute_and_remove(df, name="ERS_Arousal", method="mean")
df <- compute_and_remove(df, name="ERS_Sensitivity", method="mean")
df <- compute_and_remove(df, name="ERS_Persistence", method="mean")

df <- compute_and_remove(df, name="CERQ_Acceptance", method="mean")
df <- compute_and_remove(df, name="CERQ_PositiveRefocusing", method="mean")
df <- compute_and_remove(df, name="CERQ_PositiveReappraisal", method="mean")
df <- compute_and_remove(df, name="CERQ_SelfBlame", method="mean")
df <- compute_and_remove(df, name="CERQ_OtherBlame", method="mean")
df <- compute_and_remove(df, name="CERQ_Rumination", method="mean")
df <- compute_and_remove(df, name="CERQ_Catastrophizing", method="mean")
df <- compute_and_remove(df, name="CERQ_RefocusPlanning", method="mean")
df <- compute_and_remove(df, name="CERQ_Perspective", method="mean")

df <- compute_and_remove(df, name="CEFSA_Body", method="mean")
df <- compute_and_remove(df, name="CEFSA_Agency", method="mean")
df <- compute_and_remove(df, name="CEFSA_Emotion", method="mean")
df <- compute_and_remove(df, name="CEFSA_Familiarity", method="mean")
df <- compute_and_remove(df, name="CEFSA_Connection", method="mean")
df <- compute_and_remove(df, name="CEFSA_Reality", method="mean")
df <- compute_and_remove(df, name="CEFSA_Self", method="mean")

df[grepl("^PI18.*_R$", names(df))] <- 6 - df[grepl("^PI18.*_R$", names(df))]  # Reverse
df$PI_Alive <- rowMeans(df[grepl("^PI18.*A_", names(df))], na.rm=TRUE)
df <- compute_and_remove(df, name="PI_Safe", pattern="PI18_GS", method="mean")
df <- compute_and_remove(df, name="PI_Enticing", pattern="PI18_GE", method="mean")
df$PI_Good <- rowMeans(select(df, PI_Safe, PI_Enticing), na.rm=TRUE)
df <- select(df, -starts_with("PI18_"))
```

## Final Sample

::: {.panel-tabset}

### Summary

We removed `r sum(is.na(df$MINT_Awareness_RelA_29))` participants with missing data for the MINT due to a technical error.

```{r}
#| code-fold: false

df <- filter(df, !is.na(df$MINT_Awareness_RelA_29))
```

The final sample includes `r report::report_participants(df, age = "Age", sex = NA, gender = "Gender")`.

### Age and Gender

```{r p_demographics_age_gender}
p_age <- df |>
  ggplot(aes(x = Age, fill = Gender)) + 
  geom_histogram(data=df, aes(x = Age, fill=Gender), binwidth = 2) +
  geom_vline(xintercept = mean(df$Age), color = "red", linewidth=1.5, linetype="dashed") +
  scale_fill_viridis_d() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(20, max(df$Age), by = 10 )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

p_age
```

### Education

```{r p_demographics_education}
#| warning: false
#| error: false

p_edu <- df |>
  mutate(Student = ifelse(is.na(Student), FALSE, Student),
         Education = fct_relevel(Education, "High school", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Student)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level", fill = "Is currently a student") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

p_edu
```

### Ethnicity

```{r p_demographics_ethnicity}
#| warning: false
#| error: false

p_eth <- df |>
  filter(!is.na(Ethnicity)) |> 
  ggplot(aes(x = "", fill = Ethnicity)) +
  geom_bar() +
  coord_polar("y") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )

p_eth
```

### Country

```{r p_demographics_map}
p_map <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_map
```

```{r}
sort(table(df$Country)) |> 
  as.data.frame() |>
  rename(Country = Var1) |> 
  arrange(desc(Freq)) |> 
  gt::gt()
```

### BMI

```{r}
p_bmi <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=BMI)) +
  annotate("rect", xmin=10, xmax=18.5, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  annotate("rect", xmin=18.5, xmax=25, ymin=0, ymax=Inf, fill="green", alpha=0.25) +
  annotate("rect", xmin=25, xmax=30, ymin=0, ymax=Inf, fill="yellow", alpha=0.25) +
  annotate("rect", xmin=30, xmax=40, ymin=0, ymax=Inf, fill="orange", alpha=0.25) +
  annotate("rect", xmin=40, xmax=Inf, ymin=0, ymax=Inf, fill="red", alpha=0.1) +
  geom_density(aes(color=Gender, fill=Gender), alpha=0.5, linewidth=2) +
  labs(title = "BMI distribution") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank())

p_bmi_a <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=Age, y=BMI, color=Gender)) +
  geom_point(size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender), alpha=0.3, method = 'loess', formula = 'y ~ x') +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none")

p_bmi_b <- df |> 
  filter(!is.na(BMI), Gender != "Other") |> 
  ggplot(aes(x=Physical_Active, y=BMI)) +
  geom_jitter(aes(color=Gender), width=0.1, size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender, color=Gender), alpha=0.3, method = 'lm', formula = 'y ~ x') +
  labs(x = "Physically Active", y = "") +
  scale_x_continuous(breaks=0:7) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank()) +
  ggside::geom_xsidehistogram(aes(fill=Gender), bins = 30, position="dodge") +
  ggside::theme_ggside_void()

df$Physical_Workout[df$Physical_Workout > 25] <- NA

p_bmi_c <- df |> 
  filter(!is.na(BMI) & !is.na(Physical_Workout), Gender != "Other") |> 
  ggplot(aes(x=Physical_Workout, y=BMI)) +
  geom_point(aes(color=Gender), size=3, alpha=0.5) +
  geom_smooth(aes(fill=Gender, color=Gender), alpha=0.3, method = 'lm', formula = 'y ~ x') +
  labs(x = "Hours of Active Workout", y = "") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_color_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  theme_minimal() +
  theme(legend.position = "none") +
  ggside::geom_xsidehistogram(aes(fill=Gender), bins = 30, position="dodge") +
  ggside::theme_ggside_void()

p_bmi / (p_bmi_a | p_bmi_b | p_bmi_c)
```

### Mental Health

```{r}
df$Disorders_Psychiatric_Mood <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)
df$Disorders_Psychiatric_MoodTreatment <- ifelse(
  df$Disorders_Psychiatric_Mood & !is.na(df$Disorders_PsychiatricTreatment) & str_detect(df$Disorders_PsychiatricTreatment, "Mood|Antidepressant|Anxiolytic|Psychotherapy"), TRUE, FALSE)

df$Disorders_Psychiatric_Anxiety <- ifelse(str_detect(df$Disorders_Psychiatric, "Panic|GAD|Social Phobia|Phobia"), TRUE, FALSE)
df$Disorders_Psychiatric_AnxietyTreatment <- ifelse(
  df$Disorders_Psychiatric_Anxiety & !is.na(df$Disorders_PsychiatricTreatment) & str_detect(df$Disorders_PsychiatricTreatment, "Anxiolytic|Psychotherapy|Mindfulness"), TRUE, FALSE)

df$Disorders_Psychiatric_Eating <- ifelse(str_detect(df$Disorders_Psychiatric, "Eating"), TRUE, FALSE)
df$Disorders_Psychiatric_Addiction <- ifelse(str_detect(df$Disorders_Psychiatric, "Addiction"), TRUE, FALSE)
df$Disorders_Psychiatric_Borderline <- ifelse(str_detect(df$Disorders_Psychiatric, "BPD"), TRUE, FALSE)

df$Disorders_Psychiatric_Autism <- ifelse(str_detect(df$Disorders_Psychiatric, "ASD"), TRUE, FALSE)
df$Disorders_Psychiatric_ADHD <- ifelse(str_detect(df$Disorders_Psychiatric, "ADHD"), TRUE, FALSE)

# df$Disorders_Psychiatric_Other <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)

select(df, Participant, Gender, 
             Disorders_Psychiatric_Mood, Disorders_Psychiatric_MoodTreatment,
             Disorders_Psychiatric_Anxiety, Disorders_Psychiatric_AnxietyTreatment,
             Disorders_Psychiatric_Eating, Disorders_Psychiatric_Addiction, Disorders_Psychiatric_Borderline,
             Disorders_Psychiatric_Autism, Disorders_Psychiatric_ADHD) |> 
  pivot_longer(cols = starts_with("Disorders_"), names_to = "Disorder", values_to = "Value") |> 
  mutate(Disorder = str_remove_all(Disorder, fixed("Disorders_Psychiatric_")),
         Disorder = str_replace(Disorder, "MoodTreatment", "Mood Disorder (with treatment)"),
         Disorder = str_replace(Disorder, "AnxietyTreatment", "Anxiety (with treatment)"),
         Disorder = str_replace(Disorder, "Mood$", "Mood Disorder")) |> 
  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Disorder")) |> 
  mutate(N_tot = sum(N), .by="Disorder") |>
  mutate(Disorder = fct_reorder(Disorder, desc(N_tot)),
         Treatment = ifelse(str_detect(Disorder,  "Treatment"), "With treatment", "Without treatment")) |>
  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

### Somatic

```{r}
somatic <- data.frame(
  Hypermobility = ifelse(str_detect(df$Disorders_Somatic, "Hypermobility Syndrome"), TRUE, FALSE),
  Fibromyalgia = ifelse(str_detect(df$Disorders_Somatic, "Fibromyalgia"), TRUE, FALSE),
  ChronicFatigue = ifelse(str_detect(df$Disorders_Somatic, "Chronic Fatigue Syndrome"), TRUE, FALSE),
  ChronicPain = ifelse(str_detect(df$Disorders_Somatic, "Chronic Pain Syndrome"), TRUE, FALSE),
  BackPain = ifelse(str_detect(df$Disorders_Somatic, "Back Pain"), TRUE, FALSE),
  MuscleTension = ifelse(str_detect(df$Disorders_Somatic, "Muscle Tension"), TRUE, FALSE),
  SkinRashes = ifelse(str_detect(df$Disorders_Somatic, "Skin Rashes"), TRUE, FALSE),
  Eczema = ifelse(str_detect(df$Disorders_Somatic, "Eczema"), TRUE, FALSE),
  Psoriasis = ifelse(str_detect(df$Disorders_Somatic, "Psoriasis"), TRUE, FALSE),
   # Dermatological = ifelse(str_detect(df$Disorders_Somatic, "Other Dermatological"), TRUE, FALSE),
  # Musculoskeletal = ifelse(str_detect(df$Disorders_Somatic, "Other Musculoskeletal"), TRUE, FALSE),
  Sjogrens = ifelse(str_detect(df$Disorders_Somatic, "Sjogren"), TRUE, FALSE),
  ChestPain = ifelse(str_detect(df$Disorders_Somatic, "Chest Pain"), TRUE, FALSE),
  Arrhythmia = ifelse(str_detect(df$Disorders_Somatic, "Cardiac Arrhythmia"), TRUE, FALSE),
  Hypertension = ifelse(str_detect(df$Disorders_Somatic, "Hypertension"), TRUE, FALSE),
  Hypotension = ifelse(str_detect(df$Disorders_Somatic, "Hypotension"), TRUE, FALSE),
  IBS = ifelse(str_detect(df$Disorders_Somatic, "IBS"), TRUE, FALSE),
  GERD = ifelse(str_detect(df$Disorders_Somatic, "GERD"), TRUE, FALSE),
  Crohn = ifelse(str_detect(df$Disorders_Somatic, "Crohn"), TRUE, FALSE),
  UlcerativeColitis = ifelse(str_detect(df$Disorders_Somatic, "Ulcerative Colitis"), TRUE, FALSE),
  Celiac = ifelse(str_detect(df$Disorders_Somatic, "Celiac Disease"), TRUE, FALSE),
  Gluten = ifelse(str_detect(df$Disorders_Somatic, "Gluten"), TRUE, FALSE),
  Lactose = ifelse(str_detect(df$Disorders_Somatic, "Lactose"), TRUE, FALSE),
  ShortBreath = ifelse(str_detect(df$Disorders_Somatic, "Shortness of Breath"), TRUE, FALSE),
  Asthma = ifelse(str_detect(df$Disorders_Somatic, "Asthma"), TRUE, FALSE),
  COPD = ifelse(str_detect(df$Disorders_Somatic, "COPD"), TRUE, FALSE),
  SleepApnea = ifelse(str_detect(df$Disorders_Somatic, "Sleep Apnea"), TRUE, FALSE),
  Bronchitis = ifelse(str_detect(df$Disorders_Somatic, "Chronic Bronchitis"), TRUE, FALSE),
  Nausea = ifelse(str_detect(df$Disorders_Somatic, "Nausea/Vomiting"), TRUE, FALSE),
  Dizziness = ifelse(str_detect(df$Disorders_Somatic, "Dizziness/Lightheadedness"), TRUE, FALSE),
  Migraine = ifelse(str_detect(df$Disorders_Somatic, "Migraine"), TRUE, FALSE),
  Neuropathy = ifelse(str_detect(df$Disorders_Somatic, "Neuropathy"), TRUE, FALSE),
  Epilepsy = ifelse(str_detect(df$Disorders_Somatic, "Epilepsy"), TRUE, FALSE),
  MS = ifelse(str_detect(df$Disorders_Somatic, "Multiple Sclerosis"), TRUE, FALSE),
  FrequentUrination = ifelse(str_detect(df$Disorders_Somatic, "Frequent Urination"), TRUE, FALSE),
  Endometriosis = ifelse(str_detect(df$Disorders_Somatic, "Endometriosis"), TRUE, FALSE),
  Cystitis = ifelse(str_detect(df$Disorders_Somatic, "Interstitial Cystitis"), TRUE, FALSE),
  PelvicPain = ifelse(str_detect(df$Disorders_Somatic, "Chronic Pelvic Pain Syndrome"), TRUE, FALSE)
)

somatic |> 
  select(where(~n_distinct(.) > 1)) |> 
  correlation::correlation(redundant = TRUE, p_adjust="none") |>
  correlation::cor_sort() |> 
  mutate(label = ifelse(p < .001, "*", ""),
         label = ifelse(r == 1, "", label)) |> 
  ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
  geom_tile() +
  geom_text(aes(label=label), size=3) +
  scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
  labs(title = "Co-occurence Matrix of Somatic Disorders") +
  theme_minimal()  +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_blank())

# pca <- as.data.frame(prcomp(somatic)$rotation[, 1:5])
# pca$PC <- colnames(pca)[max.col(pca, ties.method='first')]
# 
# pca |> 
#   rownames_to_column(var="Disorder") |> 
#   pivot_longer(-c(Disorder, PC)) |> 
#   filter(PC == name) |> 
#   arrange(name, desc(value)) |> 
#   ggplot(aes(x=value, y=Disorder, fill=name)) +
#   geom_bar(stat = "identity") +
#   facet_wrap(~name, scale="free_y")
```

```{r}
#| fig-keep: "first"

# somatic |> 
#   select(where(~n_distinct(.) > 1)) |> 
#   EGAnet::bootEGA(
#     corr="pearson",
#     EGA.type = "EGA.fit",
#     model = "glasso", 
#     algorithm = "leiden",
#     seed=42)
```

To facilitate mechanistic interpretation and reduce redundancy, self-reported somatic symptoms were grouped into four non-overlapping clusters based on shared physiological pathways and known etiological mechanisms. The **Afferent Sensitivity** cluster included conditions associated with heightened interoceptive awareness and neurogenic excitability, such as migraine, neuropathy, dizziness, nausea, muscle tension, epilepsy, and frequent urination. The **Central Sensitization** cluster comprised syndromes characterized by chronic pain and fatigue, likely reflecting central amplification of sensory signals and HPA-axis dysregulation; this included fibromyalgia, chronic fatigue syndrome, chronic pain, back pain, pelvic pain, endometriosis, irritable bowel syndrome (IBS), and interstitial cystitis. The **Autonomic Dysfunction** cluster captured disorders linked to dysregulation of the autonomic and cardiopulmonary systems, including joint hypermobility, cardiac arrhythmia, chest pain, shortness of breath, hypo-/hypertension, sleep apnea, chronic obstructive pulmonary disease (COPD), and chronic bronchitis. Finally, the **Immune-Inflammatory** cluster encompassed conditions associated with immune dysregulation, barrier dysfunction, and gut-brain axis disturbance, such as eczema, psoriasis, skin rashes, asthma, celiac disease, gluten and lactose sensitivity, inflammatory bowel diseases (Crohn’s disease and ulcerative colitis), gastroesophageal reflux disease (GERD), multiple sclerosis, and Sjögren’s syndrome. Each symptom was assigned to a single cluster to preserve orthogonality and support subsequent analyses linking symptom profiles to interoceptive processes.

```{r}
# df$Disorders_Somatic_Number <- ifelse(df$Disorders_Somatic == "", 0, str_count(df$Disorders_Somatic, ";")+1)

# Cluster A: Afferent Sensitivity (neurogenic excitability, interoceptive hypervigilance)
df$Disorders_Somatic_AfferentSensitivity_N <- rowSums(somatic[, c(
  "Migraine", "Neuropathy", "MuscleTension", "Dizziness", 
  "Nausea", "Epilepsy", "FrequentUrination"
)])
df$Disorders_Somatic_AfferentSensitivity <- df$Disorders_Somatic_AfferentSensitivity_N >= 1

# Cluster B: Central Sensitization (pain and fatigue syndromes)
df$Disorders_Somatic_CentralSensitization_N <- rowSums(somatic[, c(
  "Fibromyalgia", "ChronicPain", "ChronicFatigue", "BackPain", 
  "PelvicPain", "Endometriosis", "IBS", "Cystitis"
)])
df$Disorders_Somatic_CentralSensitization <- df$Disorders_Somatic_CentralSensitization_N >= 1

# Cluster C: Autonomic Dysfunction (dysautonomia and cardiorespiratory instability)
df$Disorders_Somatic_AutonomicDysfunction_N <- rowSums(somatic[, c(
  "Hypermobility", "Arrhythmia", "ChestPain", "ShortBreath", 
  "Hypotension", "Hypertension", "SleepApnea", "COPD", "Bronchitis"
)])
df$Disorders_Somatic_AutonomicDysfunction <- df$Disorders_Somatic_AutonomicDysfunction_N >= 1

# Cluster D: Immune-Inflammatory (gut-brain axis, autoimmune and barrier-related conditions)
df$Disorders_Somatic_ImmuneInflammatory_N <- rowSums(somatic[, c(
  "SkinRashes", "Eczema", "Psoriasis", "Sjogrens", "Asthma", 
  "Celiac", "Gluten", "Lactose", "Crohn", "UlcerativeColitis", 
  "GERD", "MS"
)])
df$Disorders_Somatic_ImmuneInflammatory <- df$Disorders_Somatic_ImmuneInflammatory_N >= 1

p2 <- df |> 
  select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
  select(ends_with("_N"), Gender, Participant) |>
  pivot_longer(starts_with("Disorders_Somatic_")) |> 
  summarize(n = n(), .by=c("name", "value", "Gender")) |> 
  mutate(name = str_remove(name, "Disorders_Somatic_"),
         name = str_remove(name, "_N")) |> 
  ggplot(aes(x = value, y=n, fill=Gender)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Number of Issues") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~name)
p2
# 
# p2 <- df |> 
#   select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
#   select(-ends_with("_N"), Gender, Participant) |>
#   pivot_longer(starts_with("Disorders_Somatic_")) |> 
#   summarize(n = n(), .by=c("name", "value", "Gender")) |> 
#   filter(value == TRUE) |> 
#   ggplot(aes(x = name, y=n, fill=Gender)) +
#   geom_bar(stat="identity") +
#   scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
#   scale_y_continuous(expand = c(0, 0)) +
#   labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Type") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
# p2
```

### Wearables

```{r}
#| fig-height: 12

p1 <- df |> 
  ggplot(aes(x = Wearables_Number)) +
  geom_bar() +
  scale_x_continuous(breaks = c(0:max(df$Wearables_Number)), expand=c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Number of Wearable Devices", y = "Number of Participants", x="Number Owned") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal()

p2 <- df |> 
  select(contains("Wearables"), -contains("Importance"), -Wearables_Number) |> 
  pivot_longer(everything()) |> 
  summarize(n = n(), .by=c("name", "value")) |> 
  mutate(value = fct_relevel(value, "Not owning", "Never", "Very rarely", "A few times per week", "Once a day", "A few times per day", "A few times per hour"),
         name = str_remove(name, "Wearables_")) |> 
  ggplot(aes(x=value, y=n, fill=name)) +
  geom_bar(stat = "identity") +
  labs(x = "Usage frequency", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

get_databubble <- function(df, var="Heart") {
  x <- paste0("Wearables_", var)
  y <- paste0("Wearables_", var, "Importance")
  dat <- df[df[[x]] != "Not owning", ]
  dat <- summarize(dat, n = n(), .by = all_of(c(x, y)))
  lvl <- c("Never", "Very rarely", "A few times per week", "Once a day", "A few times per day", "A few times per hour")
  dat[[x]] <- factor(dat[[x]], levels=lvl[lvl %in% unique(dat[[x]])])
  names(dat) <- c("Usage", "Importance", "n")
  dat$Device <- var
  dat
}

p3 <- rbind(
  get_databubble(df, "Heart"),
  get_databubble(df, "Sleep"),
  get_databubble(df, "Steps"),
  get_databubble(df, "Weight"),
  get_databubble(df, "CalorieIntake"),
  get_databubble(df, "CaloriesBurnt")
) |> 
  filter(!is.na(Importance)) |> 
  ggplot(aes(x=Usage, y=Importance)) +
  geom_jitter(aes(size = n, color = Device), width = 0.2, height = 0.2, alpha = 0.5) +
  geom_smooth(aes(x=as.numeric(Usage), color = Device), method = "lm", se = FALSE, formula = 'y ~ x') +
  labs(x = "Usage frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

p1 / p2 / p3
```

:::

## Save

```{r}
df |> 
  select(-contains("AttentionCheck")) |> 
  write.csv("../data/data_participants.csv", row.names = FALSE)
```