### 2025-07-01 study2/analysis/2_analysis.qmd

---
title: "InteroceptionScale (study 2) - Data Analysis"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(EGAnet)
library(tidygraph)
library(ggraph)

set.seed(42)

Sys.setenv(CHROMOTE_CHROME = "C:\\Program Files (x86)\\Microsoft\\Edge\\Application\\msedge.exe")
```

```{r}
#| code-fold: true

df <- read.csv("../data/data_participants.csv") |> 
  rename(MINT_Urin_1 = MINT_Deficit_Urin_1,
         MINT_Urin_2 = MINT_Deficit_Urin_2,
         MINT_Urin_3 = MINT_Deficit_Urin_3,
         MINT_CaCo_1 = MINT_Deficit_CaCo_4,
         MINT_CaCo_2 = MINT_Deficit_CaCo_5,
         MINT_CaCo_3 = MINT_Deficit_CaCo_6,
         MINT_CaNo_1 = MINT_Deficit_CaNo_7,
         MINT_CaNo_2 = MINT_Deficit_CaNo_8,
         MINT_CaNo_3 = MINT_Deficit_CaNo_9,
         MINT_Olfa_1 = MINT_Deficit_Olfa_10,
         MINT_Olfa_2 = MINT_Deficit_Olfa_11,
         MINT_Olfa_3 = MINT_Deficit_Olfa_12,
         MINT_Sati_1 = MINT_Deficit_Sati_13,
         MINT_Sati_2 = MINT_Deficit_Sati_14,
         MINT_Sati_3 = MINT_Deficit_Sati_15,
         MINT_SexS_1 = MINT_Awareness_SexS_19,
         MINT_SexS_2 = MINT_Awareness_SexS_20,
         MINT_SexS_3 = MINT_Awareness_SexS_21,
         MINT_SexO_1 = MINT_Awareness_SexO_22,
         MINT_SexO_2 = MINT_Awareness_SexO_23,
         MINT_SexO_3 = MINT_Awareness_SexO_24,
         MINT_UrSe_1 = MINT_Awareness_UrSe_25,
         MINT_UrSe_2 = MINT_Awareness_UrSe_26,
         MINT_UrSe_3 = MINT_Awareness_UrSe_27,
         MINT_RelA_1 = MINT_Awareness_RelA_28,
         MINT_RelA_2 = MINT_Awareness_RelA_29,
         MINT_RelA_3 = MINT_Awareness_RelA_30,
         MINT_StaS_1 = MINT_Awareness_StaS_31,
         MINT_StaS_2 = MINT_Awareness_StaS_32,
         MINT_StaS_3 = MINT_Awareness_StaS_33,
         MINT_ExAc_1 = MINT_Awareness_ExAc_34,
         MINT_ExAc_2 = MINT_Awareness_ExAc_35,
         MINT_ExAc_3 = MINT_Awareness_ExAc_36,
         MINT_Card_1 = MINT_Sensitivity_Card_37,
         MINT_Card_2 = MINT_Sensitivity_Card_38,
         MINT_Card_3 = MINT_Sensitivity_Card_39,
         MINT_Resp_1 = MINT_Sensitivity_Resp_40,
         MINT_Resp_2 = MINT_Sensitivity_Resp_41,
         MINT_Resp_3 = MINT_Sensitivity_Resp_42,
         MINT_Gast_1 = MINT_Sensitivity_Gast_46,
         MINT_Gast_2 = MINT_Sensitivity_Gast_47,
         MINT_Gast_3 = MINT_Sensitivity_Gast_48,
         MINT_Derm_1 = MINT_Sensitivity_Derm_49,
         MINT_Derm_2 = MINT_Sensitivity_Derm_50,
         MINT_Derm_3 = MINT_Sensitivity_Derm_51)

colors <- c(Mint="#00BCD4", metaMint="#673AB7", miniMint="#26A36C", microMint="#8BC34A", 
            IAS="#F44336", MAIA="#FFAD04",  iMAIA="#FF7811", BPQ="#795548")
```

## Structure Validation

```{r}
#| code-fold: false

items <- select(df, starts_with("MINT_"))
names(items) <- str_remove(names(items), "MINT_")
```

```{r}
labels <- list(
  MINT_Urin_1 = "I sometimes feel like I need to urinate or defecate but when I go to the bathroom I produce less than I expected",
  MINT_Urin_2 = "I often feel the need to urinate even when my bladder is not full",
  MINT_Urin_3 = "Sometimes I am not sure whether I need to go to the toilet or not (to urinate or defecate)",
  MINT_CaCo_1 = "Sometimes my breathing becomes erratic or shallow and I often don't know why",
  MINT_CaCo_2 = "I often feel like I can't get enough oxygen by breathing normally",
  MINT_CaCo_3 = "Sometimes my heart starts racing and I often don't know why",
  MINT_CaNo_1 = "I often only notice how I am breathing when it becomes loud",
  MINT_CaNo_2 = "I only notice my heart when it is thumping in my chest",
  MINT_CaNo_3 = "I often only notice how I am breathing when my breathing becomes shallow or irregular",
  MINT_Olfa_1 = "I often check the smell of my armpits",
  MINT_Olfa_2 = "I often check the smell of my own breath",
  MINT_Olfa_3 = "I often check the smell of my farts",
  MINT_Sati_1 = "I don't always feel the need to eat until I am really hungry",
  MINT_Sati_2 = "Sometimes I don't realise I was hungry until I ate something",
  MINT_Sati_3 = "I don't always feel the need to drink until I am really thirsty",
  # MINT_SexA_16 = "I always feel in my body if I am sexually aroused",
  # MINT_SexA_17 = "I can always tell that I am sexually aroused from the way I feel inside",
  # MINT_SexA_18 = "I always know when I am sexually aroused",
  MINT_SexS_1 = "During sex or masturbation, I often feel very strong sensations coming from my genital areas",
  MINT_SexS_2 = "When I am sexually aroused, I often notice specific sensations in my genital area (e.g., tingling, warmth, wetness, stiffness, pulsations)",
  MINT_SexS_3 = "My genital organs are very sensitive to pleasant stimulations",
  MINT_SexO_1 = "In general, I am very sensitive to changes in my genital organs",
  MINT_SexO_2 = "I can notice even very subtle changes in the state of my genital organs",
  MINT_SexO_3 = "I am always very aware of the state of my genital organs, even when I am calm",
  MINT_UrSe_1 = "In general, I am very aware of the sensations that are happening when I am urinating",
  MINT_UrSe_2 = "In general, I am very aware of the sensations that are happening when I am defecating",
  MINT_UrSe_3 = "I often experience a pleasant sensation when relieving myself when urinating or defecating)",
  MINT_RelA_1 = "I always know when I am relaxed",
  MINT_RelA_2 = "I always feel in my body if I am relaxed",
  MINT_RelA_3 = "My body is always in the same specific state when I am relaxed",
  MINT_StaS_1 = "Being relaxed is a very different bodily feeling compared to other states (e.g., feeling anxious, sexually aroused or after exercise)",
  MINT_StaS_2 = "Being sexually aroused is a very different bodily feeling compared to other states (e.g., feeling anxious, relaxed, or after physical exercise)",
  MINT_StaS_3 = "Being anxious is a very different bodily feeling compared to other states (e.g., feeling sexually aroused, relaxed or after exercise)",
  MINT_ExAc_1 = "I can always accurately feel when I am about to burp",
  MINT_ExAc_2 = "I can always accurately feel when I am about to fart",
  MINT_ExAc_3 = "I can always accurately feel when I am about to sneeze",
  MINT_Card_1 = "In general, I am very sensitive to changes in my heart rate",
  MINT_Card_2 = "I can notice even very subtle changes in the way my heart beats",
  MINT_Card_3 = "I often notice changes in my heart rate",
  MINT_Resp_1 = "I can notice even very subtle changes in my breathing",
  MINT_Resp_2 = "I am always very aware of how I am breathing, even when I am calm",
  MINT_Resp_3 = "In general, I am very sensitive to changes in my breathing",
  # MINT_Sign_43 = "When something important is happening in my life, I can feel immediately feel changes in my heart rate",
  # MINT_Sign_44 = "When something important is happening in my life, I can immediately feel changes in my breathing",
  # MINT_Sign_45 = "When something important is happening in my life, I can feel it in my body",
  MINT_Gast_1 = "I can notice even very subtle changes in what my stomach is doing",
  MINT_Gast_2 = "In general, I am very sensitive to what my stomach is doing",
  MINT_Gast_3 = "I am always very aware of what my stomach is doing, even when I am calm",
  MINT_Derm_1 = "In general, my skin is very sensitive",
  MINT_Derm_2 = "My skin is susceptible to itchy fabrics and materials",
  MINT_Derm_3 = "I can notice even very subtle stimulations to my skin (e.g., very light touches)"
  # MINT_SexC_52 = "When I am sexually aroused, I often feel changes in the way my heart beats (e.g., faster or stronger)",
  # MINT_SexC_53 = "When I am sexually aroused, I often feel changes in my breathing (e.g., faster, shallower, or less regular)",
  # MINT_SexC_54 = "When I am sexually aroused, I often feel changes in my temperature (e.g., feeling warm or cold)"
)

cleanlabels <- function(x, qname=FALSE) {
  is_factor <- FALSE
  if(is.factor(x)) {
    is_factor <- TRUE
    vec <- x
    x <- levels(x)
  }

  x <- x |> 
    str_replace("MINT_", ifelse(qname, "MINT - ", "")) |> 
    str_replace("MAIA_", ifelse(qname, "MAIA - ", "")) |> 
    str_replace("_", " - ")

  if(is_factor) {
    levels(vec) <- x
    return(vec)
  }
  x
}
```

### Full

#### EGA

```{r}
#| eval: false

ega1 <- bootEGA(
  items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  allow.singleton = TRUE,
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3, ncores = 4)

save(ega1, file="models/ega1.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega1.RData")

itemstability <- EGAnet::itemStability(ega1, IS.plot=FALSE)
plot(itemstability)
```

#### Hierarchical Clustering

```{r}
#| fig-width: 12
#| fig-height: 14
#| warning: false

hclust <- pvclust::pvclust(items,
                 method.hclust = "ward.D2",
                 method.dist = "correlation",
                 nboot = 1000, quiet=TRUE, parallel=TRUE)
plot(hclust, hang = -1, cex = 0.5)
pvclust::pvrect(hclust, alpha=0.95, max.only=TRUE)
```

### Item Reduction

::: {.panel-tabset}

```{r}
features <- data.frame(
  SexS = rowMeans(select(df, starts_with("MINT_SexS_")), na.rm = TRUE),
  StaS = rowMeans(select(df, starts_with("MINT_StaS_")), na.rm = TRUE),
  SexO = rowMeans(select(df, starts_with("MINT_SexO_")), na.rm = TRUE),
  ExaC = rowMeans(select(df, starts_with("MINT_ExAc_")), na.rm = TRUE),
  RelA = rowMeans(select(df, starts_with("MINT_RelA_")), na.rm = TRUE),
  UrSe = rowMeans(select(df, starts_with("MINT_UrSe_")), na.rm = TRUE),
  CaCo = rowMeans(select(df, starts_with("MINT_CaCo_")), na.rm = TRUE),
  Urin = rowMeans(select(df, starts_with("MINT_Urin_")), na.rm = TRUE),
  Sati = rowMeans(select(df, starts_with("MINT_Sati_")), na.rm = TRUE),
  Derm = rowMeans(select(df, starts_with("MINT_Derm_")), na.rm = TRUE),
  CaNo = rowMeans(select(df, starts_with("MINT_CaNo_")), na.rm = TRUE),
  Olfa = rowMeans(select(df, starts_with("MINT_Olfa_")), na.rm = TRUE),
  Resp = rowMeans(select(df, starts_with("MINT_Resp_")), na.rm = TRUE),
  Card = rowMeans(select(df, starts_with("MINT_Card_")), na.rm = TRUE),
  Gast = rowMeans(select(df, starts_with("MINT_Gast_")), na.rm = TRUE)
)

features$CardGast = rowMeans(select(features, Card, Gast), na.rm = TRUE)
features$CardResp = rowMeans(select(features, Card, Resp), na.rm = TRUE)
features$CardRespGast = rowMeans(select(features, Card, Resp, Gast), na.rm = TRUE)

rez <- data.frame()
for(outcome_group in c("Interoception", "Alexithymia", "Mood", "Mental", "Physical", "Biofeedback", "Dissociative", "ER", "Primals", "Lifestyle")) {
  if(outcome_group == "Interoception") {
    outcomes <- c("IAS", "BPQ", "MAIA_EmotionalAwareness", 
                  "MAIA_BodyListening", "MAIA_SelfRegulation",
                  "MAIA_AttentionRegulation", "MAIA_Trusting")
  } else if(outcome_group == "Alexithymia") {
    outcomes <- c("TAS_EOT", "TAS_DDF", "TAS_DIF")
  } else if(outcome_group == "Mood") {
    outcomes <- c("LifeSatisfaction", "PHQ4_Anxiety", "PHQ4_Depression", "ERS_Arousal", "ERS_Sensitivity", "ERS_Persistence")
  } else if (outcome_group == "Mental") {
    outcomes <- c("Disorders_Psychiatric_MoodTreatment", "Disorders_Psychiatric_AnxietyTreatment",
                  "Disorders_Psychiatric_Eating", "Disorders_Psychiatric_Addiction",
                  "Disorders_Psychiatric_Borderline", "Disorders_Psychiatric_Autism",
                  "Disorders_Psychiatric_ADHD")
  } else if (outcome_group == "Physical") {
    outcomes <- c("Disorders_Somatic_AfferentSensitivity", "Disorders_Somatic_CentralSensitization_N",
                  "Disorders_Somatic_AutonomicDysfunction", "Disorders_Somatic_ImmuneInflammatory")
  } else if (outcome_group == "Biofeedback") {
    outcomes <- c("Wearables_HeartImportance", "Wearables_StepsImportance", 
                  "Wearables_SleepImportance", "Wearables_CalorieIntakeImportance",
                  "Wearables_WeightImportance")
  } else if (outcome_group == "Dissociative") {
    outcomes <- names(select(df, starts_with("CEFSA_")))
  } else if (outcome_group == "ER") {
    outcomes <- names(select(df, starts_with("CERQ_"))) 
  } else if (outcome_group == "Primals") {
    outcomes <- names(select(df, starts_with("PI_")))
  } else if (outcome_group == "Lifestyle") {
    outcomes <- c("Physical_Active", "BMI")
  }

  out_group <- data.frame()
  for(outcome in outcomes) {

    out <- data.frame()
    for(feature in names(features)) {

      dat <- data.frame(Participant = df$Participant)
      dat[[outcome]] <- df[[outcome]]
      dat[[feature]] <- features[[feature]]
      dat <- dat[!is.na(dat[[outcome]]) & !is.na(dat[[feature]]), ]

      f <- as.formula(paste0(outcome, " ~ ", feature))
      if(length(unique(dat[[outcome]])) > 2) {
        m <- lm(f, data = dat)
      } else {
        m <- glm(f, data = dat, family = "binomial")
      }
      out <- rbind(out, data.frame(
        Group = outcome_group,
        Outcome = outcome,
        Dimension = feature,
        R2 = as.numeric(r2(m))[1]
      ))
    }

    out$R2 <- as.numeric(normalize(out$R2))
    out$Rank <- datawizard::ranktransform(1 - out$R2)
    out_group <- rbind(out_group, out)
  }

  ave <- summarize(out_group, R2 = median(R2), Rank = min(Rank), .by="Dimension") |> 
    mutate(Group = outcome_group, Outcome = outcome_group)

  rez <- rbind(rez, out_group, ave)
}

grandave <- rez |> 
  filter(Group != Outcome) |> 
  summarize(R2 = mean(R2), Rank = min(Rank), .by="Dimension") |> 
  mutate(Group = "Average", Outcome = "Average")
```

#### R2

```{r}
rbind(rez, grandave) |> 
  select(-Group, -Rank) |> 
  pivot_wider(names_from = "Outcome", values_from = "R2") |> 
  mutate(Group = case_when(
    Dimension %in% c("Card", "Resp", "Gast", "CardRespGast", "CardResp", "CardGast") ~ "Visceral", 
    Dimension %in% c("RelA", "SexS", "ExaC", "StaS", "SexO", "UrSe") ~ "sensitivity", 
    Dimension %in% c("CaCo", "Urin", "Derm", "Olfa", "Sati", "CaNo") ~ "Deficit", 
    .default = "Other")) |> 
  select(Group, Dimension, Average, Interoception, Biofeedback, ER, 
         Primals, Alexithymia, Dissociative, Mood, 
         Physical, Mental, Lifestyle) |>
  arrange(Group, desc(Average)) |> 
  # select(-Group) |> 
  gt::gt() |> 
  gt::data_color(
    columns = -c(Dimension, Group),
    palette = c("white", "green"),
    domain = c(0, 1)
  ) |>
  gt::fmt_number() |>
  gt::tab_header(
    title = "Variable Importance",
    subtitle = "Normalized Predictive Power (R2)"
  ) |> 
  gt::tab_row_group(
    label = "Sensitivity",
    rows = Group == "sensitivity"
  )  |> 
  gt::tab_row_group(
    label = "Interoception",
    rows = Group == "Visceral"
  ) |> 
  gt::tab_row_group(
    label = "Deficit",
    rows = Group == "Deficit"
  ) |> 
  gt::cols_hide(Group)
```

#### Rank

```{r}
rbind(rez, grandave) |> 
  select(-Group, -R2) |> 
  pivot_wider(names_from = "Outcome", values_from = "Rank") |> 
  mutate(Group = case_when(
    Dimension %in% c("Card", "Resp", "Gast", "CardRespGast", "CardResp", "CardGast") ~ "Visceral", 
    Dimension %in% c("RelA", "SexS", "ExaC", "StaS", "SexO", "UrSe") ~ "sensitivity", 
    Dimension %in% c("CaCo", "Urin", "Derm", "Olfa", "Sati", "CaNo") ~ "Deficit", 
    .default = "Other")) |> 
  datawizard::data_relocate(select=c("Group", "Dimension", "Average", "Interoception")) |>
  arrange(Group, Average) |> 
  gt::gt() |> 
  # gt::fmt_number(columns = "Average") |>
  gt::data_color(
    columns = -c(Dimension, Group),
    palette = c("green", "white"),
    domain = c(1, 18)
  ) |> 
  gt::opt_interactive(page_size_default=20)
```

#### Clusters

```{r}
hclust2 <- rez |>
  filter(Group != Outcome) |>
  select(-Group, -Rank) |>
  pivot_wider(names_from = "Dimension", values_from = "R2") |>
  select(-Outcome) |>
  pvclust::pvclust(
     method.hclust = "average",
     method.dist = "correlation",
     nboot = 1000, quiet=TRUE, parallel=TRUE)

plot(hclust2, hang = -1, cex = 0.5)
pvclust::pvrect(hclust2, alpha=0.95, max.only=TRUE)
```

:::

### Final

```{r}
#| fig-width: 10
#| fig-height: 8
#| fig-keep: "first"
#| code-fold: false

items <- items |> 
  select(
    -contains("UrSe"), # Unstable item
    -contains("StaS"), # Too vague and overlapping
    -contains("SexO"), # Overalapping with SexS

    -contains("CaNo") # Overlap with CaCo
    # -contains("Olfa"), # Unstable metacluster
    # -contains("Sati")
  )
```

::: {.panel-tabset}

#### EGA

```{r}
#| eval: false

ega2 <- bootEGA(
  items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  # consensus.method = "lowest_tefi",
  allow.singleton = TRUE,
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3, ncores = 4)

save(ega2, file="models/ega2.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega2.RData")

# EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega2, IS.plot=FALSE)
p_stab <- plot(itemstability)
p_stab
```

#### Item Table

```{r}
#| message: false

make_loadingtable <- function(x) {
  t <- as.data.frame(x) |>
    datawizard::data_addprefix("C")
  t$Cluster <- colnames(t)[max.col(t, ties.method='first')]

  t |>
    rownames_to_column(var="Item") |>
    rowwise() |>
    mutate(Max = max(c_across(-c(Item, Cluster)))) |>
    arrange(Cluster, desc(Max)) |>
    as.data.frame()
}

dimensions <- list(
  C01 = "SexS",       # Sexual Arousal Sensitivity
  C02 = "Derm",       # Dermal Hypersensitivity
  C03 = "Card",       # Cardioception
  C04 = "Sati",       # Satiety Awareness
  C05 = "Gast",       # Gastroception
  C06 = "Olfa",       # Olfactory Compensation
  C07 = "Resp",       # Respiroception
  C08 = "RelA",       # Relaxation Awareness
  C09 = "Urin",       # Urointestinal Inaccuracy
  C10 = "ExAc",       # Expulsion Accuracy
  C11 = "CaCo"       # Cardiorespiratory Confusion
  # C01 = "C01",
  # C02 = "C02",
  # C03 = "C03",
  # C04 = "C04",
  # C05 = "C05",
  # C06 = "C06",
  # C07 = "C07",
  # C08 = "C08",
  # C09 = "C09",
  # C10 = "C10",
  # C11 = "C11",
  # C12 = "C12"
)

higher <- make_loadingtable(t(net.loads(ega2$EGA$higher_order, loading.method="revised")$std)) |> 
  arrange(Item)
cluster_order <- names(higher)[!names(higher) %in% c("Item", "Cluster", "Max")]

lower <- make_loadingtable(net.loads(ega2$EGA$lower_order, loading.method="revised")$std) |>
  select(names(higher)) |>
  mutate(Cluster = fct_relevel(Cluster, cluster_order)) |>
  arrange(Cluster, desc(Max))
dimensions <- dimensions[names(dimensions) %in% names(lower)]

higher$Item <- paste0("M", higher$Item)
# higher$Label <- c("Interoceptive Deficit", "Interoceptive Awareness", "Interoceptive Sensitivity")
higher$Label <- c("Interoceptive Awareness", "Interoceptive Deficit", "Visceroception")
higher$ID  <- ""

lower <- mutate(lower, Label = labels[paste0("MINT_", Item)], ID=Item, Item = 1:n())

tab1 <- rbind(higher, lower) |>
  select(-ID) |>
  datawizard::data_relocate("Label", after = "Item") |>
  datawizard::data_rename(names(dimensions), dimensions[names(dimensions)]) |>
  select(-Cluster, -Max) |> 
  mutate(`|` = "") |> 
  gt::gt() |>
  gt::tab_row_group(
    label = "Items",
    rows = 1:(nrow(lower) + nrow(higher))
  ) |>
  gt::tab_row_group(
    label = "Metaclusters",
    rows = 1:nrow(higher)
  ) |>
  gt::data_color(columns=-Item,
                 method = "numeric",
                 palette = c("red", "white", "green"),
                 domain = c(-1, 1)) |>
  gt::tab_style(
    style = gt::cell_text(size="small", style="italic"),
    locations = gt::cells_body(columns="Label", rows=c(4:36))
  ) |>
  gt::tab_style(
    style = list(gt::cell_text(weight="bold"),
                 gt::cell_fill(color = "#F5F5F5")),
    locations = list(
      gt::cells_row_groups(groups = "Items"),
      gt::cells_row_groups(groups = "Metaclusters")
    )
  ) |>
  gt::fmt_number(columns=-Item, decimals=2) |>
  gt::tab_header(
    title = gt::md("**Item Loadings**"),
    subtitle = "Node centrality"
  ) |> 
  # Metacluster 1
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Item", rows=c(1))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_column_labels(columns = 3),
      gt::cells_body(columns="Item", rows=c(4:6))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#AED581"),
    locations = list(
      gt::cells_column_labels(columns = 4),
      gt::cells_body(columns="Item", rows=c(7:9))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#DCE775"),
    locations = list(
      gt::cells_column_labels(columns = 5),
      gt::cells_body(columns="Item", rows=c(10:12))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#BA68C8"),
    locations = list(
      gt::cells_body(columns="Item", rows=c(2))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#BA68C8"),
    locations = list(
      gt::cells_column_labels(columns = 6),
      gt::cells_body(columns="Item", rows=c(13:15))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#9575CD"),
    locations = list(
      gt::cells_column_labels(columns = 7),
      gt::cells_body(columns="Item", rows=c(16:18))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#7986CB"),
    locations = list(
      gt::cells_column_labels(columns = 8),
      gt::cells_body(columns="Item", rows=c(19:21))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#64B5F6"),
    locations = list(
      gt::cells_column_labels(columns = 9),
      gt::cells_body(columns="Item", rows=c(22:24))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#4DD0E1"),
    locations = list(
      gt::cells_column_labels(columns = 10),
      gt::cells_body(columns="Item", rows=c(25:27))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#FF7811"),
    locations = list(
      gt::cells_body(columns="Item", rows=c(3))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_column_labels(columns = 11),
      gt::cells_body(columns="Item", rows=c(28:30))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#E57373"),
    locations = list(
      gt::cells_column_labels(columns = 12),
      gt::cells_body(columns="Item", rows=c(31:33))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#A1887F"),
    locations = list(
      gt::cells_column_labels(columns = 13),
      gt::cells_body(columns="Item", rows=c(34:36))
    )
  ) 
tab1
```

```{r}
#| eva: false

gt::gtsave(tab1, "figures/table1.png", vwidth=1660, expand=0, quiet=TRUE)
```

#### Hierarchical Clustering

```{r}
#| fig-width: 12
#| fig-height: 14
#| warning: false

rez <- pvclust::pvclust(items,
                 method.hclust = "ward.D2",
                 method.dist = "correlation",
                 nboot = 1000, quiet=TRUE, parallel=TRUE)
plot(rez, hang = -1, cex = 0.5)
pvclust::pvrect(rez, alpha=0.95, max.only=FALSE)
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| warning: false

dendrogram <- as.dist(1 - cor(items, use = "pairwise.complete.obs")) |> 
  hclust(method = "ward.D2") |> 
  tidygraph::as_tbl_graph()

# Process Nodes
nodes <- as.list(dendrogram)$nodes |> 
  mutate(Group = case_when(str_detect(label, "Card") ~ "Card", 
                           str_detect(label, "Resp") ~ "Resp",
                           str_detect(label, "Gast") ~ "Gast",
                           str_detect(label, "SexS") ~ "SexS",
                           str_detect(label, "RelA") ~ "RelA",
                           str_detect(label, "ExAc") ~ "ExAc",
                           str_detect(label, "Urin") ~ "Urin",
                           str_detect(label, "CaCo") ~ "CaCo",
                           str_detect(label, "Derm") ~ "Derm",
                           str_detect(label, "Olfa") ~ "Olfa",
                           str_detect(label, "Sati") ~ "Sati",
                           label == "" ~ "NA",
                           .default = "Other"),
         Size = setNames(lower$Max, lower$ID)[label],
         item = label)
nodes$idx <- 1:nrow(nodes)

# Rename accoring to table
items_newnames <- lower |>
  separate(ID, into = c("Group", "Item"), sep = "_", remove = FALSE) |>
  mutate(Name = paste0(Group, "-", 1:nrow(lower)))
items_newnames <- setNames(items_newnames$Name, items_newnames$ID)

nodes$label <- items_newnames[nodes$label]
nodes$label <- ifelse(is.na(nodes$label), "", nodes$label)

# Central node 
max_size  <- max(nodes$Size, na.rm = TRUE)
nodes[nodes$height == max(nodes$height), c("Group", "Size")] <- data.frame(Group="Central", Size=2 * max_size)
# Metaclusters
nodes[17, c("Group", "Size")] <- data.frame(Group="Awareness", Size=1.5 * max_size)
nodes[34, c("Group", "Size")] <- data.frame(Group="Visceroception", Size=1.5 * max_size)
nodes[63, c("Group", "Size")] <- data.frame(Group="Deficit", Size=1.5 * max_size)

# Process Edges
edges <- as.list(dendrogram)$edges
edges$linewidth = datawizard::rescale(nodes[edges$from, ]$height, to = c(0.1, 1))

p_hclust <- tbl_graph(nodes = nodes, edges = edges) |> 
  ggraph(layout = "dendrogram", circular = TRUE) +
  # geom_edge_diagonal(strength = 0.7, linewidth = 1) +
  geom_edge_elbow2(aes(edge_width=linewidth), color="darkgrey") +
  geom_node_point(aes(filter=Group %in% c("Central"), size = Size), color="darkgrey") +
  geom_node_point(aes(filter=Group %in% c("Visceroception", "Awareness", "Deficit"), color=Group, size = Size)) +
  geom_node_text(aes(
    label = ifelse(label != "NA", label, NA),
    x = x * 1.05,
    y = y * 1.05,
    filter = label != "",
    angle = ifelse(
      x >= 0,
      asin(y) * 360 / 2 / pi,
      360 - asin(y) * 360 / 2 / pi
    ),
    hjust = ifelse(
      x >= 0, 0, 1
    ))) +
  geom_node_point(aes(filter = label != "", color=Group, size=Size), alpha = 1) +
  # geom_node_text(aes(label=idx)) +  # Debug
  scale_edge_width_continuous(range=c(1, 3), guide = "none") +
  scale_size_continuous(range=c(3, 11), guide = "none") +
  scale_color_manual(values = c(
    "Visceroception" = "#FF7811", "Card" = "#F44336", "Resp"="#FF9800", "Gast"="#795548",
    "Awareness" = "#4CAF50", "ExAc"="#4CAF50", "RelA"="#8BC34A", "SexS" = "#CDDC39",
    "Deficit" = "#9C27B0", "CaCo"="#9C27B0", "Urin" = "#673AB7", "Derm"="#3F51B5",
    "Sati" = "#2196F3", "Olfa" = "#00BCD4"),
                     breaks = c(
                       "Awareness", "ExAc", "RelA", "SexS",
                       "Deficit", "CaCo", "Urin", "Derm", "Sati", "Olfa",
                       "Visceroception", "Resp", "Card", "Gast"
                       ),
                     labels=c(
                       "***Awareness***", "ExAc", "RelA", "SexS",
                       "***Deficit***", "CaCo", "Urin", "Derm", "Sati", "Olfa",
                       "***Visceroception***", "Resp", "Card", "Gast"
                       )) +
  ggtitle("Hierarchical Clustering", subtitle = "Method = Correlation") +
  coord_equal(clip = "off", xlim = c(-1.25, 1.25), ylim = c(-1.25, 1.25)) +
  theme_void() + 
  guides(color = guide_legend(override.aes = list(size = c(7.5, 5, 5, 3.5, 7.5, 5, 5, 3.5, 3.5, 3.5, 7.5, 5, 5, 3.5)))) +
  theme(legend.text = ggtext::element_markdown(),
        legend.title = element_blank(),
        plot.title = element_blank(),  #element_text(face="bold")
        plot.subtitle = element_blank())
p_hclust
```

```{r}
#| eval: FALSE

ggsave("figures/fig1a.png", p_stab, width=10, height=10*0.8, dpi=300)
ggsave("figures/fig1b.png", p_hclust, width=12, height=8, dpi=300)
```

#### EFA

```{r}
#| fig-width: 12
#| fig-height: 14

n <- parameters::n_factors(items)
plot(n)

fa <- parameters::factor_analysis(items, n = 9, rotation = "varimax", sort = TRUE)
plot(fa)

# fa |>
#   select(Variable, Uniqueness) |>
#   mutate(Group = str_extract(Variable, "^[A-Z][a-z]+")) |>
#   mutate(Average = mean(Uniqueness), .by = "Group") |> 
#   arrange(desc(Average), desc(Uniqueness)) |> 
#   select(-Average, -Group) |> 
#   gt::gt() |> 
#   gt::data_color(
#     columns = "Uniqueness",
#     palette = c("red", "white", "green"),
#     domain = c(0, 1) 
#   ) 
```

:::

### Score Extraction

```{r}
# Awareness
df$MINT_SexS <- rowMeans(select(df, starts_with("MINT_SexS_")), na.rm = TRUE)
df$MINT_ExaC <- rowMeans(select(df, starts_with("MINT_ExaC_")), na.rm = TRUE)
df$MINT_RelA <- rowMeans(select(df, starts_with("MINT_RelA_")), na.rm = TRUE)
# df$MINT_StaS <- rowMeans(select(df, starts_with("MINT_StaS_")), na.rm = TRUE)
# df$MINT_SexO <- rowMeans(select(df, starts_with("MINT_SexO_")), na.rm = TRUE)
# df$MINT_UrSe <- rowMeans(select(df, starts_with("MINT_UrSe_")), na.rm = TRUE)

# Deficit
df$MINT_Urin <- rowMeans(select(df, starts_with("MINT_Urin_")), na.rm = TRUE)
df$MINT_CaCo <- rowMeans(select(df, starts_with("MINT_CaCo_")), na.rm = TRUE)
df$MINT_Derm <- rowMeans(select(df, starts_with("MINT_Derm_")), na.rm = TRUE)
# df$MINT_CaNo <- rowMeans(select(df, starts_with("MINT_CaNo_")), na.rm = TRUE)

df$MINT_Sati <- rowMeans(select(df, starts_with("MINT_Sati_")), na.rm = TRUE)
df$MINT_Olfa <- rowMeans(select(df, starts_with("MINT_Olfa_")), na.rm = TRUE)

# Visceral
df$MINT_Resp <- rowMeans(select(df, starts_with("MINT_Resp_")), na.rm = TRUE)
df$MINT_Card <- rowMeans(select(df, starts_with("MINT_Card_")), na.rm = TRUE)
df$MINT_Gast <- rowMeans(select(df, starts_with("MINT_Gast_")), na.rm = TRUE)

df <- df[names(df)[!grepl("MINT_.*[0-9]$", names(df))]]
```

### Mini Versions

```{r}
mint <- select(df, starts_with("MINT_"))

metamint <- data.frame(
  MINT_Awareness = rowMeans(select(df, MINT_SexS, MINT_ExaC, MINT_RelA)),
  MINT_Deficit = rowMeans(select(df, MINT_CaCo, MINT_Urin, MINT_Derm, MINT_Sati, MINT_Olfa)),
  MINT_Visceroception = rowMeans(select(df, MINT_Card, MINT_Resp, MINT_Gast))
)

minimint <- data.frame(
  MINT_Awareness = rowMeans(select(df, MINT_ExaC, MINT_RelA)),
  MINT_Deficit = rowMeans(select(df, MINT_CaCo, MINT_Urin)),
  MINT_Visceroception = rowMeans(select(df, MINT_Resp, MINT_Card))
)

micromint <- data.frame(
  MINT_ExaC = df$MINT_ExaC,
  MINT_CaCo = df$MINT_CaCo,
  MINT_Resp = df$MINT_Resp
)
```

```{r}
#| warning: false

make_cor <- function(df1, df2=NULL, qname_x=TRUE, qname_y=TRUE, textsize = 2.9, xcol = "#424242", ycol="#424242") {

  r <- df1 |> 
    correlation::correlation(data2=df2, p_adjust="none", redundant=ifelse(is.null(df2), TRUE, FALSE)) |> 
    correlation::cor_sort() |>
    mutate(label = ifelse(p < .001, paste0(insight::format_value(r)), ""),
           Parameter1 = cleanlabels(Parameter1, qname=qname_x),
           Parameter2 = cleanlabels(Parameter2, qname=qname_y))
  r[as.character(r$Parameter1) == as.character(r$Parameter2), "label"] <- ""

  xnames <- levels(r$Parameter1)
  ynames <- levels(r$Parameter2)

  xcol <- case_when(
    str_detect(xnames, "MINT") ~ "#00838F",
    str_detect(xnames, "MAIA") ~ "#EF6C00",
    str_detect(xnames, "IAS") ~ "#C62828",
    str_detect(xnames, "BPQ") ~ "#795548",
    .default = xcol
  )
  ycol <- case_when(
    str_detect(ynames, "MINT") ~ "#00838F",
    str_detect(ynames, "MAIA") ~ "#EF6C00",
    str_detect(ynames, "IAS") ~ "#C62828",
    str_detect(ynames, "BPQ") ~ "#795548",
    .default = ycol
  )
  xbold <- case_when(
    xnames %in% c("Awareness", "MINT - Awareness") ~ "bold",
    str_detect(xnames, "Visceroception") ~ "bold",
    str_detect(xnames, "Deficit") ~ "bold",
    .default = "plain"
  )
  ybold <- case_when(
    ynames %in% c("Awareness", "MINT - Awareness") ~ "bold",
    str_detect(ynames, "Visceroception") ~ "bold",
    str_detect(ynames, "Deficit") ~ "bold",
    .default = "plain"
  )

  r |> 
    ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
    geom_tile() +
    geom_text(aes(label=label), size=textsize) +
    scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust=1, color = xcol, face = xbold),
          axis.text.y = element_text(color = ycol, face = ybold),
          legend.position = "none",
          axis.title = element_blank())
}

p_cormint <- make_cor(cbind(mint, metamint), qname_x = FALSE, qname_y = FALSE, textsize = 2.5,
                      xcol = "#00838F", ycol = "#00838F") +
  labs(title="Correlations between MINT dimensions") +
  theme(axis.title = element_blank())

p_cormint
```

## Convegent Validity

::: {.panel-tabset}

### Interoception

```{r}
#| warning: false

p_corintero <- make_cor(select(df, starts_with("MAIA"), IAS, BPQ),
                        cbind(mint, metamint), qname_y = FALSE, textsize = 2.5,
                        ycol = "#00838F") +
  labs(title="MINT vs. Other Interoception Questionnaires") +
  theme(axis.title = element_blank())
p_corintero
```

Most correlated dimensions:
- IAS: ExAc
- Trusting: RelA
- Noticing: Resp
- Emotional Awareness: Resp
- Body Listening: Gast
- SelfRegulation: RelA
- AttentionRegulation: RelA
- BPQ: Card

### Alexithymia

```{r}
table_models <- function(models) {
  test <- test_performance(models)
  table <- compare_performance(models, metrics=c("R2", "BIC"))
  test$Model <- NULL
  test$Omega2 <- NULL
  test$p_Omega2 <- NULL
  test$R2 <- table$R2
  test$BIC <- table$BIC

  preds <- c()
  coefs <- c()
  for(m in models){
    x <- names(sort(abs(coef(m)[-1]), decreasing=TRUE))
    preds <- c(preds, paste0(cleanlabels(x), collapse = ", "))
    if(length(coef(m)) == 2) {
      coefs <- c(coefs, coef(m)[2])
    }
  }

  if(length(coefs) == length(models)) test$Coefficient <- coefs
  format(test, zap_small = TRUE)
}

# Individual predictors ------------------------------------------------------

compare_predictors <- function(outcome="TAS_DIF", family = "linear") {

  preds <- c("BPQ", "IAS", 
             names(select(df, starts_with("MINT"), -matches("[[:digit:]]"))),
             names(select(df, starts_with("MAIA"))),
             names(metamint))

  models <- list()
  for(p in preds) {
    if(length(unique(df[[outcome]])) > 2) {
      if (family == "linear") {
        models[[p]] <- lm(as.formula(paste0(outcome, " ~ ", p)), data=cbind(df, metamint))
      } else {
        models[[p]] <- glm(as.formula(paste0(outcome, " ~ ", p)), data=cbind(df, metamint), family=family)
      }
    } else {
      models[[p]] <- glm(as.formula(paste0(outcome, " ~ ", p)), data=standardize(cbind(df, metamint), exclude=outcome), family="binomial")
    }
  }
  perf <- compare_performance(models) |> 
    arrange(AIC)

  best_mint <- perf$Name[str_detect(perf$Name, "MINT_")][1]
  best_maia <- perf$Name[str_detect(perf$Name, "MAIA")][1]

  table <- table_models(models[c(best_mint, best_maia, "BPQ", "IAS")])

  # Make plots --------------------------------------------------
  df_pred <- data.frame()
  for(p in c(best_mint, best_maia, "BPQ", "IAS")) {
    pred <- modelbased::estimate_relation(models[[p]], by=p, length=30) 
    names(pred)[1] <- "Value"
    pred$Predictor <- p
    df_pred <- rbind(df_pred, pred)
  }

  cols <- colors
  cols[[best_mint]] <- cols["Mint"]
  cols[[best_maia]] <- cols["MAIA"]

  p <- df_pred |>
    mutate(Predictor = fct_rev(Predictor)) |> 
    ggplot(aes(x=Value, y=Predicted)) +
    geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=Predictor), alpha=0.3) +
    geom_line(aes(color=Predictor)) +
    scale_fill_manual(values=cols) +
    scale_color_manual(values=cols) +
    facet_wrap(~Predictor, scales="free_x") +
    labs(y=outcome, x="Interoception Index")

  list(table=table, p=p, outcome=outcome)
}

# Full Models ------------------------------------------------------

compare_models_lm <- function(outcome="TAS_DIF") {

  # Full Questionnaires
  m_mint <- lm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(mint, df[outcome]))
  m_metamint <- lm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(metamint, df[outcome])) 
  m_minimint <- lm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(minimint, df[outcome])) 
  m_micromint <- lm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(micromint, df[outcome])) 
  m_ias <- lm(as.formula(paste0(outcome, " ~ IAS")),
              data = df)
  m_maia <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), starts_with("MAIA"))) 
  m_imaia <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), 
                              starts_with("MAIA"), 
                              -contains("NotWorrying"),
                              -contains("NotDistracting"))) 
  m_imaia <- lm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome),
                              # MAIA_AttentionRegulation, # remove?
                              # MAIA_SelfRegulation, # remove?
                              # MAIA_Trusting, # remove?
                              MAIA_BodyListening, 
                              MAIA_EmotionalAwareness,
                              MAIA_Noticing))
  m_bpq <- lm(as.formula(paste0(outcome, " ~ BPQ")),
                data = df)

  models_full <- list(m_mint=m_mint, m_metamint=m_metamint,
                      m_minimint=m_minimint, m_micromint=m_micromint, 
                      m_maia=m_maia, m_imaia=m_imaia, 
                      m_ias=m_ias, m_bpq=m_bpq)
  table_models(models_full)
}
```

```{r}
preds_dif <- compare_predictors("TAS_DIF")
display(preds_dif$table, title="DIF")
preds_dif$p
mods_dif <- compare_models_lm("TAS_DIF")
display(mods_dif, title="DIF")

preds_ddf <- compare_predictors("TAS_DDF")
display(preds_ddf$table, title="DDF")
preds_ddf$p
mods_ddf <- compare_models_lm("TAS_DDF")
display(mods_ddf, title="DDF")

preds_eot <- compare_predictors("TAS_EOT")
display(preds_eot$table, title="EOT")
preds_eot$p
mods_eot <- compare_models_lm("TAS_EOT")
display(mods_eot, title="EOT")
```

```{r}
make_cor(select(df, starts_with("TAS_")),
         cbind(mint, metamint, select(df, starts_with("MAIA"), IAS, BPQ)))
```

### Mood

```{r}
preds_ls <- compare_predictors("LifeSatisfaction")
display(preds_ls$table, title="Life Satisfaction")
preds_ls$p
mods_ls <- compare_models_lm("LifeSatisfaction")
display(mods_ls, title="Life Satisfaction")

preds_anx <- compare_predictors("PHQ4_Anxiety")
display(preds_anx$table, title="PHQ4 Anxiety")
preds_anx$p
mods_anx <- compare_models_lm("PHQ4_Anxiety")
display(mods_anx, title="PHQ4 Anxiety")

preds_dep <- compare_predictors("PHQ4_Depression")
display(preds_dep$table, title="PHQ4 Depression")
preds_dep$p
mods_dep <- compare_models_lm("PHQ4_Depression")
display(mods_dep, title="PHQ4 Depression")
```

```{r}
make_cor(select(df, LifeSatisfaction, starts_with("PHQ4_")),
         cbind(mint, metamint, select(df, starts_with("MAIA"), IAS, BPQ)))
```

- CaCo and RelA

### Emotions

```{r}
preds_arou <- compare_predictors("ERS_Arousal")
display(preds_arou$table, title="ERS Arousal")
preds_arou$p
mods_arou <- compare_models_lm("ERS_Arousal")
display(mods_arou, title="ERS Arousal")

preds_sens <- compare_predictors("ERS_Sensitivity")
display(preds_sens$table, title="ERS Sensitivity")
preds_sens$p
mods_sens <- compare_models_lm("ERS_Sensitivity")
display(mods_sens, title="ERS Sensitivity")

preds_pers <- compare_predictors("ERS_Persistence")
display(preds_pers$table, title="ERS Persistence")
preds_pers$p
mods_pers <- compare_models_lm("ERS_Persistence")
display(mods_pers, title="ERS Persistence")
```

```{r}
make_cor(select(df, starts_with("ERS_")),
         cbind(mint, metamint, select(df, starts_with("MAIA"), IAS, BPQ)))
```

- CaCo

:::

## Exploratory

::: {.panel-tabset}

### CEFSA

```{r}
make_cor(select(df, starts_with("MINT"), -matches("[[:digit:]]")), 
         select(df, starts_with("CEFSA")))

make_cor(select(df, starts_with("MAIA"), IAS, BPQ), 
         select(df, starts_with("CEFSA")))
```

```{r}
preds_cefsabody <- compare_predictors("CEFSA_Body")
display(preds_cefsabody$table, title="CEFSA - Body")
preds_cefsabody$p
mods_cefsabody <- compare_models_lm("CEFSA_Body")
display(mods_cefsabody, title="CEFSA - Body")

preds_cefsaself <- compare_predictors("CEFSA_Self")
display(preds_cefsaself$table, title="CEFSA - Self")
preds_cefsaself$p
mods_cefsaself <- compare_models_lm("CEFSA_Self")
display(mods_cefsaself, title="CEFSA - Self")

preds_cefsaemo <- compare_predictors("CEFSA_Emotion")
display(preds_cefsaemo$table, title="CEFSA - Emotion")
preds_cefsaemo$p
mods_cefsaemo <- compare_models_lm("CEFSA_Emotion")
display(mods_cefsaemo, title="CEFSA - Emotion")

preds_cefsareal <- compare_predictors("CEFSA_Reality")
display(preds_cefsareal$table, title="CEFSA - Reality")
preds_cefsareal$p
mods_cefsareal <- compare_models_lm("CEFSA_Reality")
display(mods_cefsareal, title="CEFSA - Reality")

preds_cefsacon <- compare_predictors("CEFSA_Connection")
display(preds_cefsacon$table, title="CEFSA - Connection")
preds_cefsacon$p
mods_cefsacon <- compare_models_lm("CEFSA_Connection")
display(mods_cefsacon, title="CEFSA - Connection")

preds_cefsagency <- compare_predictors("CEFSA_Agency")
display(preds_cefsagency$table, title="CEFSA - Agency")
preds_cefsagency$p
mods_cefsagency <- compare_models_lm("CEFSA_Agency")
display(mods_cefsagency, title="CEFSA - Agency")
```

### Emotion Regulation

```{r}
make_cor(select(df, starts_with("MINT"), -matches("[[:digit:]]")), 
         select(df, starts_with("CERQ")))

make_cor(select(df, starts_with("MAIA"), IAS, BPQ), 
         select(df, starts_with("CERQ")))
```

```{r}
preds_cerqreapp <- compare_predictors("CERQ_PositiveReappraisal")
display(preds_cerqreapp$table, title="CERQ - Positive Reappraisal")
preds_cerqreapp$p
mods_cerqreapp <- compare_models_lm("CERQ_PositiveReappraisal")
display(mods_cerqreapp, title="CERQ - Positive Reappraisal")

preds_cerqrefocus <- compare_predictors("CERQ_PositiveRefocusing")
display(preds_cerqrefocus$table, title="CERQ - Positive Refocusing")
preds_cerqrefocus$p
mods_cerqrefocus <- compare_models_lm("CERQ_PositiveRefocusing")
display(mods_cerqrefocus, title="CERQ - Positive Refocusing")

preds_cerqpersp <- compare_predictors("CERQ_Perspective")
display(preds_cerqpersp$table, title="CERQ - Perspective")
preds_cerqpersp$p
mods_cerqpersp <- compare_models_lm("CERQ_Perspective")
display(mods_cerqpersp, title="CERQ - Perspective")

preds_cerqacceptance <- compare_predictors("CERQ_Acceptance")
display(preds_cerqacceptance$table, title="CERQ - Acceptance")
preds_cerqacceptance$p
mods_cerqacceptance <- compare_models_lm("CERQ_Acceptance")
display(mods_cerqacceptance, title="CERQ - Acceptance")
```

### Primals

```{r}
make_cor(select(df, starts_with("MINT"), -matches("[[:digit:]]")), 
         select(df, starts_with("PI")))

make_cor(select(df, starts_with("MAIA"), IAS, BPQ), 
         select(df, starts_with("PI")))
```

```{r}
preds_pialive <- compare_predictors("PI_Alive")
display(preds_pialive$table, title="Primals - Alive")
preds_pialive$p
mods_pialive <- compare_models_lm("PI_Alive")
display(mods_pialive, title="Primals - Alive")

preds_pisafe <- compare_predictors("PI_Safe")
display(preds_pisafe$table, title="Primals - Safe")
preds_pisafe$p
mods_pisafe <- compare_models_lm("PI_Safe")
display(mods_pisafe, title="Primals - Safe")

preds_pient <- compare_predictors("PI_Enticing")
display(preds_pient$table, title="Primals - Enticing")
preds_pient$p
mods_pient <- compare_models_lm("PI_Enticing")
display(mods_pient, title="Primals - Enticing")

preds_pigood <- compare_predictors("PI_Good")
display(preds_pigood$table, title="Primals - Good")
preds_pigood$p
mods_pigood <- compare_models_lm("PI_Good")
display(mods_pigood, title="Primals - Good")
```

:::

## Predictive

### Mental Health

::: panel-tabset

#### Mood Disorders

```{r}
compare_models_glm <- function(outcome="Disorders_Psychiatric_Mood", title="", subtitle = "") {

  # Full Questionnaires
  m_mint <- glm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(mint, df[outcome]),
              family = "binomial")
  m_metamint <- glm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(metamint, df[outcome]),
              family = "binomial")
  m_minimint <- glm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(minimint, df[outcome]),
              family = "binomial") 
  m_micromint <- glm(as.formula(paste0(outcome, " ~ .")),
              data = cbind(micromint, df[outcome]),
              family = "binomial") 
  m_ias <- glm(as.formula(paste0(outcome, " ~ IAS")),
              data = df,
              family = "binomial")
  m_maia <- glm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), starts_with("MAIA")),
              family = "binomial") 
  m_imaia <- glm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome), 
                              starts_with("MAIA"), 
                              -contains("NotWorrying"),
                              -contains("NotDistracting")),
              family = "binomial")
  m_imaia <- glm(as.formula(paste0(outcome, " ~ .")),
                data = select(df, all_of(outcome),
                              # MAIA_AttentionRegulation, # remove?
                              # MAIA_SelfRegulation, # remove?
                              # MAIA_Trusting, # remove?
                              MAIA_BodyListening, 
                              MAIA_EmotionalAwareness,
                              MAIA_Noticing),
              family = "binomial")
  m_bpq <- glm(as.formula(paste0(outcome, " ~ BPQ")),
               data = df,
               family = "binomial")

  models_full <- list(m_mint=m_mint, m_metamint=m_metamint, 
                      m_minimint=m_minimint, m_micromint=m_micromint,
                      m_maia=m_maia, m_imaia=m_imaia, 
                      m_ias=m_ias, m_bpq=m_bpq)

  table <- table_models(models_full)

  get_auc <- function(m) {
    roc <- as.data.frame(performance_roc(m))
    auc <- bayestestR::area_under_curve(roc$Specificity, roc$Sensitivity)
    paste0("AUC = ", insight::format_percent(auc))
  }

  mods_order <- c("Mint", "metaMint", "miniMint", "microMint", "MAIA", "iMAIA", "IAS", "BPQ")

  p <- rbind(
    mutate(as.data.frame(performance_roc(m_mint)), Predictor = "Mint"),
    mutate(as.data.frame(performance_roc(m_metamint)), Predictor = "metaMint"),
    mutate(as.data.frame(performance_roc(m_minimint)), Predictor = "miniMint"),
    mutate(as.data.frame(performance_roc(m_micromint)), Predictor = "microMint"),
    mutate(as.data.frame(performance_roc(m_maia)), Predictor = "MAIA"),
    mutate(as.data.frame(performance_roc(m_imaia)), Predictor = "iMAIA"),
    mutate(as.data.frame(performance_roc(m_ias)), Predictor = "IAS"),
    mutate(as.data.frame(performance_roc(m_bpq)), Predictor = "BPQ")
  ) |>
    mutate(Predictor = fct_relevel(Predictor, mods_order)) |>
    ggplot(aes(x=Specificity)) +
    geom_abline(intercept=0, slope=1, color="gray", linewidth=1) +
    geom_line(aes(y=Sensitivity, color=Predictor), linewidth=1.5, alpha=0.9) +
    geom_label(data=data.frame(
      label = c(get_auc(m_mint), get_auc(m_metamint), 
                get_auc(m_minimint), get_auc(m_micromint), 
                get_auc(m_maia), get_auc(m_imaia), get_auc(m_ias), get_auc(m_bpq)),
      Predictor = mods_order,
      x = 0.8,
      y = c(0.55, 0.475, 0.425, 0.375, 0.30, 0.25, 0.20, 0.15)),
      aes(x=x, y=y, label=label), color=colors[mods_order], size=3) +
    labs(x= "1 - Specificity (False Positive Rate)", y="Sensitivity (True Positive Rate)",
         color="Questionnaire", title = title, subtitle=subtitle) +
    scale_color_manual(values=colors) +
    scale_x_continuous(labels=scales::percent) +
    scale_y_continuous(labels=scales::percent) +
    theme_minimal() +
    theme(plot.title = element_text(face="bold"))

  list(table=table, p=p)
}
```

```{r}
preds_mood <- compare_predictors("Disorders_Psychiatric_Mood")
display(preds_mood$table, title="Mood Disorders")
preds_mood$p
mods_mood <- compare_models_glm("Disorders_Psychiatric_Mood")
display(mods_mood$table, title="Mood Disorders")
mods_mood$p
```

```{r}
preds_moodt <- compare_predictors("Disorders_Psychiatric_MoodTreatment")
display(preds_moodt$table, title="Mood Disorders (with Treatment)")
preds_moodt$p
mods_moodt <- compare_models_glm("Disorders_Psychiatric_MoodTreatment",
                                 title="Mood Disorders", subtitle = "MDD, GAD, Bipolar (with Treatment)")
display(mods_moodt$table, title="Mood Disorders (with Treatment)")
mods_moodt$p
```

#### Anxiety

```{r}
preds_anxdis <- compare_predictors("Disorders_Psychiatric_Anxiety")
display(preds_anxdis$table, title="Anxiety Disorders")
preds_anxdis$p
mods_anxdis <- compare_models_glm("Disorders_Psychiatric_Anxiety")
display(mods_anxdis$table, title="Anxiety Disorders")
mods_anxdis$p
```

```{r}
preds_anxdist <- compare_predictors("Disorders_Psychiatric_AnxietyTreatment")
display(preds_anxdist$table, title="Anxiety Disorders (with Treatment)")
preds_anxdist$p
mods_anxdist <- compare_models_glm("Disorders_Psychiatric_AnxietyTreatment")
display(mods_anxdist$table, title="Anxiety Disorders (with Treatment)")
mods_anxdist$p
```

#### Eating

```{r}
preds_eat <- compare_predictors("Disorders_Psychiatric_Eating")
display(preds_eat$table, title="Eating Disorders")
preds_eat$p
mods_eat <- compare_models_glm("Disorders_Psychiatric_Eating")
display(mods_eat$table, title="Eating Disorders")
mods_eat$p
```

#### Addiction

```{r}
preds_addict <- compare_predictors("Disorders_Psychiatric_Addiction")
display(preds_addict$table, title="Addiction")
preds_addict$p
mods_addict <- compare_models_glm("Disorders_Psychiatric_Addiction")
display(mods_addict$table, title="Addiction")
mods_addict$p
```

#### ADHD

```{r}
preds_adhd <- compare_predictors("Disorders_Psychiatric_ADHD")
display(preds_adhd$table, title="ADHD")
preds_adhd$p
mods_adhd <- compare_models_glm("Disorders_Psychiatric_ADHD",
                                title = "ADHD", subtitle = "Attention Deficit/Hyperactivity Disorder")
display(mods_adhd$table, title="ADHD")
mods_adhd$p
```

#### Autism

```{r}
preds_asd <- compare_predictors("Disorders_Psychiatric_Autism")
display(preds_asd$table, title="Autism")
preds_asd$p
mods_asd <- compare_models_glm("Disorders_Psychiatric_Autism",
                                title = "ASD", subtitle = "Autism Spectrum Disorder")
display(mods_asd$table, title="Autism")
mods_asd$p
```

#### Borderline

```{r}
preds_bpd <- compare_predictors("Disorders_Psychiatric_Borderline")
display(preds_bpd$table, title="Borderline Personality Disorder")
preds_bpd$p
mods_bpd <- compare_models_glm("Disorders_Psychiatric_Borderline")
display(mods_bpd$table, title="Borderline Personality Disorder")
mods_bpd$p
```

::::

### Somatic Issues

::: panel-tabset

#### Afferent Sensitivity

```{r}
preds_afferent <- compare_predictors("Disorders_Somatic_AfferentSensitivity")
display(preds_afferent$table, title="Afferent Sensitivity Symptoms")
preds_afferent$p
mods_afferent <- compare_models_glm("Disorders_Somatic_AfferentSensitivity",
                                    title="Afferent Sensitivity Symptoms",
                                    subtitle = "Migraine, neuropathy, muscle tension, dizziness, ...")
display(mods_afferent$table, title="Afferent Sensitivity Symptoms")
mods_afferent$p
```

#### Central Sensitization

```{r}
preds_central <- compare_predictors("Disorders_Somatic_CentralSensitization")
display(preds_central$table, title="Central Sensitization Symptoms")
preds_central$p
mods_central <- compare_models_glm("Disorders_Somatic_CentralSensitization",
                                   title="Central Sensitization Symptoms",
                                   subtitle = "Fibromyalgia, chronic fatigue, back pain, IBS, ...")
display(mods_central$table, title="Central Sensitization Symptoms")
mods_central$p
```

#### Autonomic Dysfunction

```{r}
preds_autonomic <- compare_predictors("Disorders_Somatic_AutonomicDysfunction")
display(preds_autonomic$table, title="Autonomic Dysfunction")
preds_autonomic$p
mods_autonomic <- compare_models_glm("Disorders_Somatic_AutonomicDysfunction",
                                     title="Autonomic Dysfunction",
                                     subtitle = "Hypermobility, chest pain, hypo/hypertension, ....")
display(mods_autonomic$table, title="Autonomic Dysfunction")
mods_autonomic$p
```

#### Immune-Inflammatory

```{r}
preds_immune <- compare_predictors("Disorders_Somatic_ImmuneInflammatory")
display(preds_immune$table, title="Immune-Inflammatory Symptoms")
preds_immune$p
mods_immune <- compare_models_glm("Disorders_Somatic_ImmuneInflammatory",
                                  title="Immune-Inflammatory Symptoms",
                                  subtitle = "Allergies, asthma, eczema, autoimmune, ...")
display(mods_immune$table, title="Immune-Inflammatory Symptoms")
mods_immune$p
```

:::

### Lifestyle and Demographic

::: {.panel-tabset}

#### Age

```{r}
preds_age <- compare_predictors("Age")
display(preds_age$table, title="Age")
preds_age$p
mods_age <- compare_models_lm("Age")
display(mods_age, title="Age")
```

#### Sex

```{r}
df$Sex <- ifelse(df$Gender == "Female", 1, ifelse(df$Gender == "Male", 0, NA))

preds_sex <- compare_predictors("Sex")
display(preds_sex$table, title="Sex")
preds_sex$p
mods_sex <- compare_models_glm("Sex")
display(mods_sex$table, title="Sex")
mods_sex$p
```

#### BMI

```{r}
preds_bmi <- compare_predictors("BMI")
display(preds_bmi$table, title="BMI")
preds_bmi$p
mods_bmi <- compare_models_lm("BMI")
display(mods_bmi, title="BMI")
```

#### Physical Activity

```{r}
preds_physactiv <- compare_predictors("Physical_Active")
display(preds_physactiv$table, title="Physical Activity")
preds_physactiv$p
mods_physactiv <- compare_models_lm("Physical_Active")
display(mods_physactiv, title="Physical Activity")
```

:::

### Biofeedback Usage (Wearables)

::: {.panel-tabset}

#### Number of Wearables

```{r}
preds_wearN <- compare_predictors("Wearables_Number", family = "poisson")
display(preds_wearN$table, title="Wearables Number")
preds_wearN$p
```

#### Heart

```{r}
df$Wearables_HeartOwn <- ifelse(df$Wearables_Heart == "Not owning", 0, 1) 
preds_wearHeartOwn <- compare_predictors("Wearables_HeartOwn")
display(preds_wearHeartOwn$table, title="Wearables Heart Ownership")
preds_wearHeartOwn$p
mods_wearHeartOwn <- compare_models_glm("Wearables_HeartOwn")
display(mods_wearHeartOwn$table, title="Wearables Heart Ownership")
mods_wearHeartOwn$p

preds_wearHeartImp <- compare_predictors("Wearables_HeartImportance")
display(preds_wearHeartImp$table, title="Wearables Heart Importance")
preds_wearHeartImp$p
mods_wearHeartImp <- compare_models_lm("Wearables_HeartImportance")
display(mods_wearHeartImp, title="Wearables Heart Importance")
```

#### Sleep

```{r}
df$Wearables_SleepOwn <- ifelse(df$Wearables_Sleep == "Not owning", 0, 1)
preds_wearSleepOwn <- compare_predictors("Wearables_SleepOwn")
display(preds_wearSleepOwn$table, title="Wearables Sleep Ownership")
preds_wearSleepOwn$p
mods_wearSleepOwn <- compare_models_glm("Wearables_SleepOwn")
display(mods_wearSleepOwn$table, title="Wearables Sleep Ownership")

preds_wearSleepImp <- compare_predictors("Wearables_SleepImportance")
display(preds_wearSleepImp$table, title="Wearables Sleep Importance")
preds_wearSleepImp$p
mods_wearSleepImp <- compare_models_lm("Wearables_SleepImportance")
display(mods_wearSleepImp, title="Wearables Sleep Importance")
```

#### Steps

```{r}
df$Wearables_StepsOwn <- ifelse(df$Wearables_Steps == "Not owning", 0, 1)
preds_wearStepsOwn <- compare_predictors("Wearables_StepsOwn")
display(preds_wearStepsOwn$table, title="Wearables Steps Ownership")
preds_wearStepsOwn$p
mods_wearStepsOwn <- compare_models_glm("Wearables_StepsOwn")
display(mods_wearStepsOwn$table, title="Wearables Steps Ownership")

preds_wearStepsImp <- compare_predictors("Wearables_StepsImportance")
display(preds_wearStepsImp$table, title="Wearables Steps Importance")
preds_wearStepsImp$p
mods_wearStepsImp <- compare_models_lm("Wearables_StepsImportance")
display(mods_wearStepsImp, title="Wearables Steps Importance")
```

#### Weight

```{r}
df$Wearables_WeightOwn <- ifelse(df$Wearables_Weight == "Not owning", 0, 1)
preds_wearWeightOwn <- compare_predictors("Wearables_WeightOwn")
display(preds_wearWeightOwn$table, title="Wearables Weight Ownership")
preds_wearWeightOwn$p
mods_wearWeightOwn <- compare_models_glm("Wearables_WeightOwn")
display(mods_wearWeightOwn$table, title="Wearables Weight Ownership")

preds_wearWeightImp <- compare_predictors("Wearables_WeightImportance")
display(preds_wearWeightImp$table, title="Wearables Weight Importance")
preds_wearWeightImp$p
mods_wearWeightImp <- compare_models_lm("Wearables_WeightImportance")
display(mods_wearWeightImp, title="Wearables Weight Importance")
```

:::

## Summary Table

```{r}
make_table1 <- function(preds=preds_dif, mods=mods_dif, group = "Alexithymia", outcome = "TAS - DIF") {
  yvar <- preds$outcome
  preds <- preds$table

  best_pred_mint <- preds[str_detect(preds$Name, "MINT"), ]
  best_pred_nonmint <- preds[!str_detect(preds$Name, "MINT"), ]
  best_pred_nonmint <- best_pred_nonmint[best_pred_nonmint$R2 == max(best_pred_nonmint$R2), ][1,]
  best_models_R2 <- str_remove(arrange(mods, desc(as.numeric(R2)))$Name, "m_")
  best_model_BIC <- str_remove(arrange(mods, as.numeric(BIC))$Name, "m_")

  if(length(unique(df[[yvar]])) > 2) {
    # LMs - display correlation
    r_mint <- cor.test(cbind(df, metamint)[[best_pred_mint$Name[1]]], df[[yvar]])
    r_mint <- paste0(insight::format_value(r_mint$estimate, zap_small = TRUE))
    r_nonmint <- cor.test(df[[best_pred_nonmint$Name[1]]], df[[yvar]])
    r_nonmint <- paste0(insight::format_value(r_nonmint$estimate, zap_small = TRUE))
  } else {
    # GLMs - display log-odds
    r_mint <- best_pred_mint$Coefficient
    r_nonmint <- best_pred_nonmint$Coefficient
  }

  # Format
  best_models_R2[str_detect(best_models_R2, "mint")] <- paste0("<b>", best_models_R2[str_detect(best_models_R2, "mint")], "</b>")
  best_model_BIC[str_detect(best_model_BIC, "mint")] <- paste0("<b>", best_model_BIC[str_detect(best_model_BIC, "mint")], "</b>")
  best_models_R2 <- paste0("<small>", paste0(best_models_R2, collapse = " > "), "</small>")
  best_model_BIC <- paste0("<small>", paste0(best_model_BIC, collapse = " > "), "</small>")

  data.frame(
    Group = group,
    Outcome = outcome, 
    Best_MINT_Predictor = paste0(str_remove(best_pred_mint$Name[1], "MINT_"), "<br><small><i>(", r_mint, ")</i></small>"),
    Best_NonMINT_Predictor = paste0(str_replace(best_pred_nonmint$Name[1], "_", " - "), "<br><small><i>(",  r_nonmint, ")</i></small>"),
    Best_Models_R2 = best_models_R2,
    Best_Models_BIC = best_model_BIC
    )
}

rez <- rbind(
  make_table1(preds_dif, mods_dif, group = "Alexithymia", outcome = "Difficulty Identifying Feelings (DIF)<small><p align='right'><i>TAS-20</i></small><p>"),
  make_table1(preds_ddf, mods_ddf, group = "Alexithymia", outcome = "Difficulty Describing Feelings (DDF)<small><p align='right'><i>TAS-20</i></small><p>"),
  make_table1(preds_eot, mods_eot, group = "Alexithymia", outcome = "Externally Oriented Thinking (EOT)<small><p align='right'><i>TAS-20</i></small><p>"),

  make_table1(preds_arou, mods_arou, group = "Emotion Reactivity", outcome = "Arousal<small><p align='right'><i>ERS</i></small><p>"),
  make_table1(preds_sens, mods_sens, group = "Emotion Reactivity", outcome = "Sensitivity<small><p align='right'><i>ERS</i></small><p>"),
  make_table1(preds_pers, mods_pers, group = "Emotion Reactivity", outcome = "Persistence<small><p align='right'><i>ERS</i></small><p>"),

  make_table1(preds_ls, mods_ls, group = "Mood", outcome = "Life Satisfaction"),
  make_table1(preds_anx, mods_anx, group = "Mood", outcome = "Anxiety<small><p align='right'><i>PHQ-4</i></small><p>"),
  make_table1(preds_dep, mods_dep, group = "Mood", outcome = "Depression<small><p align='right'><i>PHQ-4</i></small><p>"),

  make_table1(preds_moodt, mods_moodt$table, group = "Mental Health", outcome = "Mood Disorder"),
  make_table1(preds_adhd, mods_adhd$table, group = "Mental Health", outcome = "ADHD"),
  make_table1(preds_asd, mods_asd$table, group = "Mental Health", outcome = "Autism"),

  make_table1(preds_afferent, mods_afferent$table, group = "Somatic Health", outcome = "Afferent Sensitivity<small><small><p align='right'><i>Migraine, neuropathy, muscle tension, dizziness, ...</i></small></small><p>"),
  make_table1(preds_central, mods_central$table, group = "Somatic Health", outcome = "Central Sensitization<small><small><p align='right'><i>Fibromyalgia, chronic fatigue, back pain, IBS, ...</i></small></small><p>"),
  make_table1(preds_autonomic, mods_autonomic$table, group = "Somatic Health", outcome = "Autonomic Dysfunction<small><small><p align='right'><i>Hypermobility, chest pain, hypo/hypertension, ...</i></small></small><p>"),
  make_table1(preds_immune, mods_immune$table, group = "Somatic Health", outcome = "Immune-Inflammatory<small><small><p align='right'><i>Allergies, eczema, autoimmune, ...</i></small></small><p>"),

  make_table1(preds_cefsabody, mods_cefsabody, group = "Dissociative Symptoms", outcome = "Body<small><p align='right'><i>CEFSA</i></small><p>"),
  make_table1(preds_cefsaself, mods_cefsaself, group = "Dissociative Symptoms", outcome = "Self<small><p align='right'><i>CEFSA</i></small><p>"),
  make_table1(preds_cefsaemo, mods_cefsaemo, group = "Dissociative Symptoms", outcome = "Emotions<small><p align='right'><i>CEFSA</i></small><p>"),
  make_table1(preds_cefsareal, mods_cefsareal, group = "Dissociative Symptoms", outcome = "Reality<small><p align='right'><i>CEFSA</i></small><p>"),

  # make_table1(preds_pialive, mods_pialive, group = "Primals", outcome = "Alive<small><p align='right'><i>PI-18</i></small><p>"),
  # make_table1(preds_pisafe, mods_pisafe, group = "Primals", outcome = "Safe<small><p align='right'><i>PI-18</i></small><p>"),
  # make_table1(preds_pient, mods_pient, group = "Primals", outcome = "Enticing<small><p align='right'><i>PI-18</i></small><p>"),
  # make_table1(preds_pigood, mods_pigood, group = "Primals", outcome = "Good<small><p align='right'><i>PI-18</i></small><p>")

  make_table1(preds_bmi, mods_bmi, group = "Lifestyle", outcome = "BMI"),
  make_table1(preds_physactiv, mods_physactiv, group = "Lifestyle", outcome = "Physical Activity"),
  # make_table1(preds_age, mods_age, group = "Lifestyle", outcome = "Age"),
  make_table1(preds_wearHeartImp, mods_wearHeartImp, group = "Lifestyle", outcome = "Cardiac Monitoring"),
  make_table1(preds_wearSleepImp, mods_wearSleepImp, group = "Lifestyle", outcome = "Sleep Monitoring"),
  make_table1(preds_wearStepsImp, mods_wearStepsImp, group = "Lifestyle", outcome = "Steps Monitoring")
  # make_table1(preds_wearWeightImp, mods_wearWeightImp, group = "Lifestyle", outcome = "Weight Monitoring")
)

tab2 <- rez |> 
  gt::gt() |> 
  gt::tab_header(
    title = gt::md("**Interoceptive Questionnaires Comparison**"),
    subtitle = "MINT vs. MAIA, IAS, BPQ"
  ) |> 
  gt::cols_width(
    Outcome ~ gt::pct(23),
    Best_MINT_Predictor ~ gt::pct(12),
    Best_NonMINT_Predictor ~ gt::pct(15),
    Best_Models_R2 ~ gt::pct(25),
    Best_Models_BIC ~ gt::pct(25)
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight="bold", v_align = "middle"),
    locations = gt::cells_column_labels()
  )  |>
  gt::tab_style(
    style = gt::cell_text(align = "center", v_align = "middle"),
    locations = gt::cells_body(columns=c("Best_MINT_Predictor", "Best_NonMINT_Predictor"))
  )  |>
  gt::tab_row_group(
    label = "Lifestyle",
    rows = Group == "Lifestyle"
  ) |> 
  gt::tab_row_group(
    label = "Dissociative Symptoms",
    rows = Group == "Dissociative Symptoms"
  ) |> 
  gt::tab_row_group(
    label = "Primal World Beliefs",
    rows = Group == "Primals"
  ) |> 
  gt::tab_row_group(
    label = "Somatic Health",
    rows = Group == "Somatic Health"
  ) |> 
  gt::tab_row_group(
    label = "Mental Health",
    rows = Group == "Mental Health"
  ) |>
  gt::tab_row_group(
    label = "Mood",
    rows = Group == "Mood"
  ) |> 
  gt::tab_row_group(
    label = "Emotional Reactivity",
    rows = Group == "Emotion Reactivity"
  ) |> 
  gt::tab_row_group(
    label = "Alexithymia",
    rows = Group == "Alexithymia"
  ) |> 
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "gray85"),
      gt::cell_text(weight="bold", style="italic")),
    locations = gt::cells_row_groups()
  )  |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(1:4, 11:18, 20:21, 24))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(5:9, 10, 19, 22))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#EF9A9A"),
    locations = list(
      gt::cells_body(columns="Best_MINT_Predictor", rows=c(23, 25))
    )
  ) |> 
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_Models_R2", rows=c(1:3, 11:13, 15:21, 23:25))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_Models_R2", rows=c(4:10, 14, 22))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#81C784"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(1:3, 11:20, 25))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#FFB74D"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(4:10, 22))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#EF9A9A"),
    locations = list(
      gt::cells_body(columns="Best_Models_BIC", rows=c(21, 23, 24))
    )
  ) |>
  gt::cols_label(
    Best_MINT_Predictor = "Best Predictor\n(MINT)",
    Best_NonMINT_Predictor = "Best Predictor\n(Non-MINT)",
    Best_Models_R2 = "Best Models (R2)",
    Best_Models_BIC = "Best Models (BIC)"
  ) |> 
  gt::cols_hide(Group) |> 
  gt::fmt_markdown() |> 
  gt::tab_options(
    table.font.size = gt::pct(125)
  )

tab2 |> 
  gt::tab_options(
    table.font.size = gt::pct(50)
  ) |> 
  gt::opt_interactive()

gt::gtsave(tab2, "figures/table2.png", vwidth=1400, expand=0, quiet=TRUE)
```

## Figures

```{r}
#| warning: false
#| eval: false

fig1a <- patchwork::wrap_elements(
  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/fig1a.png"))) +
  patchwork::plot_annotation(title = "Exploratory Graphical Analysis (EGA)",
                             subtitle = "Bootstrapped replication of hierarchical clusters (Methhod = leiden)",
                             theme = theme(plot.title = element_text(face="bold", hjust = 0, size = rel(1.4)),
                                           plot.subtitle = element_text(hjust = 0, size = rel(1)))))
fig1b <- patchwork::wrap_elements(
  patchwork::wrap_elements(grid::rasterGrob(png::readPNG("figures/fig1b.png"))) +
  patchwork::plot_annotation(title = "Hierarchical Clustering",
                             subtitle = "Method = Correlation",
                             theme = theme(plot.title = element_text(face="bold", hjust = 0, size = rel(1.4)),
                                           plot.subtitle = element_text(hjust = 0, size = rel(1)))))
ggsave("figures/fig1.png", fig1a / fig1b, width=8, height=12, dpi=300)

p_cor <- df |> 
  select(starts_with("TAS_"), LifeSatisfaction, starts_with("PHQ4_"), starts_with("CEFSA"), starts_with("PI"), starts_with("ERS_")) |> 
  mutate(LifeSatisfaction = datawizard::reverse_scale(LifeSatisfaction),
         PI_Safe = datawizard::reverse_scale(PI_Safe),
         PI_Good = datawizard::reverse_scale(PI_Good),
         PI_Enticing = datawizard::reverse_scale(PI_Enticing)) |>
  rename(`Life Satisfaction*` = LifeSatisfaction,
         `Primals - Alive` = PI_Alive,
         `Primals - Safe*` = PI_Safe,
         `Primals - Good*` = PI_Good,
         `Primals - Enticing*` = PI_Enticing) |>
  make_cor(cbind(mint, metamint, select(df, starts_with("MAIA"), IAS, BPQ))) +
  labs(title="Correlates of Interoception") 
p_cor

fig2 <- (p_cormint | p_corintero) / p_cor + plot_layout(heights = c(0.4, 0.6))
ggsave("figures/fig2.png", fig2, width=10, height=12, dpi=300)

fig3 <- (mods_moodt$p + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) | 
    mods_adhd$p + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) | 
    mods_asd$p + theme(axis.text.x = element_blank(), axis.title.x = element_blank())) / 
  (mods_central$p | mods_autonomic$p | mods_immune$p) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "Predictive Performance of Interoceptive Questionnaires", 
       subtitle = "ROC Curves",
       theme = theme(plot.title = element_text(face="bold", hjust = 0.5, size = rel(1.4)),
                     plot.subtitle = element_text(hjust = 0.5, size = rel(1.25))))
fig3
ggsave("figures/fig3.png", fig3, width=13, height=9, dpi=300)
```