### 2025-07-01 paper/manuscript.qmd

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  TO DO.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

# Introduction

TODO: write general intro.

Main issues in existing questionnaires:

- Either heavily based on theories (e.g., focusing on specific "alleged" dimensions), despite shaky evidence for said-theories
- Do not control for context (which leads to variability in interpretation and occurence)
- Often quite narrow in the modalities covered (or scattered, like the IAS)

# Study 1: Item Selection

TODO: write intro.

Goal of study 1: to generate a lot of items, analyze its structure and reduce them to a balanced set of items.

## Methods

### Participants

We recruited 760 English-speaking participants using Prolific\textcopyright. We excluded 191 for failing at least one attention check, and 10 based on measures significantly related to the probability of failing attention checks [namely, the multivariate distance obtained with the OPTICS algorithm, @theriault2024check]. The final sample includes 559 participants (age = 37.0 $\pm$ 12.2 [18, 77]; 50.8\% women; Country of residence: 63.86% UK, 26.65% USA).
This study was approved by the University of Sussex' Ethics Committee (ER/MB2021/1).

### Item Generation

Based on the two goals outlined for this scale, namely to include different interoceptive modalities, and to explicitly state the context of the interoceptive experience (e.g., whether negative or positive), we generated items in a systematic way following a combinatory approach, where each item's category corresponds to the combination of a specific modality and context (@fig-one).

```{r}
#| label: fig-one
#| fig-cap: "The conceptual grid used to generate the 120 initial items (top-left). Each item belong both to an interoceptive modalitiy and a facet, with the number of each item per category indicated in the circles. The askterisk denotes the additional presence of an attention check item in that category. In the experiment, these items were presented on different pages grouped either by modality (bottom-left), by facet (top-right), or entirely randomly. The Correlation Similarity (bottom-right) analysis suggested that the correlation matrix obtained from the participants assigned to the random-grouping condition was slightly more similar (but non-significantly) to the one obtained in the modality-grouping condition, suggesting that 1) the scale's structure is robust to different presentation conditions; 2) modality-grouping might tend to facilitate the emergence of the underlying item structure (and thus be interpreted as being more natural)."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../study1/analysis/figures/figure1.png")
```

We identified 7 "modalities" (cardiac, respiratory, gastric, genital, skin & temperature, bladder & colon, and a "general state" category corresponding to a holistic and general awareness of an interoceptive state or dimension). Through iterative refinement (e.g., splitting or merging different categories together), we settled on 6 "facets", which encompass both *contexts* of experience (negative and positive arousal, namely anxious and sexual states), and potential distinct *mechanisms* (nociception & pleasure, sensitivity, accuracy, and confusion).

Using this orthogonal 7x6 modality/facet grid as a conceptual scaffolding, we generated 120 initial items, striving for a balanced number of items with consistent phrasing within modalities and facets^[The initial item list is available at [realitybending.github.io/InteroceptionScale/study1/analysis/2_analysis.html](https://realitybending.github.io/InteroceptionScale/study1/analysis/2_analysis.html)]. We additionally crafted 8 "attention check" items blending in (and distributed across) each category.

### Procedure

To avoid presenting all the 120 items on a single long and discouraging page, we split them into different pages.
Participants were randomly assigned to one of three conditions, driving how items were grouped on the same page: 1) items grouped by modality (i.e., all cardiac items on the first page, all colon & bladder  items on the second, etc.), 2) items grouped by facet, or 3) items presented fully randomly (but balanced randomly across 6 pages). The order of the item on any given page and the order of the modalities/facets were randomized.
Each participant completed the full set of 120 items, with the 8 attention check items interspersed throughout.
The online experiment was implemented using JsPsych [@de2015jspsych], and item responses were recorded using 7-points Likert scales (0 = Disagree, 6 = Agree).

### Data Analysis

In order to test whether the grouping condition had an effect on the structure (i.e., how items relate to one-another), we compared the correlation matrix obtained in the random condition to the ones obtained in the modality and facet conditions, focusing on 3 indices of correlation matrix similarity - the Procrustes Similarity Index [PSI, @sibson1978studies], the Adjusted RV [Rvadj, @mayer2011exploratory], and the Similarity of Matrices Index [SMI, @indahl2018similarity]. For each index, we bootstrapped the difference between the similarity with the facet and modality conditions to test whether the correlation matrix in the random-grouping condition is significantly more similar to any of the two other conditions.

Items deemed "redundant" (which can distort the item structure estimation by introducing multicollinearity or local dependencies) will be identified (using the recommended threshold of 0.25) using Unique Variable Analysis [UVA, @christensen2023unique], a novel and principled method derived from network psychometrics.

The structure of the items will be analyzed using the recently-developed Exploratory Graph Analysis [EGA, @golino2017exploratory] framework, which allows to jointly estimate the number of dimensions (i.e., clusters of items), the structure, as well as its stability using bootstrapping [@golino2020investigating]. At a fundamental level, EGA conceptualizes variables as nodes in a network, with connections (edges) reflecting associations between them. Evidence has underlined its suitability as an alternative to traditional factor analysis, addressing some of its limitations such as the assumption of a "latent" source of variability, issues in estimation of the optimal factor numbers, and poor performance in complex population structures, while remaining comparable and interpretable [@christensen2021equivalency; @jimenez2023dimensionality]. In particular, nodes communities (i.e., clusters of items) can be in practice interpreted as distinct "dimensions", similarly to traditional latent factors - but without explicitly assuming their existence [@christensen2021equivalency].

After removing redundant items using UVA, we will iteratively fit hierarchical EGA models (which additionally estimates higher-order "meta" clusters) using "glasso" [**REF**] and the "leiden" algorithm [**REF**] for community detection, refining the item pool at each step. We will start by removing items with a low (< 80%) cluster stability (i.e., volatile items which jump between clusters across bootstrapped samples), followed by odd items belonging to no clusters or pairs of items (i.e., we keep items belonging to clusters of more than 2 items). Finally, for each lower-level cluster, we will select the 3 items with the highest node centrality (i.e., the highest loading in the cluster).

## Results

The correlation matrix similarity analysis yielded no significant differences between the similarity of the random-grouping condition with the modality-grouping and facet-grouping conditions ($PSI_{\text{Random vs. Facet}} = 0.81$, $PSI_{\text{Random vs. Modality}} = 0.82$, $p = .45$; $RVadj_{\text{Random vs. Facet}} = 0.77$, $RVadj_{\text{Random vs. Modality}} = 0.78$, $p = .74$; $SMI_{\text{Random vs. Facet}} = 0.49$, $SMI_{\text{Random vs. Modality}} = 0.51$, $p = .52$).

From the 120 initial items, UVA flagged 4 redundant items that we removed. We then removed 40 items that showed low cluster stability, and 9 items that were part of clusters with less than 3 items. Finally, We kept the 3 items with the highest loading in their lower-level structure (removing 13 items in the process), resulting in 54 items in the final item pool.

The final hierarchical EGA model (Generalized Total Entropy Fit Index = -119.18) - in which all 54 items yielded a high cluster stability (> 90%) suggested 3 metaclusters and 15 lower-level clusters (each containing 3 items): "Interoceptive Deficits" (containing 5 clusters: *Urointestinal Inaccuracy* - UrIn; *Cardiorespiratory Confusion* - CaCo; *Cardiorespiratory Noticing* - CaNo; *Olfactory Compensation* - Olfa; *Satiety Noticing* - Sati), "Interoceptive Awareness" (containing 7 clusters: *Sexual Arousal Awareness* - SexA; *Sexual Arousal Sensitivity* - SexS; *Sexual Organs Sensitivity* - SexO; *Urointestinal Sensitivity* - UrSe; *Relaxation Awareness* - RelA; *State Specificity* - StaS; *Expulsion Accuracy* - ExAc), and "Interoceptive Sensitivity" (containing 6 clusters: *Cardioception* - Card; *Respiroception* - Resp; *Signalling* - Sign; *Gastroception* - Gast; *Dermal Hypersensitivity* - Derm; *Sexual Arousal Changes* - SexC).

To further reduce and balance the remaining items, we collectively decided on removing the lower-level clusters *Sexual Arousal Awareness* (too general and overlapping with the other more specific sex-related items), *Signalling* (which items started with "when something important is happening in my life", which meaning we deemed too much open to interpretation), and *Sexual Arousal Changes* (low fit with the other modality-focused clusters of its group). The final set included 45 items.

<!-- Interoceptive Deficits -->
<!--   C01 = "UrIn",       # Urointestinal Inaccuracy -->
<!--   C09 = "CaCo",       # Cardiorespiratory Confusion -->
<!--   C13 = "CaNo",       # Cardiorespiratory Noticing -->
<!--   C18 = "Olfa",       # Olfactory Compensation -->
<!--   C03 = "Sati",       # Satiety Noticing -->
<!-- Interoceptive Awareness -->
<!--   C04 = "SexA",       # Sexual Arousal Awareness -->
<!--   C06 = "SexS",       # Sexual Arousal Sensitivity -->
<!--   C15 = "SexO",       # Sexual Organs Sensitivity -->
<!--   C08 = "UrSe",       # Urointestinal Sensitivity -->
<!--   C10 = "RelA",       # Relaxation Awareness -->
<!--   C05 = "StaS",       # State Specificity -->
<!--   C02 = "ExAc",       # Expulsion Accuracy -->
<!-- Interoceptive Sensitivity -->
<!--   C17 = "Card",       # Cardioception -->
<!--   C14 = "Resp",       # Respiroception -->
<!--   C12 = "Sign",       # Signalling -->
<!--   C16 = "Gast",       # Gastroception -->
<!--   C11 = "Derm",       # Dermal Hypersensitivity -->
<!--   C07 = "SexC"        # Sexual Arousal Changes -->
<!-- ) -->

## Discussion

Possible (although not at all significant) bias consistent across indices in favour of a greater similarity with the modality-grouping condition.

We grouped some of the dimensions under "Awareness", despite being similar to items included in the IAS typically conceptualized as "Accuracy". However, we believe that subjective scales cannot be pure measures of accuracy, as they are inherently subjective and thus reflect a degree of awareness (of one's accurate noticing or sensation).

# Study 2: Validation

TODO: write intro.
Goal of study 2: to validate the Mint scale against other interoception scales.

## Methods

### Participants

We recruited 921 English-speaking participants via Prolific\textcopyright and SONA, from which 118 were excluded for failing at least one attention check and 6 based on multivariate distance (using the same procedure as for study 1). 
60 participants were further excluded due to missing data following a technical error.
The final sample includes 737 participants (age = 36.8 $\pm$ 14.7 [18, 87]; 57.3\% women; Country of residence: 75.17% United Kingdom, 24.83% other).
This study was approved by the University of Sussex' Ethics Committee (ER/EB672/2).

### Procedure

The experiment included the following questionnaires.

**TODO: What about attention checks?**

#### Interoception Questionnaires

##### The Multimodal Interoception Questionnaire (Mint)

The 45 items of the Mint scale were presented in a random order, with the same 7-point Likert scale as in study 1 (0 = Disagree, 6 = Agree).

##### Multidimensional Assessment of Interoceptive Awareness - Second Version (MAIA-2)

The 37 items of the MAIA-2 [@mehling2018multidimensional] on a 7-point Likert scale from *Never* (0) to *Always* (6). The MAIA-2 consists of 8 scales: Noticing, Not-Distracting, Not-Worrying, Attention Regulation, Emotional Awareness, Self-Regulation, Body Listening and Trust. The scales address 5 dimensions of body awareness: Awareness of Body Sensations, Emotional Reaction and Attentional Response to Sensations, Capacity to Regulate Attention, Awareness of Mind-Body Integration, Trusting Body Sensations.

##### The Body Perception Questionnaire - Very Short Form (BPQ- VSF)

The 12 item BPQ-VSF [@cabrera2018assessing] related to bodily awareness on a 5-point Likert scale from *Never* (0) to *Always* (5).

##### The Interoceptive Accuracy Scale (IAS)

The 21 items of the IAS [@murphy2020testing] assessing interoceptive accuracy to physical sensations on a 5-point Likert scale ranging from *Disagree Strongly* (1) to *Strongly Agree*.

#### Emotions and Cognition

##### Toronto Alexithymia Scale (TAS-20)

The 20-items TAS [@leising2009toronto] assessing Alexithymia traits grouped into three dimensions: Difficulty Identifying Feelings, Difficulty Describing Feelings, and Externally Oriented Thinking. Items are rated in a 5-point Likert scale from *Strongly Disagree* (1) to *Strongly Agree*(5).

##### Emotion Reactivity Scale - Brief Version (B-ERS)

The 6-item B-ERS [@veilleux2024development] assessing emotional sensitivity on a 5-point Likert scale from *Not like me at all* (0) to *Extremely like me* (4).

##### Cognitive Emotion Regulation Questionnaire-short (CERQ-short)

The 18 items CERQ-short [@garnefski2006cognitive] on a 5-point Likert scale ranging from *Almost Never* (1) to *Almost Always* (5). This questionnaire measures eight cognitive responses to threatening or stressful events, including: Self-blame, Other-blame, Rumination, Catastrophizing, Putting into Perspective, Positive Refocusing, Positive Reappraisal, Acceptance, and Planning.

##### Černis Felt Sense of Anomaly scale - short-form (CEFSA-14)

The 14 item CEFSA [@vcernis2024measuring] assessing a specific sub-type of dissociative experiences in adolescence and adulthood - i.e., felt sense of anomaly-type
dissociation. the items are rated ON A 5-point Likert scale from *Never* (0) to *Always* (4), forming seven factors: Anomalous Experience of the Self, Anomalous Experience of the Body , Anomalous Experiences of Emotion , Altered Sense of Familiarity , Altered Sense of Connection, Altered Sense of Agency , and Altered Sense of Reality

##### Primal World Beliefs (PI-18)

the 18 items PI [@clifton2021brief] measuring highest-order primal world beliefs - i.e., beliefs about the basic character of the world: Good, Safe, Enticing, Alive. Responses are scored on a 6-point Likert scale from *Strongly Disagree* (0) to *Strongly Agree* (5).

#### Health and Wellbeing

##### Life Satisfaction, Depression and Anxiety (PHQ-4)

The 4 item scale of the PHQ [@kroenke2009ultra] assessing depression and anxiety, with responses scored as *Not at all* (0), *Once or twice* (0.5), *Several days* (1), *More than half of the days* (2), *nearly every day* (3). This revised version added as a response option (i.e., Once or twice) as a way to improve the scale's sensitivity to milder mood fluctuations [@makowskimeasuring].

<!-- PHQ-4 + LS: presented together -->

##### Mental Health

Participants were asked to self-report any current, medically diagnosed psychiatric disorders using a checklist based on DSM-5 categories. If one or more conditions were endorsed, participants were asked to indicate any current treatments, including pharmacological (e.g., antidepressants, anxiolytics, antipsychotics, mood stabilizers), psychological (e.g., psychotherapy, mindfulness), or lifestyle interventions. Binary variables (0 = absent, 1 = present) were created to identify participants reporting mood disorders (MDD, GAD, Bipolar Disorder; with a stricter subgroup of participants undergoing a pharma- or psychological treatment), anxiety-centred disorders (GAD, Panic Disorder, Social Anxiety Disorder, Specific Phobias), eating disorders, addiction-related disorders, borderline personality disorder, autism spectrum disorder (ASD), and ADHD.

##### Somatic Health

Participants were asked to select somatic symptoms or conditions they experienced from a list of 36 options.
To facilitate mechanistic interpretation and reduce redundancy, answers were grouped into four non-overlapping clusters based on shared physiological pathways and known etiological mechanisms. The *Afferent Sensitivity* cluster included conditions associated with heightened interoceptive awareness and neurogenic excitability, such as migraine, neuropathy, dizziness, nausea, muscle tension, epilepsy, and frequent urination. The *Central Sensitization* cluster comprised syndromes characterized by chronic pain and fatigue, likely reflecting central amplification of sensory signals and HPA-axis dysregulation, such as fibromyalgia, chronic fatigue syndrome, chronic pain, back pain, pelvic pain, irritable bowel syndrome (IBS). The *Autonomic Dysfunction* cluster captured disorders linked to dysregulation of the autonomic and cardiopulmonary systems, including joint hypermobility, cardiac arrhythmia, chest pain, shortness of breath, hypo-/hypertension, sleep apnea, chronic obstructive pulmonary disease (COPD), and chronic bronchitis. Finally, the *Immune-Inflammatory* cluster encompassed conditions associated with immune dysregulation, barrier dysfunction, and gut-brain axis disturbance, such as eczema, psoriasis, skin rashes, asthma, celiac disease, gluten and lactose sensitivity, inflammatory bowel diseases (Crohn’s disease and ulcerative colitis), gastroesophageal reflux disease (GERD), multiple sclerosis, and Sjögren’s syndrome. 
<!-- Each symptom was assigned to a single cluster to preserve orthogonality and support subsequent analyses linking symptom profiles to interoceptive processes. -->
We scored each cluster as a binary variable based on whether the participant selected at least one symptom from that cluster.

##### Patient Health Questionnaire - Somatic Symptoms (PHQ-15)

The PHQ-15 [@kroenke2002phq] is comprised of 15 somatic symptoms, with responses scored as *Not at all* (0), *Bothered a little* (1) and *bothered a lot* (2).

#### Lifestyle

Participants were asked to report on their use of wearable devices to monitor several health indices: heart rate, number of steps, calories burned, sleep quality, and weight (including via a regular scale). For each selected index, a question appeared asking them to rate how often they check that index; responses were scored as *Never*, *Very rarely*, *A few times per week*, *Once a day*, *A few times per day*, and *A few times per hour*. Additionally, participants were asked to rate how important this information was to them on an 8-point Likert scale ranging from *Not at all* (0) to *Very much* (7).

Questions regarding physical activity were also included. Participants were asked to rate how physically active they consider themselves on an 8-point Likert scale, from *Not at all* (0) to *Very much* (7), and to self-report how many hours of active workout they engage in per week.

#### Randomization

In order to avoid the repetition of similar types of questions and and balance longer and shorter questionnaires, we partitioned them into three groups (and randomized their order within them): (1) interoception questionnaires (MAIA, IAS, BPQ), (2) emotions (TAS, CERQ, ERS, PI-18), and (3) health (somatic symptoms, mental health, LS + PHQ-4, CEFSA). 
After completing demographic questions, participants always started with the Mint scale, and each following interoception questionnaire was interspersed with two questionnaires from the emotions and pathology groups.
In order to make the experiment more enjoyable, the experiment ended with a radar chart summarizing the participants' responses to the Mint scale^[The experiment can be tested by following the link on https://github.com/RealityBending/InteroceptionScale].

### Data Analysis

We will start by confirming and further refining the structure of the Mint scale (see @fig-two) using the same EGA model as in Study 1. 
We will then compute the lower-level dimensions as well as the higher-level metaclusters' scores by averaging the items in each cluster.
The convergent validity of the final set of items will be assessed by computing the correlations between the Mint scale and the other interoception questionnaires (MAIA, BPQ, IAS).

```{r}
#| label: fig-two
#| fig-cap: "TODO."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../study2/analysis/figures/fig1.png")
```

Next, we will test the predictive power of the Mint scale relative to the other interoception questionnaires.
We will first assess the importance of each interoceptive dimension (from all the scales) as a unique predictor by fitting regression models (linear for continuous measures - e.g., depression score - and logistic for binary variables - e.g., presence *vs.* absence of mood disorders) to predict various outcomes, and compare the best between the 4 interoception scales (based on the R2).
We will then assess the predictive performance of each scale as a whole by comparing regression models with all the dimensions entered as predictors (note that the IAS and BPQ only have one total score variable). We will evaluate models based on their total explained variance (R2), as well as on the Bayesian Information Criterion (BIC), which penalizes for predictor numbers, thus offering a balance between model parsimony and predictive power. For the logistic model, we will quantify the discriminative power by computing the Area Under the Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which assesses the model's discriminative power (the combination of sensitivity and specificity).

In order to evaluate the potential of a short version of our scale, we will compare 4 variants of the Mint models: the (full) Mint (including all lower-level dimensions), the "metaMint" (only including the metaclusters), the "miniMint" (including the metaclusters computed from a reduced number of items), and the "microMint" (including only the most representative dimension of each metacluster).
Moreover, we will also include an alternative "interoception-only" version of the MAIA ("iMAIA") that only includes the 3 most interoceptive dimensions (Body Listening, Noticing, and Emotional Awareness).

**TODO: Ana: do you think that this selection of dims for the iMAIA is fair? Or should I include other dimensions that you think are also very low-level and interoceptive?**

## Results

The first application of EGA to the initial set of 42 items, which reproduced the expected 14 lower-level clusters of triplets and 3 higher-order metaclusters, with the exception of the *UrSe* (Urointestinal Sensitivity), for which one item moved to the *Olfa* (Olfactory Compensation) cluster.
In order to further balance and reduce the items, we removedt the *UrSe* cluster, as well as *StaS* (State Specificity; which in comparison to the other items stood out as containing vague and overlapping items (e.g., "Being sexually aroused is a very different bodily feeling compared to other states (e.g., feeling anxious, relaxed, or after physical exercise)"); *SexO* (Sexual Organs Sensitivity), which overlapped with the *SexS* (Sexual Arousal Sensitivity) cluster; and *CaNo* (Cardiorespiratory Noticing), which overlapped with the *CaCo* (Cardiorespiratory Confusion) cluster.

The final set of 36 items yielded a good fit (Generalized Total Entropy Fit Index = -77.23), with all items showing high cluster stability (> 90\%), with the exception of *Olfa*. The final structure included 11 lower-level clusters, grouped into 3 higher-order metaclusters: "Interoceptive Deficit" (containing *CaCo*, *UrIn*, *Derm*, *Sati*, *Olfa*), "Interoceptive Awareness" (containing *ExAc*, *RelA*, *SexS*), and "Visceroception" (containing *Card*, *Resp*, *Gast*). Item centrality (intepretable as loadings) are shown in @fig-three. Additionally, we also ran a hierarchical clustering analysis which replicated this structure (see @fig-two).

```{r}
#| label: fig-three
#| fig-cap: "TODO."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../study2/analysis/figures/table1.png")
```

The correlation matrix of the Mint dimensions (i.e., clusters)

```{r}
#| label: fig-four
#| fig-cap: "TODO."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../study2/analysis/figures/fig2.png")
```

```{r}
#| label: fig-five
#| fig-cap: "Stars indicate dimensions that have been score-reversed for better in-context interpretation." 
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../study2/analysis/figures/fig3.png")
```

```{r}
#| label: fig-six
#| fig-cap: "TODO."
#| fig-align: "center"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../study2/analysis/figures/table2.png")
```

## Discussion

# General Discussion

# Data Availability

Data, code, and all materials are available at https://github.com/RealityBending/InteroceptionScale.

# Acknowledgements

We would like to thank the dissertation students from the University of Sussex for their help in data collection.
DM would also like to thank ... for the motivation provided to write this paper.

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::