### 2025-07-02 paper/manuscript.tex

\documentclass[
  jou,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\usepackage{subcaption}

\renewcommand\thesubfigure{\Alph{subfigure}}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{\textbf{The Mint Scale: A Fresh Validation of the Multimodal
Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS}}

\shorttitle{Mint Validation}

\usepackage{etoolbox}

\authorsnames[{1,2},{1},{1},{1},{1}]{Dominique Makowski,Ana Neves,Emma
Benn,Maisie Bennett,Giulia Poreiro}

\authorsaffiliations{
{School of Psychology, University of Sussex},{Sussex Centre for
Consciousness Science, University of Sussex}}

\leftheader{Makowski, Neves, Benn, Bennett and Poreiro}

\abstract{TO DO. }

\keywords{Interoception questionnaire, interoceptive accuracy scale,
MAIA, Mint Validation, Body Awareness}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, opacityback=0, left=2mm, arc=.35mm, leftrule=.75mm, breakable, bottomrule=.15mm, rightrule=.15mm]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}

\end{tcolorbox}  Author roles were classified using the Contributor Role
Taxonomy (CRediT; https://credit.niso.org/) as follows:  Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; Ana Neves:   Investigation, Methodology, Project
administration, Supervision, Writing -- original draft, Writing --
review \& editing; Emma Benn:   Investigation; Maisie
Bennett:   Investigation; Giulia
Poreiro:   Investigation, Methodology, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique
Makowski, Email: \href{mailto:D.Makowski@sussex.ac.uk}{D.Makowski@sussex.ac.uk}}
}

\usepackage{pbalance}
% \usepackage{float}
\makeatletter
\let\oldtpt\ThreePartTable
\let\endoldtpt\endThreePartTable
\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.485\textwidth}
\oldtpt[#1]
}
\def\ThreePartTable@ii{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.48\textwidth}
\oldtpt
}
\def\endThreePartTable{
\endoldtpt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\makeatletter
\let\endoldlt\endlongtable		
\def\endlongtable{
\hline
\endoldlt}
\makeatother

\newenvironment{twocolumntable}% environment name
{% begin code
\begin{table*}[!htbp]%
\onecolumn%
}%
{%
\twocolumn%
\end{table*}%
}% end code

\urlstyle{same}

\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

Main issues in existing questionnaires:

\begin{itemize}
\tightlist
\item
  Either heavily based on theories (e.g., focusing on specific
  ``alleged'' dimensions), despite shaky evidence for said-theories
\item
  Do not control for context (which leads to variability in
  interpretation and occurence)
\item
  Often quite narrow in the modalities covered (or scattered, like the
  IAS)
\end{itemize}

\section{Study 1: Item Selection}\label{study-1-item-selection}

TODO: write intro.

\begin{itemize}
\tightlist
\item
  Goal of study 1: to generate a lot of items, analyze its structure and
  reduce them to a balanced set of items.
\end{itemize}

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We recruited 760 English-speaking participants using
Prolific\textcopyright. We excluded 191 for failing at least one
attention check, and 10 based on measures significantly related to the
probability of failing attention checks (namely, the multivariate
distance obtained with the OPTICS algorithm,
\citeproc{ref-theriault2024check}{Thériault et al., 2024}). The final
sample includes 559 participants (age = 37.0 \(\pm\) 12.2 {[}18, 77{]};
50.8\% women; country of residence: 63.86\% UK, 26.65\% USA). This study
was approved by the University of Sussex' Ethics Committee
(ER/MB2021/1).

\subsubsection{Item Generation}\label{item-generation}

Based on the two goals outlined for this scale, namely to include
different interoceptive modalities, and to explicitly state the context
of the interoceptive experience (e.g., whether negative or positive), we
generated items in a systematic way following a combinatory approach,
where each item's category corresponds to the union of a specific
modality and context (Figure~\ref{fig-one}).

\begin{figure*}[!htbp]

{\caption{{The conceptual grid used to generate the 120 initial items
(top-left). Each item belongs both to an interoceptive modality
(vertical) and a facet (horizontal), with the number of each item per
category indicated in the circles. The asterisk denotes the additional
presence of an attention check item in that category. In the experiment,
these items were presented on different pages grouped either by modality
(bottom-left), by facet (top-right), or randomly. The Correlation
Similarity (bottom-right) analysis suggested that the correlation matrix
obtained from the participants assigned to the random-grouping condition
was slightly more similar (but non-significantly) to the one obtained in
the modality-grouping condition, suggesting that the scale's structure
is robust to different presentation conditions, and that the
modality-grouping might potentially tend to facilitate the emergence of
the underlying item structure (and thus be interpreted as being more
``natural'').}{\label{fig-one}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study1/analysis/figures/figure1.png}
\end{center}

\end{figure*}

We identified 7 ``modalities'' (cardiac, respiratory, gastric, genital,
skin \& temperature (thermoregulation), bladder \& colon, and a
``general state'' category corresponding to a holistic and general
awareness of an interoceptive state or dimension). Through iterative
refinement (e.g., splitting or merging different categories together),
we settled on 6 ``facets'', which encompass both \emph{contexts} of
experience (negative and positive arousal, e.g., anxious and sexual
states), and potential distinct \emph{mechanisms} (nociception \&
pleasure, sensitivity, accuracy, and confusion).

Using this orthogonal 7x6 modality/facet grid as a conceptual
scaffolding, we generated 120 initial items, striving for a balanced
number of items with consistent phrasing within modalities and
facets\footnote{The initial item list is available at
  \href{https://realitybending.github.io/InteroceptionScale/study1/analysis/2_analysis.html}{realitybending.github.io/InteroceptionScale/study1/analysis/2\_analysis.html}}.
We additionally crafted 8 ``attention check'' items blending in (and
distributed across) each category.

\subsubsection{Procedure}\label{procedure}

To avoid presenting all the 120 items on a single long and discouraging
page, we split them into different pages. Participants were randomly
assigned to one of three conditions, driving how items were grouped on
the same page: 1) items grouped by modality (i.e., all cardiac items on
the first page, all colon \& bladder items on the second, etc.), 2)
items grouped by facet, or 3) items presented fully randomly (but with
their number balanced across 6 pages). The order of the item on any
given page and the order of the modalities/facets was randomized. Each
participant completed the full set of 120 items, with 8 attention check
items interspersed throughout. The online experiment was implemented
using JsPsych (\citeproc{ref-de2015jspsych}{De Leeuw, 2015}), and item
responses were recorded using 7-points Likert scales (0 = Disagree, 6 =
Agree).

\subsubsection{Data Analysis}\label{data-analysis}

In order to test whether the grouping condition had an effect on the
structure (i.e., how items relate to one-another), we compared the
correlation matrix obtained in the random condition to the ones obtained
in the modality and facet conditions, focusing on 3 indices of
correlation matrix similarity - the Procrustes Similarity Index (PSI,
\citeproc{ref-sibson1978studies}{Sibson, 1978}), the Adjusted RV (Rvadj,
\citeproc{ref-mayer2011exploratory}{Mayer et al., 2011}), and the
Similarity of Matrices Index (SMI,
\citeproc{ref-indahl2018similarity}{Indahl et al., 2018}). For each
index, we bootstrapped the difference between the similarity with the
facet and modality conditions to test whether the correlation matrix in
the random-grouping condition is significantly more similar to any of
the two other conditions.

Items deemed ``redundant'' (which can distort the item structure
estimation by introducing multicollinearity or local dependencies) were
identified (using the recommended threshold of 0.25) using Unique
Variable Analysis (UVA, \citeproc{ref-christensen2023unique}{Christensen
et al., 2023}), a novel and principled method derived from network
psychometrics.

The structure of the items was analyzed using the recently-developed
Exploratory Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{H.
F. Golino \& Epskamp, 2017}) framework, which allows to jointly estimate
the number of dimensions (i.e., clusters of items), the structure, as
well as its stability using bootstrapping
(\citeproc{ref-golino2020investigating}{H. Golino et al., 2020}). At a
fundamental level, EGA conceptualizes variables as nodes in a network,
with connections (edges) reflecting associations between them. Evidence
has underlined its suitability as an alternative to traditional factor
analysis, addressing some of its limitations such as the assumption of a
``latent'' source of variability, issues in estimation of the optimal
factor numbers, and poor performance in complex population structures,
while remaining comparable and interpretable
(\citeproc{ref-christensen2021equivalency}{Christensen \& Golino, 2021};
\citeproc{ref-jimenez2023dimensionality}{Jiménez et al., 2023}). In
particular, nodes communities (i.e., clusters of items) can be in
practice interpreted as distinct ``dimensions'', similarly to
traditional latent factors - but without explicitly assuming their
existence (\citeproc{ref-christensen2021equivalency}{Christensen \&
Golino, 2021}).

After removing redundant items using UVA, we iteratively fitted
hierarchical EGA models (which additionally estimates higher-order
``meta'' clusters) using ``glasso''
(\citeproc{ref-friedman2008sparse}{Friedman et al., 2008}) and the
``Leiden'' algorithm (\citeproc{ref-traag2019louvain}{Traag et al.,
2019}) for community detection, refining the item pool at each step. We
started by removing items with a low (\textless{} 80\%) cluster
stability (i.e., volatile items which jump between clusters across
bootstrapped samples), followed by odd items belonging to no clusters or
pairs of items (i.e., we keep items belonging to clusters of more than 2
items). Finally, for each lower-level cluster, we selected the 3 items
with the highest node centrality (i.e., the highest loading in the
cluster).

\subsection{Results}\label{results}

The correlation matrix similarity analysis yielded no significant
differences between the similarity of the random-grouping condition with
the modality-grouping and facet-grouping conditions
(\(PSI_{\text{Random vs. Facet}} = 0.81\),
\(PSI_{\text{Random vs. Modality}} = 0.82\), \(p = .45\);
\(RVadj_{\text{Random vs. Facet}} = 0.77\),
\(RVadj_{\text{Random vs. Modality}} = 0.78\), \(p = .74\);
\(SMI_{\text{Random vs. Facet}} = 0.49\),
\(SMI_{\text{Random vs. Modality}} = 0.51\), \(p = .52\)), despite a
consistent bias in favour of the modality-grouping condition.

From the 120 initial items, 4 redundant items were flagged by UVA. We
then removed 40 items that showed low cluster stability, and 9 items
that were part of clusters with less than 3 items. Finally, We kept the
3 items with the highest loading in their lower-level structure
(removing 13 items in the process), resulting in 54 items in the final
item pool.

The final hierarchical EGA model (Generalized Total Entropy Fit Index =
-119.18) - in which all 54 items yielded a high cluster stability
(\textgreater{} 90\%) suggested 3 metaclusters and 15 lower-level
clusters (each containing 3 items): ``Interoceptive Deficits''
(containing 5 clusters: \emph{Urointestinal Inaccuracy} - UrIn;
\emph{Cardiorespiratory Confusion} - CaCo; \emph{Cardiorespiratory
Noticing} - CaNo; \emph{Olfactory Compensation} - Olfa; \emph{Satiety
Noticing} - Sati), ``Interoceptive Awareness'' (containing 7 clusters:
\emph{Sexual Arousal Awareness} - SexA; \emph{Sexual Arousal
Sensitivity} - SexS; \emph{Sexual Organs Sensitivity} - SexO;
\emph{Urointestinal Sensitivity} - UrSe; \emph{Relaxation Awareness} -
RelA; \emph{State Specificity} - StaS; \emph{Expulsion Accuracy} -
ExAc), and ``Interoceptive Sensitivity'' (containing 6 clusters:
\emph{Cardioception} - Card; \emph{Respiroception} - Resp;
\emph{Signalling} - Sign; \emph{Gastroception} - Gast; \emph{Dermal
Hypersensitivity} - Derm; \emph{Sexual Arousal Changes} - SexC).

To further reduce and balance the remaining items, we collectively
decided on removing the lower-level clusters \emph{Sexual Arousal
Awareness} (too general and overlapping with the other more specific
sex-related items), \emph{Signalling} (which items started with ``when
something important is happening in my life'', which meaning we deemed
too much open to interpretation), and \emph{Sexual Arousal Changes} (low
fit with the other modality-focused clusters of its group). The final
set included 45 items.

\subsection{Discussion}\label{discussion}

\begin{itemize}
\tightlist
\item
  Possible (although not at all significant) bias consistent across
  indices in favour of a greater similarity with the modality-grouping
  condition. Futher research with more data (here we only had
  \textasciitilde200 observation per condition) might be able to yield
  more conclusive results. In general, testing the presentation mode is
  a novel approach which should be more widely applied to psychometric
  validations.
\item
  We grouped some of the dimensions under ``Awareness'', despite being
  similar to items included in the IAS typically conceptualized as
  ``Accuracy''. However, we believe that subjective scales cannot be
  pure measures of accuracy, as they are inherently subjective and thus
  reflect a degree of awareness (of one's accurate noticing or
  sensation).
\item
  Olfactory compensation: pseudo-interoceptive pathway whereby people
  with interoceptive deficits would compensate by gathering evidence
  about their internal state externally (e.g., via smell).
\item
  The metacluster ``Sensitivity'' seems to be grouping iterms related to
  visceropcetion and (hyper)sensitivity (Derm).
\end{itemize}

\section{Study 2: Validation}\label{study-2-validation}

\begin{itemize}
\tightlist
\item
  Goal of study 2: to validate the Mint scale against other
  interoception scales.
\end{itemize}

\subsection{Methods}\label{methods-1}

\subsubsection{Participants}\label{participants-1}

We recruited 921 English-speaking participants via SONA and
Prolific\textcopyright, from which 118 were excluded for failing at
least one attention check and 6 based on multivariate distance (using
the same procedure as for study 1). 60 participants were further
excluded due to missing data following a technical issue. The final
sample includes 737 participants (age = 36.8 \(\pm\) 14.7 {[}18, 87{]};
57.3\% women; Country of residence: 75.17\% United Kingdom, 24.83\%
other). This study was approved by the University of Sussex' Ethics
Committee (ER/EB672/2).

\subsubsection{Measures}\label{measures}

\begin{figure*}[!htbp]

{\caption{{Structure analysis of the final set of 33 Mint items.
Hierarchical Exploratory Graph Analysis (EGA) was applied to jointly
identify clusters of items, as well as higher-order metaclusters. The
reliability of each item can be quantified by the proportion of
structure replication across bootstrapped samples, with high values
indicative of high structural stability. The Mint scale displayed an
excellent structural consistency, with the exception of the \emph{Olfa}
cluster which belonging to the \emph{Deficit} metaclusters warrants
further research. Bottom plot shows the result of hierarchical
clustering, providing evidence for structure statbility across methods,
and allowing for more granular understanding of the relationships
between variables.}{\label{fig-two}}}}

\begin{center}
\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig1.png}
\end{center}

\end{figure*}

\paragraph{Interoception
Questionnaires.}\label{interoception-questionnaires}

The 45 items of the Multimodal Interoception Questionnaire (Mint) scale
were presented in a random order, with the same 7-point Likert scale as
in study 1 (0 = \emph{Disagree}, 6 = \emph{Agree}).

The Multidimensional Assessment of Interoceptive Awareness (MAIA-2,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018})
measures 8 dimensions with 37 items presented on a 7-point Likert scale
(0 = \emph{Never}, 6 = \emph{Always}). It includes bodily dimensions
such as Body Listening, Noticing, Emotional Awareness, dimensions
related to emotion regulation, such as Trust, Not-worrying, and
Self-Regulation, as well as dimensions related to attention, such as
Attention Regulation and Not-Distracting. The relationship of some of
these dimensions to interoception remains debated, particularly
Not-Distracting and Not-Worrying, which tend to show weak or
non-existent correlations with the other MAIA subscales
(\citeproc{ref-ferentzi2021examining}{Ferentzi et al., 2021}).

The Body Perception Questionnaire - Very Short Form (BPQ-VSF,
\citeproc{ref-cabrera2018assessing}{Cabrera et al., 2018}) measures a
general score of bodily awareness with 12 items presented on 5-point
Likert scale (0 = \emph{Never}, 5 = \emph{Always}).

The Interoceptive Accuracy Scale (IAS,
\citeproc{ref-murphy2020testing}{Murphy et al., 2020}) measures a
general score of interoceptive accuracy to physical sensations with 21
items presented on a 5-point Likert scale (1 = \emph{Disagree Strongly},
5 = \emph{Strongly Agree}).

\paragraph{Emotions and Cognition.}\label{emotions-and-cognition}

The Toronto Alexithymia Scale TAS (TAS-20,
\citeproc{ref-leising2009toronto}{Leising et al., 2009}) measures 3
Alexithymia dimensions with 20 items 5-point Likert scale (1 =
\emph{Strongly Disagree}, 5 = \emph{Strongly Agree}: Difficulty
Identifying Feelings (DIF), Difficulty Describing Feelings (DDF), and
Externally Oriented Thinking (EOR).

The Emotion Reactivity Scale - Brief Version (B-ERS,
\citeproc{ref-veilleux2024development}{Veilleux et al., 2024}) measures
3 dimensions with 6 items on a 5-point Likert scale (0 = \emph{Not like
me at all}, 4 = \emph{Extremely like me}): Arousal (\emph{``My moods are
very strong and powerful''}), Sensitivity (\emph{``I tend to get very
emotional very easily''}), and Persistence (\emph{``When I am
angry/upset, it takes me much longer than most people to calm down''}).

The Cognitive Emotion Regulation Questionnaire (CERQ-short,
\citeproc{ref-garnefski2006cognitive}{Garnefski \& Kraaij, 2006})
measures 9 adaptive and maladaptive emotion regulation strategies with
18 items (1 = \emph{Almost Never}, 5 = \emph{Almost Always}). The
strategies include Self-blame, Other-blame, Rumination, Catastrophizing,
Putting into Perspective, Positive Refocusing, Positive Reappraisal,
Acceptance, and Planning.

The Černis Felt Sense of Anomaly scale - short form (CEFSA-14,
\citeproc{ref-vcernis2024measuring}{Černis et al., 2024}) measures 7
dissociative experiences with 14 items on a 5-point Likert scale (0 =
\emph{Never}, 4 = \emph{Always}). This scale is designed to measure
dissociative experiences in adolescence and adulthood, specifically
focusing on the felt sense of anomaly-type dissociation. It includes
Anomalous Experience of the Self, Anomalous Experience of the Body,
Anomalous Experiences of Emotion, Altered Sense of Familiarity, Altered
Sense of Connection, Altered Sense of Agency, and Altered Sense of
Reality.

The Primal World Beliefs Inventory (PI-18,
\citeproc{ref-clifton2021brief}{Clifton \& Yaden, 2021}) measures
higher-order beliefs about the basic character of the world: Good, Safe,
Enticing, and Alive. The 18 items are presented on a 6-point Likert
scale (0 = \emph{Strongly Disagree}, 5 = \emph{Strongly Agree}).

\paragraph{Health and Wellbeing.}\label{health-and-wellbeing}

The Single Item Life Satisfaction scale (SILS,
\citeproc{ref-jovanovic2020longer}{Jovanović \& Lazić, 2020}) was
followed by the PHQ-4 (\citeproc{ref-kroenke2009ultra}{Kroenke et al.,
2009}) assessing depression and anxiety symptoms with 4 items. We used
the refined version of the PHQ-4
(\citeproc{ref-makowski2025measuring}{Makowski et al., 2025}), which
adds an additional response option (``Once or twice''), increasing
sensitivity to subclinical fluctuations.

Participants were asked to self-report any current, medically diagnosed
psychiatric disorders using a checklist based on DSM-5 categories. If
one or more conditions were endorsed, participants were asked to
indicate any current treatments, including pharmacological (e.g.,
antidepressants, anxiolytics, antipsychotics, mood stabilizers),
psychological (e.g., psychotherapy, mindfulness), or lifestyle
interventions. Binary variables (0 = absent, 1 = present) were created
to identify participants reporting mood disorders (MDD, GAD, Bipolar
Disorder; with a stricter subgroup of participants undergoing a pharma-
or psychological treatment), anxiety-centred disorders (GAD, Panic
Disorder, Social Anxiety Disorder, Specific Phobias), eating disorders,
addiction-related disorders, borderline personality disorder, autism
spectrum disorder (ASD), and ADHD.

Participants were asked to select somatic symptoms or conditions they
experienced from a list of 36 options. To facilitate mechanistic
interpretation and reduce redundancy, answers were grouped into four
non-overlapping clusters based on shared physiological pathways and
known etiological mechanisms. The \emph{Afferent Sensitivity} cluster
included conditions associated with heightened interoceptive awareness
and neurogenic excitability, such as migraine, neuropathy, dizziness,
nausea, muscle tension, epilepsy, and frequent urination. The
\emph{Central Sensitization} cluster comprised syndromes characterized
by chronic pain and fatigue, likely reflecting central amplification of
sensory signals and HPA-axis dysregulation, such as fibromyalgia,
chronic fatigue syndrome, chronic pain, back pain, pelvic pain,
irritable bowel syndrome (IBS). The \emph{Autonomic Dysfunction} cluster
captured disorders linked to dysregulation of the autonomic and
cardiopulmonary systems, including joint hypermobility, cardiac
arrhythmia, chest pain, shortness of breath, hypo-/hypertension, sleep
apnea, chronic obstructive pulmonary disease (COPD), and chronic
bronchitis. Finally, the \emph{Immune-Inflammatory} cluster encompassed
conditions associated with immune dysregulation, barrier dysfunction,
and gut-brain axis disturbance, such as eczema, psoriasis, skin rashes,
asthma, celiac disease, gluten and lactose sensitivity, inflammatory
bowel diseases (Crohn's disease and ulcerative colitis),
gastroesophageal reflux disease (GERD), multiple sclerosis, and
Sjögren's syndrome. We scored each cluster as a binary variable based on
whether the participant selected at least one symptom from that cluster.

\paragraph{Lifestyle.}\label{lifestyle}

Participants reported about owning any wearable devices to monitor
health indices such as heart rate, number of steps, calories burned,
sleep quality, and weight. For each selected device, a follow-up
question inquired about the frequency of usage and subjective importance
of that measure.

Participants were asked to rate how physically active they consider
themselves and how many hours of active workout they engage in per week.
Participant's BMI was computed using height and weight.

\subsubsection{Procedure}\label{procedure-1}

In order to avoid the repetition of similar types of questions and and
balance longer and shorter questionnaires, we partitioned the measures
into three groups (and randomized the order within them): 1)
interoception questionnaires (MAIA, IAS, BPQ), 2) emotions (TAS, CERQ,
ERS, PI-18), and 3) health (somatic symptoms, mental health, LS + PHQ-4,
CEFSA). After completing demographic questions, participants always
started with the Mint scale, and each following interoception
questionnaire was interspersed with two questionnaires from the emotions
and pathology groups. 7 attention checks were embedded throughout (one
within each major questionnaire). In order to make the experiment more
enjoyable, the experiment ended with a radar chart summarizing the
participants' responses to the Mint scale\footnote{The experiment can be
  tested by following the link on
  https://github.com/RealityBending/InteroceptionScale}.

\begin{figure*}[!htbp]

{\caption{{Item loadings. The table shows each item of the Mint with its
cluster centrality, which is the EGA equivalent (in terms of
interpretation) of factor loadings. It also shows how each lower-level
cluster is related to higher-order metaclusters.}{\label{fig-three}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/table1.png}
\end{center}

\end{figure*}

\subsubsection{Data Analysis}\label{data-analysis-1}

We started by confirming and further refining the structure of the Mint
scale (see Figure~\ref{fig-two}) using the same EGA model as in Study 1.
We then computed the lower-level dimensions and the higher-level
metaclusters' scores by averaging their corresponding items. The
convergent validity of the final set of items was assessed by computing
the correlations between the Mint scale and the other interoception
questionnaires (MAIA, BPQ, IAS).

Next, we tested the predictive power of the Mint scale relative to other
interoception questionnaires. We will assessed the importance of each
interoceptive dimension (from all the scales) as a unique predictor by
fitting regression models (linear for continuous measures - e.g.,
depression score - and logistic for binary variables - e.g., presence
\emph{vs.} absence of mood disorders) to predict different outcomes, and
compare the best between the 4 interoception scales (based on the R2).
We then evaluated the predictive performance of each scale as a whole by
comparing regression models with all the dimensions entered as
predictors (note that the IAS and BPQ only have one total score
variable). We assessed the models based on their total explained
variance (R2), as well as on the Bayesian Information Criterion (BIC),
which penalizes for predictor number, thus offering a balance between
model parsimony and predictive power. For the logistic models, we
quantified the discriminative power by computing the Area Under the
Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which
assesses the model's discriminative power (the combination of
sensitivity and specificity).

In order to evaluate the potential for a short version of our scale, we
compared 4 variants of the Mint: the (full) \emph{Mint} (including all
lower-level dimensions), the \emph{metaMint} (only including the
metaclusters), the \emph{miniMint} (including the metaclusters computed
from a reduced number of items), and the \emph{microMint} (including
only the most representative dimension of each metacluster). Moreover,
we also included an alternative ``interoception-focused'' version of the
MAIA (\emph{iMAIA}) that only contains the 3 most interoceptive
dimensions (Body Listening, Noticing, and Emotional
Awareness)\footnote{See correlation results to further justify this
  selection.}.

\subsection{Results}\label{results-1}

The application of EGA to the initial set of 42 items reproduced the
expected 14 lower-level clusters of triplets and 3 higher-order
metaclusters, with the exception of the \emph{UrSe} (Urointestinal
Sensitivity), for which one item moved to the \emph{Olfa} (Olfactory
Compensation) cluster. In order to further balance and reduce the items,
we removed the \emph{UrSe} cluster, as well as \emph{StaS} (State
Specificity), which in comparison to the other items stood out as
containing vague and overlapping items ; \emph{SexO} (Sexual Organs
Sensitivity), which overlapped with the \emph{SexS} (Sexual Arousal
Sensitivity) cluster; and \emph{CaNo} (Cardiorespiratory Noticing),
which overlapped with the \emph{CaCo} (Cardiorespiratory Confusion)
cluster.

The final set of 33 items yielded a good fit (Generalized Total Entropy
Fit Index = -77.23), with all items showing high cluster stability
(\textgreater{} 90\%), with the exception of \emph{Olfa}. The final
structure included 11 lower-level clusters, grouped into 3 higher-order
metaclusters: ``Interoceptive Deficit'' (containing \emph{CaCo},
\emph{UrIn}, \emph{Derm}, \emph{Sati}, \emph{Olfa}), ``Interoceptive
Awareness'' (containing \emph{ExAc}, \emph{RelA}, \emph{SexS}), and
``Visceroception'' (containing \emph{Card}, \emph{Resp}, \emph{Gast}).
Item centrality (intepretable as loadings in traditional factor analysis
frameworks) are shown in Figure~\ref{fig-three}. Additionally, we
applied hierarchical clustering analysis which replicated this structure
(see Figure~\ref{fig-two}), suggesting consistency across methods.

\begin{figure*}[!htbp]

{\caption{{Correlation Matrices between the Mint dimensions
(upper-left), and between the Mint and other interoception
questionnaires (MAIA, BPQ and IAS; upper-right). The bottom matrix shows
the relationships between interoception and other measures included in
the study, such as alexithymia (TAS-20), depression and anxiety (PHQ-4),
emotion reactivity (ERS), abnormal and dissociative experiences (CEFSA),
and Primal World Beliefs. Stars indicate dimensions that have been
score-reversed for better in-context interpretation. Correlation
coefficients are shown for significant
correlations.}{\label{fig-four}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig2.png}
\end{center}

\end{figure*}

The correlation matrix of the Mint dimensions revealed an interesting
and rich tapestry of relationships (see Figure~\ref{fig-four}), with
contrasting patterns of associations with dimensions from the same
group. For instance, \emph{Derm} and \emph{Olfa}, despite being
positively correlated with the other dimensions from the \emph{Deficit}
cluster, were not negatively correlated to \emph{Awareness} and its
dimensions. Similarly, \emph{Sati}, unlike the other \emph{Deficit}
dimensions, was not positively correlated to \emph{Visceroception}. This
complex structure suggests that the dimensions indeed capture distinct
aspects of interoception, and that the metaclusters, rather than being
simple aggregates of rather-redundant elements, might actually capture
unique combinations (greater than the sum of their parts). It also
provides evidence against a simplistic adaptive \emph{vs.} maladaptive
dichotomy, as \emph{Deficit} was not negatively bit positively
correlated with \emph{Visceroception}. Interestingly, while
\emph{Awareness} was also positively correlated with
\emph{Visceroception}, it yielded an insignificant negative correlation
with \emph{Deficit} (this finding was also aligned with the hierarchical
clustering results, showing a greater proximity between \emph{Deficit}
and \emph{Visceroception} than with \emph{Awareness}), underlining again
the complex web of relationships captured by the Mint.

The correlation matrix between the Mint and the other interoception
scales revealed high levels of overlap, as well as some unique
contributions (see Figure~\ref{fig-four}). The BPQ was positively
correlated with most Mint dimensions (the highest with
\emph{Visceroception}, \(r = .46\)). The IAS was positively correlated
with \emph{Visceroception} and \emph{Awareness} dimensions (the highest
with \emph{Awareness}, \(r = .63\)), but negatively with most
\emph{Deficit} dimensions (with the exception of \emph{Olfa} and
\emph{Derm}). In turn, the \emph{Visceroception} metacluster most
strongly correlated with MAIA's Noticing (\(r = .55\)). Interestingly,
MAIA's TTrusting correlated selectively with \emph{Awareness}, and
negatively with \emph{Deficit} dimensions, but not with
\emph{Visceroceptive} dimensions (underlining its high-level
metacognitive nature). MAIA's Emotional Awareness and Body Listening
displayed a similar pattern to Noticing, and Attention Regulation and
Self Regulation positively correlated with \emph{Awareness} and
\emph{Visceroception} dimensions, but negatively with some
\emph{Deficit} dimensions (\emph{CaCo} and \emph{UrIn}). Not-distracting
only yielded mild negative associations with \emph{Sati} and
\emph{Olfa}. Overall, the Mint dimensions were able to capture most of
the (relevant) variance and intricacies present in the other
interoception scales.

Exploratory correlations with the emotion regulation (CERQ) dimensions
revealed stronger associations with most of the MAIA dimensions
(supporting its proximity with emotion regulation). Interestingly,
Rumination (and Self-Blame) stood out as selectively related to the
Mint's \emph{Visceroception} and \emph{Deficit} dimensions (note that
Rumination was also related to the MAIA's interoceptive dimensions,
namely Noticing, Emotional Awareness and Body Listening). The only
primal world belief that correlated particularly with the Mint's
\emph{Visceroception} and the MAIA's interoceptive dimension was the
belief that the world was alive (\textbf{TODO: Ana maybe throw in a
couple more words about what this belief means and is defined}).

\begin{figure*}[!htbp]

{\caption{{Summary table of the comparison between interoception
questionnaires (Mint, MAIA, BPQ and IAS). For various outcomes included
the study, we tested the Mint and Non-Mint dimension that had the
strongest link (as a unique predictor). Values in parenthesis represents
correlation coefficient for continuous variables and log-odds ratios for
binary variables (the occurrence of mental and somatic health
conditions). We also compared the predictive performance of regression
models including multiple predictors, and present their ranking based on
raw predictive power (R2) and BIC, which favours more parsimonious
models (with less predictors). Green background indicates an advantage
for the Mint, while red backgrounds indicate an advantage for another
interoception scale. Orange backgrounds indicate an advantage for the
MAIA driven by its emotion-regulation dimensions (Trusting and
Not-Worrying).}{\label{fig-five}}}}

\begin{center}
\includegraphics[width=0.82\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/table2.png}
\end{center}

\end{figure*}

Most of the target outcomes measured in the study to assess validity
were best predicted by one of the Mint dimension (see
Figure~\ref{fig-five}). This included Alexithymia (best predicted by
\emph{Deficit}); ERS' Arousal (best predicted by \emph{Deficit});
CEFSA's anomalous experiences of the body (best predicted by
\emph{CaCo}), Self and Reality (best predicted by \emph{Deficit}); ADHD
and Autism (best predicted by \emph{Deficit}), Somatic symptoms (best
predicted by \emph{Deficit}); BMI (best predicted by \emph{Sati}). Most
of the exceptions showed an advantage for MAIA dimensions, in particular
Not-worrying (which best predicted ERS' Sensitivity and Persistence) and
Trusting (which best predicted LS, Depression and Anxiety, CEFSA's
Emotion and Connection, and self-reported physical activity).

Model comparison included 4 variants of the Mint scale, the full
\emph{Mint} (including all lower-level clusters - 33 items), the
\emph{metaMint} (including only the 3 metaclusters based on all items),
the \emph{miniMint} (including the 3 metaclusters computed from 2 of its
most loading triplets - 18 items), and the \emph{microMint} (including
only the most representative dimension of each metacluster - 9 items).
This analysis confirmed the clear advantage for the Mint over the other
interoception scales. The full \emph{Mint} model was typically the best
model based on R2, and the \emph{metaMint} was the best model when
parsimony was taken into account (i.e., based on BIC). Moreover, most of
the instances were the \emph{MAIA} was the best model were explainable
by the inclusion of Trusting, and the interoception-only version
\emph{iMAIA} typically yielded worse performance than the Mint. In many
instances, the \emph{miniMint} yielded reasonable performance, although
the \emph{microMint} was typically less promising.

The predictive performance for the mental health outcomes (see
Figure~\ref{fig-six}) displayed a consistent pattern, with an advantage
for the \emph{MAIA} for depression and anxiety which dropped below the
\emph{Mint} versions for its \emph{iMAIA} variant. The \emph{Mint},
however, displayed a clear advantage for the prediction of autism, ADHD,
and all somatic groups of symptoms (with the exception of Central
Sensitization, which was best predicted by the \emph{MAIA}).

Finally, the importance of heart monitoring (for owners of such
wearable) was best predicted by MAIA's Body listening (\(r = .36\)),
followed the \emph{Card} dimension of the Mint (\(r = .31\)). The
importance of sleep monitoring importance was best predicted by
\emph{RelA} (\(r = .22\)), followed by MAIA's not worrying
(\(r = -.219\)). Daily activity via steps monitoring was best predicted
by MAIA's Noticing (\(r = .28\)), followed by \emph{Awareness}
(\(r = .26\)).

\begin{figure*}[!htbp]

{\caption{{Predictive performance of interoception questionnaires for
various self-reported mental health conditions (mood disorders with
psychopharmacological treatment, ADHD and ASD) and somatic symptoms. The
Receiver Operating Characteristic (ROC) curves are shown for the
logistic regression models for various questionnaires and versions. A
high Area Under the Curve (AUC) indicates a good discriminative power of
the model (i.e., a strong combination of sensitivity and specificity).
The Mint scale (in blue-green) consistently outperformed the other
interoception scales, with the exception of the MAIA which, driven by
its emotion-regulation dimensions (Trusting and Not-worrying), performed
better for mood disorders and central sensitization symptoms. In several
instances, the short versions of the Mint scale (miniMint and microMint)
yielded reasonable performance, still outperforming other
measures.}{\label{fig-six}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig3.png}
\end{center}

\end{figure*}

\subsection{Discussion}\label{discussion-1}

\begin{itemize}
\tightlist
\item
  Structure: Low stability of \emph{Olfa} as belonging in the deficit
  group. More studies are needed to investigate its place and cluster
  invariance: it is possible that its association might depend on other
  factors or categories.
\item
  Main contender is MAIA. But its most sensitive dimensions seem to be
  the least related to interoception, which is dangerous as it might
  lead to findings with associations misattributed to ``Interoception''
  (just because it correlate with some MAIA dimensions), but it might
  have nothing to do with interoception per se and more with emotion
  regulation.
\item
  Basically the pattern is clear: everything high-level, metacognitive
  and highly subjective related to wellbeing and general affective state
  is the most strongly related to the MAIA through some of its Emotion
  regulation dimensions. All the other outcomes actually related to
  interoceptive processes are most strongly predicted by the Mint.
\end{itemize}

\section{General Discussion}\label{general-discussion}

\begin{itemize}
\tightlist
\item
  Future directions: needs to test relationship with physiological
  measures of interoception.
\item
  Also, the structure might still require a bit of deeper investigation,
  in particular with some of its facets (\emph{Olfa} and \emph{Derm})
  which might display different associations depending on the context or
  the population.
\end{itemize}

\section{Data Availability}\label{data-availability}

Data, code, and all materials are available at
https://github.com/RealityBending/InteroceptionScale.

\section{Acknowledgements}\label{acknowledgements}

We would like to thank the dissertation students from the University of
Sussex for their help in data collection. DM would also like to thank
\ldots{} for the motivation provided to write this paper.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-cabrera2018assessing}
Cabrera, A., Kolacz, J., Pailhez, G., Bulbena-Cabre, A., Bulbena, A., \&
Porges, S. W. (2018). Assessing body awareness and autonomic reactivity:
Factor structure and psychometric properties of the body perception
questionnaire-short form (BPQ-SF). \emph{International Journal of
Methods in Psychiatric Research}, \emph{27}(2), e1596.

\bibitem[\citeproctext]{ref-vcernis2024measuring}
Černis, E., Loe, B. S., Lofthouse, K., Waite, P., Molodynski, A.,
Ehlers, A., \& Freeman, D. (2024). Measuring dissociation across
adolescence and adulthood: Developing the short-form {Č}ernis felt sense
of anomaly scale ({Č}EFSA-14). \emph{Behavioural and Cognitive
Psychotherapy}, \emph{52}(2), 163--177.

\bibitem[\citeproctext]{ref-christensen2023unique}
Christensen, A. P., Garrido, L. E., \& Golino, H. (2023). Unique
variable analysis: A network psychometrics method to detect local
dependence. \emph{Multivariate Behavioral Research}, \emph{58}(6),
1165--1182.

\bibitem[\citeproctext]{ref-christensen2021equivalency}
Christensen, A. P., \& Golino, H. (2021). On the equivalency of factor
and network loadings. \emph{Behavior Research Methods}, \emph{53}(4),
1563--1580.

\bibitem[\citeproctext]{ref-clifton2021brief}
Clifton, J. D., \& Yaden, D. B. (2021). Brief measures of the four
highest-order primal world beliefs. \emph{Psychological Assessment},
\emph{33}(12), 1267.

\bibitem[\citeproctext]{ref-de2015jspsych}
De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating
behavioral experiments in a web browser. \emph{Behavior Research
Methods}, \emph{47}, 1--12.

\bibitem[\citeproctext]{ref-ferentzi2021examining}
Ferentzi, E., Olaru, G., Geiger, M., Vig, L., Köteles, F., \& Wilhelm,
O. (2021). Examining the factor structure and validity of the
multidimensional assessment of interoceptive awareness. \emph{Journal of
Personality Assessment}, \emph{103}(5), 675--684.

\bibitem[\citeproctext]{ref-friedman2008sparse}
Friedman, J., Hastie, T., \& Tibshirani, R. (2008). Sparse inverse
covariance estimation with the graphical lasso. \emph{Biostatistics},
\emph{9}(3), 432--441.

\bibitem[\citeproctext]{ref-garnefski2006cognitive}
Garnefski, N., \& Kraaij, V. (2006). Cognitive emotion regulation
questionnaire--development of a short 18-item version (CERQ-short).
\emph{Personality and Individual Differences}, \emph{41}(6), 1045--1053.

\bibitem[\citeproctext]{ref-golino2017exploratory}
Golino, H. F., \& Epskamp, S. (2017). Exploratory graph analysis: A new
approach for estimating the number of dimensions in psychological
research. \emph{PloS One}, \emph{12}(6), e0174035.

\bibitem[\citeproctext]{ref-golino2020investigating}
Golino, H., Shi, D., Christensen, A. P., Garrido, L. E., Nieto, M. D.,
Sadana, R., Thiyagarajan, J. A., \& Martinez-Molina, A. (2020).
Investigating the performance of exploratory graph analysis and
traditional techniques to identify the number of latent factors: A
simulation and tutorial. \emph{Psychological Methods}, \emph{25}(3),
292.

\bibitem[\citeproctext]{ref-indahl2018similarity}
Indahl, U. G., Næs, T., \& Liland, K. H. (2018). A similarity index for
comparing coupled matrices. \emph{Journal of Chemometrics},
\emph{32}(10), e3049.

\bibitem[\citeproctext]{ref-jimenez2023dimensionality}
Jiménez, M., Abad, F. J., Garcia-Garzon, E., Golino, H., Christensen, A.
P., \& Garrido, L. E. (2023). Dimensionality assessment in bifactor
structures with multiple general factors: A network psychometrics
approach. \emph{Psychological Methods}.

\bibitem[\citeproctext]{ref-jovanovic2020longer}
Jovanović, V., \& Lazić, M. (2020). Is longer always better? A
comparison of the validity of single-item versus multiple-item measures
of life satisfaction. \emph{Applied Research in Quality of Life},
\emph{15}, 675--692.

\bibitem[\citeproctext]{ref-kroenke2009ultra}
Kroenke, K., Spitzer, R. L., Williams, J. B., \& Löwe, B. (2009). An
ultra-brief screening scale for anxiety and depression: The PHQ--4.
\emph{Psychosomatics}, \emph{50}(6), 613--621.

\bibitem[\citeproctext]{ref-leising2009toronto}
Leising, D., Grande, T., \& Faber, R. (2009). The toronto alexithymia
scale (TAS-20): A measure of general psychological distress.
\emph{Journal of Research in Personality}, \emph{43}(4), 707--710.

\bibitem[\citeproctext]{ref-makowski2025measuring}
Makowski, D., Te, A., Neves, A., \& Chen, S. A. (2025). \emph{Measuring
depression and anxiety with 4 items? Adaptation of the PHQ-4 to increase
its sensitivity to subclinical variability}. PsyArXiv.
\url{https://doi.org/10.31234/osf.io/436np_v2}

\bibitem[\citeproctext]{ref-mayer2011exploratory}
Mayer, C.-D., Lorent, J., \& Horgan, G. W. (2011). Exploratory analysis
of multiple omics datasets using the adjusted RV coefficient.
\emph{Statistical Applications in Genetics \& Molecular Biology},
\emph{10}(1).

\bibitem[\citeproctext]{ref-mehling2018multidimensional}
Mehling, W. E., Acree, M., Stewart, A., Silas, J., \& Jones, A. (2018).
The multidimensional assessment of interoceptive awareness, version 2
(MAIA-2). \emph{PloS One}, \emph{13}(12), e0208034.

\bibitem[\citeproctext]{ref-murphy2020testing}
Murphy, J., Brewer, R., Plans, D., Khalsa, S. S., Catmur, C., \& Bird,
G. (2020). Testing the independence of self-reported interoceptive
accuracy and attention. \emph{Quarterly Journal of Experimental
Psychology}, \emph{73}(1), 115--133.

\bibitem[\citeproctext]{ref-sibson1978studies}
Sibson, R. (1978). Studies in the robustness of multidimensional
scaling: Procrustes statistics. \emph{Journal of the Royal Statistical
Society: Series B (Methodological)}, \emph{40}(2), 234--238.

\bibitem[\citeproctext]{ref-theriault2024check}
Thériault, R., Ben-Shachar, M. S., Patil, I., Lüdecke, D., Wiernik, B.
M., \& Makowski, D. (2024). Check your outliers! An introduction to
identifying statistical outliers in r with easystats. \emph{Behavior
Research Methods}, \emph{56}(4), 4162--4172.

\bibitem[\citeproctext]{ref-traag2019louvain}
Traag, V. A., Waltman, L., \& Van Eck, N. J. (2019). From louvain to
leiden: Guaranteeing well-connected communities. \emph{Scientific
Reports}, \emph{9}(1), 1--12.

\bibitem[\citeproctext]{ref-veilleux2024development}
Veilleux, J. C., Schreiber, R. E., Warner, E. A., \& Brott, K. H.
(2024). Development and validation of a brief version of the emotion
reactivity scale: The b-ERS. \emph{Current Psychology}, \emph{43}(14),
12586--12600.

\end{CSLReferences}

\end{document}