### 2025-08-12 paper/manuscript.tex

\documentclass[
  jou,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\usepackage{subcaption}

\renewcommand\thesubfigure{\Alph{subfigure}}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{\textbf{The Mint Scale: A Fresh Validation of the Multimodal
Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS}}

\shorttitle{Mint Validation}

\usepackage{etoolbox}

\authorsnames[{1,2},{1},{1},{1},{1}]{Dominique Makowski,Ana Neves,Emma
Benn,Maisie Bennett,Giulia Poreiro}

\authorsaffiliations{
{School of Psychology, University of Sussex},{Sussex Centre for
Consciousness Science, University of Sussex}}

\leftheader{Makowski, Neves, Benn, Bennett and Poreiro}

\abstract{TO DO. }

\keywords{Interoception questionnaire, interoceptive accuracy scale,
MAIA, Mint Validation, Body Awareness, Health}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, toprule=.15mm, arc=.35mm, left=2mm, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0, colback=white]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}

\end{tcolorbox}  Author roles were classified using the Contributor Role
Taxonomy (CRediT; https://credit.niso.org/) as follows:  Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; Ana Neves:   Investigation, Methodology, Project
administration, Supervision, Writing -- original draft, Writing --
review \& editing; Emma Benn:   Investigation; Maisie
Bennett:   Investigation; Giulia
Poreiro:   Investigation, Methodology, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique
Makowski, Email: \href{mailto:D.Makowski@sussex.ac.uk}{D.Makowski@sussex.ac.uk}}
}

\usepackage{pbalance}
% \usepackage{float}
\makeatletter
\let\oldtpt\ThreePartTable
\let\endoldtpt\endThreePartTable
\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.485\textwidth}
\oldtpt[#1]
}
\def\ThreePartTable@ii{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.48\textwidth}
\oldtpt
}
\def\endThreePartTable{
\endoldtpt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\makeatletter
\let\endoldlt\endlongtable		
\def\endlongtable{
\hline
\endoldlt}
\makeatother

\newenvironment{twocolumntable}% environment name
{% begin code
\begin{table*}[!htbp]%
\onecolumn%
}%
{%
\twocolumn%
\end{table*}%
}% end code

\urlstyle{same}

\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

Interoception - the sensing and interpretation of the body's internal
physiological signals - plays a fundamental role in processes ranging
from homeostatic regulation to emotion, decision-making, and mental
health (\citeproc{ref-craig2002how}{Craig, 2002};
\citeproc{ref-critchley2017interoception}{Critchley \& Garfinkel,
2017}). Despite its growing recognition, the field is hindered by
inconsistent definitions and fragmented measurement approaches
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022},
\citeproc{ref-desmedt2025discrepancies}{2025};
\citeproc{ref-khalsa2018interoception}{Khalsa et al., 2018}). Available
tools include ``objective'' physiological tasks (e.g., heartbeat
counting or tracking) and ``subjective'' self-reports, and encompass
both ``explicit'' paradigms that direct participants' attention and
consciousness to internal sensations and ``implicit'' paradigms that do
not (e.g., heartbeat evoked potentials recorded during rest). These
assessments are often targeted at only one of several interoceptive
modalities (e.g., cardioception, respiroception, gastroception), and are
mapped onto theoretical dimensions (e.g., accuracy, sensitivity,
awareness, \citeproc{ref-garfinkel2015knowing}{Garfinkel et al., 2015})
within a rapidly evolving theoretical scaffolding. However, no consensus
exists on a gold-standard measure, and each approach has distinct
limitations: physiological tasks often require specialised equipment and
expertise, are time-consuming, and are challenging to administer at
scale, whereas questionnaires are constrained by their inherent
subjectivity. Crucially, the psychometric quality of many of these
measures is questionable (e.g., see criticism about the heartbeat
counting task, \citeproc{ref-zamariola2018interoceptive}{Zamariola et
al., 2018}), yielding low inter-task correlations
(\citeproc{ref-murphy2018classifying}{Murphy et al., 2018}), raising
serious concerns about their reliability and validity (e.g., measures
presented as measuring one theoretical dimension but likely measuring
something else). This measurement conundrum hinders progress in
comparing, replicating, and interpreting interoception-related findings,
underscoring the need to develop improved tools that align with modern
psychometric standards.

The present work focuses on self-report questionnaires, which offer a
scalable and accessible tool complementing physiological tasks. Among
the numerous questionnaires available, a handful seem to dominate the
literature (\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}),
including the Body Perception Questionnaire (BPQ,
\citeproc{ref-kolacz2018body}{Kolacz et al., 2018}), the
Multidimensional Assessment of Interoceptive Awareness (MAIA,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}), and
the Interoceptive Accuracy Scale (IAS,
\citeproc{ref-murphy2018classifying}{Murphy et al., 2018}). These tools
were developed in different eras of a rapidly evolving field, each with
its own objectives and philosophies. The BPQ, one of the earliest
measures, emphasizes the assessment of body awareness and autonomic
nervous system (ANS) reactivity, focusing on subjective experiences of
physiological stress reactions in organs innervated by the ANS, such as
those triggered by stress or trauma
(\citeproc{ref-porges1993body}{\textbf{porges1993body?}}). The MAIA
incorporated ideas from the then-booming emotion regulation and
mindfulness fields, including subscales assessing adaptive coping
strategies and metacognitive insights about the body. By contrast, the
IAS was recently crafted to sidestep such high-level - possibly
tangential - constructs and instead capture subjective reports of
concrete bodily events (e.g., coughing, sneezing) as proxies for
interoceptive accuracy.

These differing philosophies have practical consequences, leading to
poor convergence between scales that are assumed to measure the same
construct (\citeproc{ref-vig2022questionnaires}{Vig et al., 2022}) - a
classic example of the ``jingle fallacy'' where the shared label of
``interoception'' can mask underlying measurand differences, which in
turn hinders replication and theoretical integration. Recent work has
been focused on aligning subjective measures and facets with theoretical
models, debating whether they measure interoceptive ``sensitivity'' or
other dimensions (accuracy, attention, deficits). However, this focus on
a theory-driven approach might be premature when theoretical models are
in flux, and when a topographic mapping to self-report dimensions is not
guaranteed (e.g.~distinctions that are relevant at a neuroanatomical or
cognitive level might not necessarily translate into the same structural
landscape of subjective experiences). More fundamentally, existing
questionnaires suffer from two major structural blind spots. First, they
are often narrow or unsystematic in their coverage of bodily
\emph{modalities}, with a heavy bias towards cardio-respiratory
sensations and a neglect of others like gastric, urogenital, or
thermoregulatory signals. Second, and perhaps most critically, they fail
to control for the \emph{context} in which an interoceptive experience
occurs, introducing significant variability in how participants
interpret and respond to items. For example, an item about ``feeling
your heart beat'' is ambiguous; its frequency, experience and its
interpretation differ vastly depending on whether it occurs during
physical exercise, a moment of anxiety, or quiet rest (e.g., someone who
exercises regularly, or is anxious, might more easily recall an instance
of occurrence of feeling their heart beating and respond based on that
contextual saliency).

To address these gaps, we introduce the Multimodal Interoception
Questionnaire (Mint). The overarching aim of this work is to develop and
validate a new questionnaire that systematically addresses these
limitations. In Study 1, we detail the bottom-up development of the
Mint, using a ``modality-by-context'' framework to generate a
comprehensive item pool, which is then refined using cutting-edge
psychometric network analysis. In Study 2, we validate the final version
of the Mint against the established interoception questionnaires,
demonstrating its robust structure and predictive validity for a range
of clinically-relevant outcomes.

\section{Study 1: Item Selection}\label{study-1-item-selection}

The primary goal of Study 1 was to develop a new interoception
questionnaire by generating a comprehensive item pool and refining it
into a psychometrically robust scale. To overcome the limitations of
existing measures, we adopted a bottom-up, theoretically agnostic
approach. Instead of imposing a pre-defined theoretical structure, we
focused on two tangible and systematically controllable aspects of
interoceptive experience: the \emph{modality} of the sensation and the
\emph{context} in which it occurs. The objective was not to create
\emph{yet another scale} measuring an alleged distinct construct or
variation, but rather to synthesize existing ideas and frameworks into a
comprehensive and modern tool that captures the richness of
interoceptive experiences and encompasses the most relevant facets of
interoception.

Our item generation process was guided by a ``modality-by-context'' grid
designed to overcome the limitations of existing scales. For
\emph{modalities}, we expanded beyond the typical cardio-respiratory
focus to include a broader range of bodily systems: cardiac,
respiratory, gastric, genital, bladder \& colon, and skin \& temperature
(the latter of which being increasingly recognized as a key
interoceptive channel, \citeproc{ref-crucianelli2023role}{Crucianelli \&
Ehrsson, 2023}). We also included a ``general state'' category to
capture the integrated sense of the body's overall physiological
condition (\citeproc{ref-craig2002how}{Craig, 2002};
\citeproc{ref-craig2003interoception}{Craig, 2003}). For
\emph{contexts}, we designed facets to capture specific situational
contexts like negative (anxious) and positive (sexual) arousal, which
reflect distinct bodily states with differing interoceptive signatures,
as well as other experiential qualities tapping into elaboration and
threshold of perception, such as sensitivity, accuracy, and confusion.
To maintain continuity with established work while making targeted
improvements, we purposefully adapted items from existing questionnaires
(e.g., IAS-like items related to observable behaviors). This systematic
framework yielded a diverse pool of 120 initial items, ensuring
comprehensive coverage while minimizing the ambiguity inherent in
context-free questions.

The final step involved a data-driven reduction of this item pool. We
employed cutting-edge psychometric network-based techniques to identify
and remove redundant items and to uncover the most stable and coherent
dimensional structure. Additionally, we designed the study so that we
could explore an often overlooked psychometric quality: the invariance
of the structure to the presentation format (how items are grouped
together on the screen, whether by modality, facet, or randomly)
influences the resulting psychometric structure. This rigorous,
multi-stage process was designed to produce a subset of items with a
clear, stable, and theoretically rich hierarchical structure, ready for
formal validation.

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We recruited 760 English-speaking participants using
Prolific\textcopyright. We excluded 191 for failing at least one
attention check, and 10 based on measures significantly related to the
probability of failing attention checks (namely, the multivariate
distance obtained with the OPTICS algorithm,
\citeproc{ref-theriault2024check}{Thériault et al., 2024}). The final
sample includes 559 participants (age = 37.0 \(\pm\) 12.2 {[}18, 77{]};
50.8\% women; country of residence: 63.86\% UK, 26.65\% USA). This study
was approved by the University of Sussex' Ethics Committee
(ER/MB2021/1).

\subsubsection{Item Generation}\label{item-generation}

Based on the two goals outlined for this scale, namely to include
different interoceptive modalities, and to explicitly state the context
of the interoceptive experience (e.g., whether negative or positive), we
generated items in a systematic way following a combinatory approach,
where each item's category corresponds to the union of a specific
modality and context (Figure~\ref{fig-one}).

\begin{figure*}[!htbp]

{\caption{{The conceptual grid used to generate the 120 initial items
(top-left). Each item belongs both to an interoceptive modality
(vertical) and a facet (horizontal), with the number of each item per
category indicated in the circles. The asterisk denotes the additional
presence of an attention check item in that category. In the experiment,
these items were presented on different pages grouped either by modality
(bottom-left), by facet (top-right), or randomly. The Correlation
Similarity (bottom-right) analysis suggested that the correlation matrix
obtained from the participants assigned to the random-grouping condition
was slightly more similar (but non-significantly) to the one obtained in
the modality-grouping condition, suggesting that the scale's structure
is robust to different presentation conditions, and that the
modality-grouping might potentially tend to facilitate the emergence of
the underlying item structure (and thus be interpreted as being more
``natural'').}{\label{fig-one}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study1/analysis/figures/figure1.png}
\end{center}

\end{figure*}

We identified 7 ``modalities'' (cardiac, respiratory, gastric, genital,
skin \& temperature (thermoregulation), bladder \& colon, and a
``general state'' category corresponding to a holistic and general
awareness of an interoceptive state or dimension). Through iterative
refinement (e.g., splitting or merging different categories together),
we settled on 6 ``facets'', which encompass both \emph{contexts} of
experience (negative and positive arousal, e.g., anxious and sexual
states), and potential distinct \emph{mechanisms} (nociception \&
pleasure, sensitivity, accuracy, and confusion).

Using this orthogonal 7x6 modality/facet grid as a conceptual
scaffolding, we generated 120 initial items, striving for a balanced
number of items with consistent phrasing within modalities and
facets\footnote{The initial item list is available at
  \href{https://realitybending.github.io/InteroceptionScale/study1/analysis/2_analysis.html}{realitybending.github.io/InteroceptionScale/study1/analysis/2\_analysis.html}}.
We additionally crafted 8 ``attention check'' items blending in (and
distributed across) each category.

\subsubsection{Procedure}\label{procedure}

To avoid presenting all the 120 items on a single long and discouraging
page, we split them into different pages. Participants were randomly
assigned to one of three conditions, driving how items were grouped on
the same page: 1) items grouped by modality (i.e., all cardiac items on
the first page, all colon \& bladder items on the second, etc.), 2)
items grouped by facet, or 3) items presented fully randomly (but with
their number balanced across 6 pages). The order of the item on any
given page and the order of the modalities/facets was randomized. Each
participant completed the full set of 120 items, with 8 attention check
items interspersed throughout. The online experiment was implemented
using JsPsych (\citeproc{ref-de2015jspsych}{De Leeuw, 2015}), and item
responses were recorded using 7-points Likert scales (0 = Disagree, 6 =
Agree).

\subsubsection{Data Analysis}\label{data-analysis}

In order to test whether the grouping condition had an effect on the
structure (i.e., how items relate to one-another), we compared the
correlation matrix obtained in the random condition to the ones obtained
in the modality and facet conditions, focusing on 3 indices of
correlation matrix similarity - the Procrustes Similarity Index (PSI,
\citeproc{ref-sibson1978studies}{Sibson, 1978}), the Adjusted RV (Rvadj,
\citeproc{ref-mayer2011exploratory}{Mayer et al., 2011}), and the
Similarity of Matrices Index (SMI,
\citeproc{ref-indahl2018similarity}{Indahl et al., 2018}). For each
index, we bootstrapped the difference between the similarity with the
facet and modality conditions to test whether the correlation matrix in
the random-grouping condition is significantly more similar to any of
the two other conditions.

Items deemed ``redundant'' (which can distort the item structure
estimation by introducing multicollinearity or local dependencies) were
identified (using the recommended threshold of 0.25) using Unique
Variable Analysis (UVA, \citeproc{ref-christensen2023unique}{Christensen
et al., 2023}), a novel and principled method derived from network
psychometrics.

The structure of the items was analyzed using the recently-developed
Exploratory Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{H.
F. Golino \& Epskamp, 2017}) framework, which allows to jointly estimate
the number of dimensions (i.e., clusters of items), the structure, as
well as its stability using bootstrapping
(\citeproc{ref-golino2020investigating}{H. Golino et al., 2020}). At a
fundamental level, EGA conceptualizes variables as nodes in a network,
with connections (edges) reflecting associations between them. Evidence
has underlined its suitability as an alternative to traditional factor
analysis, addressing some of its limitations such as the assumption of a
``latent'' source of variability, issues in estimation of the optimal
factor numbers, and poor performance in complex population structures,
while remaining comparable and interpretable
(\citeproc{ref-christensen2021equivalency}{Christensen \& Golino, 2021};
\citeproc{ref-jimenez2023dimensionality}{Jiménez et al., 2023}). In
particular, nodes communities (i.e., clusters of items) can be in
practice interpreted as distinct ``dimensions'', similarly to
traditional latent factors - but without explicitly assuming their
existence (\citeproc{ref-christensen2021equivalency}{Christensen \&
Golino, 2021}).

After removing redundant items using UVA, we iteratively fitted
hierarchical EGA models (which additionally estimate higher-order
``meta'' clusters) using ``glasso''
(\citeproc{ref-friedman2008sparse}{Friedman et al., 2008}) and the
``Leiden'' algorithm (\citeproc{ref-traag2019louvain}{Traag et al.,
2019}) for community detection, refining the item pool at each step. We
started by removing items with a low (\textless{} 80\%) cluster
stability (i.e., volatile items which jump between clusters across
bootstrapped samples), followed by odd items belonging to no clusters or
pairs of items (i.e., we keep items belonging to clusters of more than 2
items). Finally, for each lower-level cluster, we selected the 3 items
with the highest node centrality (i.e., the highest loading in the
cluster).

\subsection{Results}\label{results}

The correlation matrix similarity analysis yielded no significant
differences between the similarity of the random-grouping condition with
the modality-grouping and facet-grouping conditions
(\(PSI_{\text{Random vs. Facet}} = 0.81\),
\(PSI_{\text{Random vs. Modality}} = 0.82\), \(p = .45\);
\(RVadj_{\text{Random vs. Facet}} = 0.77\),
\(RVadj_{\text{Random vs. Modality}} = 0.78\), \(p = .74\);
\(SMI_{\text{Random vs. Facet}} = 0.49\),
\(SMI_{\text{Random vs. Modality}} = 0.51\), \(p = .52\)), despite a
consistent bias in favour of the modality-grouping condition.

From the 120 initial items, 4 redundant items were flagged by UVA. We
then removed 40 items that showed low cluster stability, and 9 items
that were part of clusters with less than 3 items. Finally, We kept the
3 items with the highest loading in their lower-level structure
(removing 13 items in the process), resulting in 54 items in the final
item pool.

The final hierarchical EGA model (Generalized Total Entropy Fit Index =
-119.18) - in which all 54 items yielded a high cluster stability
(\textgreater{} 90\%) suggested 3 metaclusters and 15 lower-level
clusters (each containing 3 items): ``Interoceptive Deficits''
(containing 5 clusters: \emph{Urointestinal Inaccuracy} - UrIn;
\emph{Cardiorespiratory Confusion} - CaCo; \emph{Cardiorespiratory
Noticing} - CaNo; \emph{Olfactory Compensation} - Olfa; \emph{Satiety
Noticing} - Sati), ``Interoceptive Awareness'' (containing 7 clusters:
\emph{Sexual Arousal Awareness} - SexA; \emph{Sexual Arousal
Sensitivity} - SexS; \emph{Sexual Organs Sensitivity} - SexO;
\emph{Urointestinal Sensitivity} - UrSe; \emph{Relaxation Awareness} -
RelA; \emph{State Specificity} - StaS; \emph{Expulsion Accuracy} -
ExAc), and ``Interoceptive Sensitivity'' (containing 6 clusters:
\emph{Cardioception} - Card; \emph{Respiroception} - Resp;
\emph{Signalling} - Sign; \emph{Gastroception} - Gast; \emph{Dermal
Hypersensitivity} - Derm; \emph{Sexual Arousal Changes} - SexC).

To further reduce and balance the remaining items, we collectively
decided on removing the lower-level clusters \emph{Sexual Arousal
Awareness} (too general and overlapping with the other more specific
sex-related items), \emph{Signalling} (which items started with ``when
something important is happening in my life'', which meaning we deemed
too much open to interpretation), and \emph{Sexual Arousal Changes} (low
fit with the other modality-focused clusters of its group). The final
set included 45 items.

\subsection{Discussion}\label{discussion}

The primary aim of Study 1 was to develop and refine a large,
systematically generated pool of interoceptive items into a
psychometrically sound and structurally coherent questionnaire. Starting
with 120 items derived from a modality-by-context grid, we employed a
multi-step, data-driven reduction process using cutting-edge structure
analysis methods. This rigorous procedure yielded a final, balanced set
of 45 items, organized within a stable hierarchical structure comprising
three higher-order dimensions (Interoceptive Deficits, Interoceptive
Awareness, Interoceptive Sensitivity) and 12 distinct lower-level
facets.

One innovation of this study was to test the effect of item presentation
by comparing structures across different grouping conditions. While the
analysis revealed no significant effect, a consistent (though
non-significant) trend suggested that grouping items by modality may be
slightly more conducive to revealing the underlying psychometric
structure than grouping by facet or randomly. This finding, while
preliminary due to the sample size per condition (\(N \pm 200\)),
highlights a potentially important and under-investigated aspect of
questionnaire design that warrants future research.

The resulting hierarchical structure offers novel yet interpretable
insights into self-reported interoception, with flexible degrees of
granularity. Notably, we categorised several facets (typically related
to objective external behaviours, like coughing or sneezing) under the
label ``Awareness'', that might traditionally be named ``Accuracy'' in
other scales. Our rationale is that subjective self-report, by its very
nature, cannot be a pure measure of objective accuracy, but at best a
proxy of an individual's subjective certainty or awareness of their
interoceptive processes. Furthermore, our bottom-up approach revealed
intriguing facets, such as ``Olfactory Compensation''. This finding
provide further evidence for a pseudo-interoceptive pathway, whereby
individuals with potentially poorer access to internal signals may
compensate by relying on external cues (e.g., odours) to infer their
bodily state. The ``Sensitivity'' metacluster also grouped core
visceroceptive facets (e.g., ``Cardioception'', ``Gastroception'') with
``Dermal Hypersensitivity'', pointing to a potential common dimension of
heightened perceptual sensitivity to bodily sensations.

In summary, Study 1 successfully leveraged a systematic item generation
process and modern analytical techniques to produce a 45-item
questionnaire with a clear, stable, and theoretically rich hierarchical
structure. The next crucial step is to formally validate this new
instrument by examining its psychometric properties and its relationship
with existing measures in a new, independent sample, which is the
objective of Study 2.

\section{Study 2: Validation}\label{study-2-validation}

Following the generation and selection of a set of 45 interoceptive
items, the next objective is to formally validate this new instrument in
a large, independent sample, a crucial step for establishing the
psychometric soundness and unique contribution of the Mint scale to the
field of interoception research. To this end, we sought to 1) re-assess
the structure of the shorter version and 2) establish its convergent
validity by examining the correlations with other established and
popular interoception questionnaires (namely the MAIA-2, BPQ, and IAS),
as well as other related measures (e.g., alexithymia). Finally, we
assessed its criterion validity by directly comparing the predictive
power of the interoceptive scales across a broad range of outcomes,
including emotional, cognitive, and clinical variables.

\subsection{Methods}\label{methods-1}

\subsubsection{Participants}\label{participants-1}

We recruited 921 English-speaking participants via SONA and
Prolific\textcopyright, from which 118 were excluded for failing at
least one attention check and 6 based on multivariate distance (using
the same procedure as for study 1). 60 participants were further
excluded due to missing data following a technical issue. The final
sample includes 737 participants (age = 36.8 \(\pm\) 14.7 {[}18, 87{]};
57.3\% women; Country of residence: 75.17\% United Kingdom, 24.83\%
other). This study was approved by the University of Sussex' Ethics
Committee (ER/EB672/2).

\subsubsection{Measures}\label{measures}

\begin{figure*}[!htbp]

{\caption{{Structure analysis of the final set of 33 Mint items.
Hierarchical Exploratory Graph Analysis (EGA) was applied to jointly
identify clusters of items, as well as higher-order metaclusters. The
reliability of each item can be quantified by the proportion of
structure replication across bootstrapped samples, with high values
indicative of high structural stability. The Mint scale displayed an
excellent structural consistency, with the exception of the \emph{Olfa}
cluster which belonging to the \emph{Deficit} metaclusters warrants
further research. Bottom plot shows the result of hierarchical
clustering, providing evidence for structure statbility across methods,
and allowing for more granular understanding of the relationships
between variables.}{\label{fig-two}}}}

\begin{center}
\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig1.png}
\end{center}

\end{figure*}

\paragraph{Interoception
Questionnaires.}\label{interoception-questionnaires}

The 45 items of the Multimodal Interoception Questionnaire (Mint) scale
were presented in a random order, with the same 7-point Likert scale as
in study 1 (0 = \emph{Disagree}, 6 = \emph{Agree}).

The Multidimensional Assessment of Interoceptive Awareness (MAIA-2,
\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018})
measures 8 dimensions with 37 items presented on a 7-point Likert scale
(0 = \emph{Never}, 6 = \emph{Always}). It includes bodily dimensions
such as Body Listening, Noticing, Emotional Awareness, dimensions
related to emotion regulation, such as Trust, Not-worrying, and
Self-Regulation, as well as dimensions related to attention, such as
Attention Regulation and Not-Distracting. The relationship of some of
these dimensions to interoception remains debated, particularly
Not-Distracting and Not-Worrying, which tend to show weak or
non-existent correlations with the other MAIA subscales
(\citeproc{ref-ferentzi2021examining}{Ferentzi et al., 2021}).

The Body Perception Questionnaire - Very Short Form (BPQ-VSF,
\citeproc{ref-cabrera2018assessing}{Cabrera et al., 2018}) measures a
general score of bodily awareness with 12 items presented on 5-point
Likert scale (0 = \emph{Never}, 5 = \emph{Always}).

The Interoceptive Accuracy Scale (IAS,
\citeproc{ref-murphy2020testing}{Murphy et al., 2020}) measures a
general score of interoceptive accuracy to physical sensations with 21
items presented on a 5-point Likert scale (1 = \emph{Disagree Strongly},
5 = \emph{Strongly Agree}).

\paragraph{Emotions and Cognition.}\label{emotions-and-cognition}

The Toronto Alexithymia Scale TAS (TAS-20,
\citeproc{ref-leising2009toronto}{Leising et al., 2009}) measures 3
Alexithymia dimensions with 20 items 5-point Likert scale (1 =
\emph{Strongly Disagree}, 5 = \emph{Strongly Agree}: Difficulty
Identifying Feelings (DIF), Difficulty Describing Feelings (DDF), and
Externally Oriented Thinking (EOR).

The Emotion Reactivity Scale - Brief Version (B-ERS,
\citeproc{ref-veilleux2024development}{Veilleux et al., 2024}) measures
3 dimensions with 6 items on a 5-point Likert scale (0 = \emph{Not like
me at all}, 4 = \emph{Extremely like me}): Arousal (\emph{``My moods are
very strong and powerful''}), Sensitivity (\emph{``I tend to get very
emotional very easily''}), and Persistence (\emph{``When I am
angry/upset, it takes me much longer than most people to calm down''}).

The Cognitive Emotion Regulation Questionnaire (CERQ-short,
\citeproc{ref-garnefski2006cognitive}{Garnefski \& Kraaij, 2006})
measures 9 adaptive and maladaptive emotion regulation strategies with
18 items (1 = \emph{Almost Never}, 5 = \emph{Almost Always}). The
strategies include Self-blame, Other-blame, Rumination, Catastrophizing,
Putting into Perspective, Positive Refocusing, Positive Reappraisal,
Acceptance, and Planning.

The Černis Felt Sense of Anomaly scale - short form (CEFSA-14,
\citeproc{ref-vcernis2024measuring}{Černis et al., 2024}) measures 7
dissociative experiences with 14 items on a 5-point Likert scale (0 =
\emph{Never}, 4 = \emph{Always}). This scale is designed to measure
dissociative experiences in adolescence and adulthood, specifically
focusing on the felt sense of anomaly-type dissociation. It includes
Anomalous Experience of the Self, Anomalous Experience of the Body,
Anomalous Experiences of Emotion, Altered Sense of Familiarity, Altered
Sense of Connection, Altered Sense of Agency, and Altered Sense of
Reality.

The Primal World Beliefs Inventory (PI-18,
\citeproc{ref-clifton2021brief}{Clifton \& Yaden, 2021}) measures
higher-order beliefs about the basic character of the world: Good, Safe,
Enticing, and Alive. The 18 items are presented on a 6-point Likert
scale (0 = \emph{Strongly Disagree}, 5 = \emph{Strongly Agree}).

\paragraph{Health and Wellbeing.}\label{health-and-wellbeing}

The Single Item Life Satisfaction scale (SILS,
\citeproc{ref-jovanovic2020longer}{Jovanović \& Lazić, 2020}) was
followed by the PHQ-4 (\citeproc{ref-kroenke2009ultra}{Kroenke et al.,
2009}) assessing depression and anxiety symptoms with 4 items. We used
the refined version of the PHQ-4
(\citeproc{ref-makowski2025measuring}{Makowski et al., 2025}), which
adds an additional response option (``Once or twice''), increasing
sensitivity to subclinical fluctuations.

Participants were asked to self-report any current, medically diagnosed
psychiatric disorders using a checklist based on DSM-5 categories. If
one or more conditions were endorsed, participants were asked to
indicate any current treatments, including pharmacological (e.g.,
antidepressants, anxiolytics, antipsychotics, mood stabilizers),
psychological (e.g., psychotherapy, mindfulness), or lifestyle
interventions. Binary variables (0 = absent, 1 = present) were created
to identify participants reporting mood disorders (MDD, GAD, Bipolar
Disorder; with a stricter subgroup of participants undergoing a pharma-
or psychological treatment), anxiety-centred disorders (GAD, Panic
Disorder, Social Anxiety Disorder, Specific Phobias), eating disorders,
addiction-related disorders, borderline personality disorder, autism
spectrum disorder (ASD), and ADHD.

Participants were asked to select somatic symptoms or conditions they
experienced from a list of 36 options. To facilitate mechanistic
interpretation and reduce redundancy, answers were grouped into four
non-overlapping clusters based on shared physiological pathways and
known etiological mechanisms. The \emph{Afferent Sensitivity} cluster
included conditions associated with heightened interoceptive awareness
and neurogenic excitability, such as migraine, neuropathy, dizziness,
nausea, muscle tension, epilepsy, and frequent urination. The
\emph{Central Sensitization} cluster comprised syndromes characterized
by chronic pain and fatigue, likely reflecting central amplification of
sensory signals and HPA-axis dysregulation, such as fibromyalgia,
chronic fatigue syndrome, chronic pain, back pain, pelvic pain,
irritable bowel syndrome (IBS). The \emph{Autonomic Dysfunction} cluster
captured disorders linked to dysregulation of the autonomic and
cardiopulmonary systems, including joint hypermobility, cardiac
arrhythmia, chest pain, shortness of breath, hypo-/hypertension, sleep
apnea, chronic obstructive pulmonary disease (COPD), and chronic
bronchitis. Finally, the \emph{Immune-Inflammatory} cluster encompassed
conditions associated with immune dysregulation, barrier dysfunction,
and gut-brain axis disturbance, such as eczema, psoriasis, skin rashes,
asthma, celiac disease, gluten and lactose sensitivity, inflammatory
bowel diseases (Crohn's disease and ulcerative colitis),
gastroesophageal reflux disease (GERD), multiple sclerosis, and
Sjögren's syndrome. We scored each cluster as a binary variable based on
whether the participant selected at least one symptom from that cluster.

\paragraph{Lifestyle.}\label{lifestyle}

Participants reported about owning any wearable devices to monitor
health indices such as heart rate, number of steps, calories burned,
sleep quality, and weight. For each selected device, a follow-up
question inquired about the frequency of usage and subjective importance
of that measure.

Participants were asked to rate how physically active they consider
themselves and how many hours of active workout they engage in per week.
Participant's BMI was computed using height and weight.

\subsubsection{Procedure}\label{procedure-1}

In order to avoid the repetition of similar types of questions and and
balance longer and shorter questionnaires, we partitioned the measures
into three groups (and randomized the order within them): 1)
interoception questionnaires (MAIA, IAS, BPQ), 2) emotions (TAS, CERQ,
ERS, PI-18), and 3) health (somatic symptoms, mental health, LS + PHQ-4,
CEFSA). After completing demographic questions, participants always
started with the Mint scale, and each following interoception
questionnaire was interspersed with two questionnaires from the emotions
and pathology groups. 7 attention checks were embedded throughout (one
within each major questionnaire). In order to make the experiment more
enjoyable, the experiment ended with a radar chart summarizing the
participants' responses to the Mint scale\footnote{The experiment can be
  tested by following the link on
  https://github.com/RealityBending/InteroceptionScale}.

\begin{figure*}[!htbp]

{\caption{{Item loadings. The table shows each item of the Mint with its
cluster centrality, which is the EGA equivalent (in terms of
interpretation) of factor loadings. It also shows how each lower-level
cluster is related to higher-order metaclusters. The lower-level
clusters are Expulsion Accuracy (ExAc), Relaxation Awareness (RelA),
Sexual Arousal Sensitivity (SexS), Cardiorespiratory Confusion (CaCo),
Urointestinal Inaccuracy (UrIn), Dermal Hypersensitivity (Derm), Satiety
Noticing (Sati), Olfactory Compensation (Olfa), Respiroception (Resp),
Cardioception (Card), Gastroception (Gast).}{\label{fig-three}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/table1.png}
\end{center}

\end{figure*}

\subsubsection{Data Analysis}\label{data-analysis-1}

We started by confirming and further refining the structure of the Mint
scale (see Figure~\ref{fig-two}) using the same EGA model as in Study 1.
We then computed the lower-level dimensions and the higher-level
metaclusters' scores by averaging their corresponding items. The
convergent validity of the final set of items was assessed by computing
the correlations between the Mint scale and the other interoception
questionnaires (MAIA, BPQ, IAS).

Next, we tested the predictive power of the Mint scale relative to other
interoception questionnaires. We will assessed the importance of each
interoceptive dimension (from all the scales) as a unique predictor by
fitting regression models (linear for continuous measures - e.g.,
depression score - and logistic for binary variables - e.g., presence
\emph{vs.} absence of mood disorders) to predict different outcomes, and
compare the best between the 4 interoception scales (based on the R2).
We then evaluated the predictive performance of each scale as a whole by
comparing regression models with all the dimensions entered as
predictors (note that the IAS and BPQ only have one total score
variable). We assessed the models based on their total explained
variance (R2), as well as on the Bayesian Information Criterion (BIC),
which penalizes for predictor number, thus offering a balance between
model parsimony and predictive power. For the logistic models, we
quantified the discriminative power by computing the Area Under the
Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which
assesses the model's discriminative power (the combination of
sensitivity and specificity).

In order to evaluate the potential for a short version of our scale, we
compared 4 variants of the Mint: the (full) \emph{Mint} (including all
lower-level dimensions), the \emph{metaMint} (only including the
metaclusters), the \emph{miniMint} (including the metaclusters computed
from a reduced number of items), and the \emph{microMint} (including
only the most representative dimension of each metacluster). Moreover,
we also included an alternative ``interoception-focused'' version of the
MAIA (\emph{iMAIA}) that only contains the 3 most interoceptive
dimensions (Body Listening, Noticing, and Emotional
Awareness)\footnote{See correlation results to further justify this
  selection.}.

\subsection{Results}\label{results-1}

The application of EGA to the initial set of 42 items reproduced the
expected 14 lower-level clusters of triplets and 3 higher-order
metaclusters, with the exception of the \emph{UrSe} (Urointestinal
Sensitivity), for which one item moved to the \emph{Olfa} (Olfactory
Compensation) cluster. In order to further balance and reduce the items,
we removed the \emph{UrSe} cluster, as well as \emph{StaS} (State
Specificity), which in comparison to the other items stood out as
containing vague and overlapping items ; \emph{SexO} (Sexual Organs
Sensitivity), which overlapped with the \emph{SexS} (Sexual Arousal
Sensitivity) cluster; and \emph{CaNo} (Cardiorespiratory Noticing),
which overlapped with the \emph{CaCo} (Cardiorespiratory Confusion)
cluster.

The final set of 33 items yielded a good fit (Generalized Total Entropy
Fit Index = -77.23), with all items showing high cluster stability
(\textgreater{} 90\%), with the exception of \emph{Olfa}. The final
structure included 11 lower-level clusters, grouped into 3 higher-order
metaclusters: ``Interoceptive Deficit'' (containing \emph{CaCo},
\emph{UrIn}, \emph{Derm}, \emph{Sati}, \emph{Olfa}), ``Interoceptive
Awareness'' (containing \emph{ExAc}, \emph{RelA}, \emph{SexS}), and
``Visceroception'' (containing \emph{Card}, \emph{Resp}, \emph{Gast}).
Item centrality (intepretable as loadings in traditional factor analysis
frameworks) are shown in Figure~\ref{fig-three}. Additionally, we
applied hierarchical clustering analysis which replicated this structure
(see Figure~\ref{fig-two}), suggesting consistency across methods.

\begin{figure*}[!htbp]

{\caption{{Correlation Matrices between the Mint dimensions
(upper-left), and between the Mint and other interoception
questionnaires (MAIA, BPQ and IAS; upper-right). The bottom matrix shows
the relationships between interoception and other measures included in
the study, such as alexithymia (TAS-20), depression and anxiety (PHQ-4),
emotion reactivity (ERS), abnormal and dissociative experiences (CEFSA),
and Primal World Beliefs. Stars indicate dimensions that have been
score-reversed for better in-context interpretation. Correlation
coefficients are shown only for significant correlations (p \textless{}
.001).}{\label{fig-four}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig2.png}
\end{center}

\end{figure*}

The correlation matrix of the Mint dimensions revealed an interesting
and rich tapestry of relationships (see Figure~\ref{fig-four}), with
contrasting patterns of associations with dimensions from the same
group. For instance, \emph{Derm} and \emph{Olfa}, despite being
positively correlated with the other dimensions from the \emph{Deficit}
cluster, were not negatively correlated to \emph{Awareness} and its
dimensions. Similarly, \emph{Sati}, unlike the other \emph{Deficit}
dimensions, was not positively correlated to \emph{Visceroception}. This
complex structure suggests that the dimensions indeed capture distinct
aspects of interoception, and that the metaclusters, rather than being
simple aggregates of rather-redundant elements, might actually capture
unique combinations (greater than the sum of their parts). It also
provides evidence against a simplistic adaptive \emph{vs.} maladaptive
dichotomy, as \emph{Deficit} was not negatively bit positively
correlated with \emph{Visceroception}. Interestingly, while
\emph{Awareness} was also positively correlated with
\emph{Visceroception}, it yielded an insignificant negative correlation
with \emph{Deficit} (this finding was also aligned with the hierarchical
clustering results, showing a greater proximity between \emph{Deficit}
and \emph{Visceroception} than with \emph{Awareness}), underlining again
the complex web of relationships captured by the Mint.

The correlation matrix between the Mint and the other interoception
scales revealed high levels of overlap, as well as some unique
contributions (see Figure~\ref{fig-four}). The BPQ was positively
correlated with most Mint dimensions (the highest with
\emph{Visceroception}, \(r = .46\)). The IAS was positively correlated
with \emph{Visceroception} and \emph{Awareness} dimensions (the highest
with \emph{Awareness}, \(r = .63\)), but negatively with most
\emph{Deficit} dimensions (with the exception of \emph{Olfa} and
\emph{Derm}). In turn, the \emph{Visceroception} metacluster most
strongly correlated with MAIA's Noticing (\(r = .55\)). Interestingly,
MAIA's TTrusting correlated selectively with \emph{Awareness}, and
negatively with \emph{Deficit} dimensions, but not with
\emph{Visceroceptive} dimensions (underlining its high-level
metacognitive nature). MAIA's Emotional Awareness and Body Listening
displayed a similar pattern to Noticing, and Attention Regulation and
Self Regulation positively correlated with \emph{Awareness} and
\emph{Visceroception} dimensions, but negatively with some
\emph{Deficit} dimensions (\emph{CaCo} and \emph{UrIn}). Not-distracting
only yielded mild negative associations with \emph{Sati} and
\emph{Olfa}. Overall, the Mint dimensions were able to capture most of
the (relevant) variance and intricacies present in the other
interoception scales.

Exploratory correlations with the emotion regulation (CERQ) dimensions
revealed stronger associations with most of the MAIA dimensions
(supporting its proximity with emotion regulation). Interestingly,
Rumination (and Self-Blame) stood out as selectively related to the
Mint's \emph{Visceroception} and \emph{Deficit} dimensions (note that
Rumination was also related to the MAIA's interoceptive dimensions,
namely Noticing, Emotional Awareness and Body Listening). The only
primal world belief that correlated particularly with the Mint's
\emph{Visceroception} and the MAIA's interoceptive dimension was the
belief that the world was alive (\textbf{TODO: Ana maybe throw in a
couple more words about what this belief means and is defined}).

\begin{figure*}[!htbp]

{\caption{{Summary table of the comparison between interoception
questionnaires (Mint, MAIA, BPQ and IAS). For various outcomes included
the study, we tested the Mint and Non-Mint dimension that had the
strongest link (as a unique predictor). Values in parenthesis represents
correlation coefficient for continuous variables and log-odds ratios for
binary variables (the occurrence of mental and somatic health
conditions). We also compared the predictive performance of regression
models including multiple predictors, and present their ranking based on
raw predictive power (R2) and BIC, which favours more parsimonious
models (with less predictors). Green background indicates an advantage
for the Mint, while red backgrounds indicate an advantage for another
interoception scale. Orange backgrounds indicate an advantage for the
MAIA driven by its emotion-regulation dimensions (Trusting and
Not-Worrying).}{\label{fig-five}}}}

\begin{center}
\includegraphics[width=0.82\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/table2.png}
\end{center}

\end{figure*}

Most of the target outcomes measured in the study to assess validity
were best predicted by one of the Mint dimension (see
Figure~\ref{fig-five}). This included Alexithymia (best predicted by
\emph{Deficit}); ERS' Arousal (best predicted by \emph{Deficit});
CEFSA's anomalous experiences of the body (best predicted by
\emph{CaCo}), Self and Reality (best predicted by \emph{Deficit}); ADHD
and Autism (best predicted by \emph{Deficit}), Somatic symptoms (best
predicted by \emph{Deficit}); BMI (best predicted by \emph{Sati}). Most
of the exceptions showed an advantage for MAIA dimensions, in particular
Not-worrying (which best predicted ERS' Sensitivity and Persistence) and
Trusting (which best predicted LS, Depression and Anxiety, CEFSA's
Emotion and Connection, and self-reported physical activity).

Model comparison included 4 variants of the Mint scale, the full
\emph{Mint} (including all lower-level clusters - 33 items), the
\emph{metaMint} (including only the 3 metaclusters based on all items),
the \emph{miniMint} (including the 3 metaclusters computed from 2 of its
most loading triplets - 18 items), and the \emph{microMint} (including
only the most representative dimension of each metacluster - 9 items).
This analysis confirmed the clear advantage for the Mint over the other
interoception scales. The full \emph{Mint} model was typically the best
model based on R2, and the \emph{metaMint} was the best model when
parsimony was taken into account (i.e., based on BIC). Moreover, most of
the instances were the \emph{MAIA} was the best model were explainable
by the inclusion of Trusting, and the interoception-only version
\emph{iMAIA} typically yielded worse performance than the Mint. In many
instances, the \emph{miniMint} yielded reasonable performance, although
the \emph{microMint} was typically less promising.

The predictive performance for the mental health outcomes (see
Figure~\ref{fig-six}) displayed a consistent pattern, with an advantage
for the \emph{MAIA} for depression and anxiety which dropped below the
\emph{Mint} versions for its \emph{iMAIA} variant. The \emph{Mint},
however, displayed a clear advantage for the prediction of autism, ADHD,
and all somatic groups of symptoms (with the exception of Central
Sensitization, which was best predicted by the \emph{MAIA}).

Finally, the importance of heart monitoring (for owners of such
wearable) was best predicted by MAIA's Body listening (\(r = .36\)),
followed the \emph{Card} dimension of the Mint (\(r = .31\)). The
importance of sleep monitoring importance was best predicted by
\emph{RelA} (\(r = .22\)), followed by MAIA's not worrying
(\(r = -.219\)). Daily activity via steps monitoring was best predicted
by MAIA's Noticing (\(r = .28\)), followed by \emph{Awareness}
(\(r = .26\)).

\begin{figure*}[!htbp]

{\caption{{Predictive performance of interoception questionnaires for
various self-reported mental health conditions (mood disorders with
psychopharmacological treatment, ADHD and ASD) and somatic symptoms. The
Receiver Operating Characteristic (ROC) curves are shown for the
logistic regression models for various questionnaires and versions. A
high Area Under the Curve (AUC) indicates a good discriminative power of
the model (i.e., a strong combination of sensitivity and specificity).
The Mint scale (in blue-green) consistently outperformed the other
interoception scales, with the exception of the MAIA which, driven by
its emotion-regulation dimensions (Trusting and Not-worrying), performed
better for mood disorders and central sensitization symptoms. In several
instances, the short versions of the Mint scale (miniMint and microMint)
yielded reasonable performance, still outperforming other
measures.}{\label{fig-six}}}}

\begin{center}
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../study2/analysis/figures/fig3.png}
\end{center}

\end{figure*}

\subsection{Discussion}\label{discussion-1}

The primary goal of Study 2 was to validate the items selected in Study
1. This involved re-assessing its structure in a new, large sample, and
establishing its convergent and criterion validity against other
established interoception questionnaires. The analysis led to a final,
refined 33-item Mint scale with a stable and theoretically coherent
hierarchical structure, which demonstrated substantial predictive power
across a range of outcomes.

Re-validating the structure after item selection is a crucial step in
questionnaire development. When a large item pool is reduced, the
context changes, which can alter the perceived meaning and relationships
between the remaining items. For instance, a questionnaire with many
items suggesting abnormality or pathology might colour other neutral
items with a negative connotation. While the original structure was
mostly replicated, the re-validation allowed for a clarification of some
of the higher-order dimensions (for instance, the ``Sensitivity''
metacluster became more clearly defined around core interoceptive
clusters). The final structure, comprising 11 facets and 3 metaclusters,
showed excellent structural consistency and stability, providing a solid
foundation for the scale. However, the \emph{Olfa} (Olfactory
Compensation) cluster showed lower stability in its placement within the
\emph{Deficit} metacluster. This suggests its role may be more complex,
and future studies are needed to investigate its place and invariance
across different populations and contexts.

The criterion validity analysis revealed a clear and compelling pattern
of results. The Mint scale, particularly its \emph{Deficit} metacluster,
consistently outperformed other scales in predicting outcomes directly
related to alterations in bodily experience and self-regulation, such as
alexithymia, dissociative experiences, ADHD, autism, and various somatic
symptoms. The main contender, the MAIA-2, showed an advantage primarily
for outcomes related to general well-being, such as life satisfaction,
depression, and anxiety. However, this advantage was almost entirely
driven by dimensions highly related to emotion regulation abilities,
namely \emph{Trusting} and \emph{Not-Worrying}, whose inclusion as core
interoceptive facets has been questioned
(\citeproc{ref-ferentzi2021examining}{Ferentzi et al., 2021}). When
these dimensions were removed (in the \emph{iMAIA} version), the
MAIA-2's predictive power for these outcomes dropped considerably, often
below that of the Mint.

This highlights a crucial issue in the field: the potential for
misattributing findings to ``interoception'' when they may be driven by
broader emotion regulation capacities
(\citeproc{ref-murphy2020testing}{Murphy et al., 2020}). While valuable,
the MAIA-2's most predictive subscales appear less related to core
interoceptive processing and more to metacognitive beliefs about one's
body and emotional coping. The Mint, in contrast, appears to tap more
directly into distinct facets of interoceptive experience, from visceral
sensitivity to perceived deficits. The pattern is clear: for high-level,
metacognitive aspects of affective state and well-being, the MAIA-2's
emotion regulation facets are highly relevant. For outcomes tied more
specifically to bodily processes, clinical conditions characterized by
sensory dysregulation, and nuanced interoceptive experiences, the Mint
offers superior predictive validity and specificity.

\section{General Discussion}\label{general-discussion}

Across two studies, we developed and validated the Multimodal
Interoception Questionnaire (Mint), a new 33-item self-report measure of
interoception. Our approach addressed key limitations of existing scales
by starting with a systematically generated, broad item pool that
explicitly controlled for context and modality, and then using a
data-driven reduction process. The result is a psychometrically robust
instrument with a stable, hierarchical structure comprising three
higher-order metaclusters, namely \emph{Interoceptive Deficit},
\emph{Interoceptive Awareness}, and \emph{Visceroception}, and 11
distinct lower-level facets. The validation study demonstrated that the
Mint not only captures variance present in other popular scales (MAIA-2,
BPQ and IAS) but also offers superior predictive power for a range of
clinical and cognitive outcomes, particularly those tied to altered
bodily perception.

A key contribution of the Mint is its potential ability to disentangle
core interoceptive experiences from more general emotion regulation
capacities. Our findings show that while the MAIA-2 is a potent
predictor of general well-being, this is largely driven by its
\emph{Trusting} and \emph{Not-Worrying} subscales, which heavily tap
into emotion regulation skills and coping strategies. The Mint,
conversely, excelled at predicting conditions like alexithymia, ADHD,
autism, and somatic symptoms, where alterations in bodily sensing are a
core feature. This suggests the Mint provides a more specific measure of
interoceptive processing, offering researchers a finer-grained tool to
probe the role of bodily awareness in health and disease.

Despite these strengths, the present work has limitations that open
avenues for future research. First, as a self-report measure, the Mint
captures subjective beliefs about interoceptive abilities, not objective
accuracy. A crucial next step is to investigate the relationship between
Mint scores and physiological or behavioural measures of interoception
(e.g., heartbeat tracking tasks, heartbeat evoked potentials). This
would help establish the extent to which subjective reports on the Mint
correspond to objective interoceptive performance. Future research could
also systematically examine the cost of shorter versions, by
investigating the tradeoffs incurred by versions with only the one or
two most central facets per metacluster and their impact on psychometric
qualities and predictive power.

While the overall structure of the Mint is stable, some facets warrant
deeper investigation. The \emph{Olfa} (Olfactory Compensation) and
\emph{Derm} (Dermal Hypersensitivity) clusters, for instance, showed
more complex patterns of association. Their placement within the
\emph{Deficit} metacluster is theoretically intriguing but requires
further validation. The former aligns with predictive coding models
suggesting that individuals may rely on exteroceptive cues (like odours)
to infer internal states (\citeproc{ref-paulus2010interoception}{Paulus
\& Stein, 2010}), a form of pseudo-interoception (or
extero-interoception). The latter points to a potential link between
visceral and cutaneous sensitivity, possibly reflecting a general trait
of heightened bodily awareness or a shared mechanism of central
sensitization, which has been observed in conditions like chronic pain
(\citeproc{ref-di2016alexithymia}{Di Tella \& Castelli, 2016}). Future
studies should examine the invariance of the Mint's structure across
different populations (e.g., clinical vs.~non-clinical groups) and
contexts (e.g., heightened arousal vs.~rest) to clarify the role of
these more nuanced facets. Such work could reveal whether their
relevance and pattern of associations differs in conditions like anxiety
disorders, chronic pain, or neurodevelopmental disorders.

In conclusion, the Mint scale represents a significant methodological
advancement for the field of interoception research. By providing a
validated, structurally coherent, and nuanced tool, it allows for a more
precise assessment of self-reported interoceptive experiences. Its
various forms (full, meta, mini, and micro) offer flexibility for
researchers and clinicians, paving the way for a deeper understanding of
the multifaceted role of interoception in human psychology and
pathophysiology.

\section{Data Availability}\label{data-availability}

Data, code, and all materials are available at
https://github.com/RealityBending/InteroceptionScale.

\section{Acknowledgements}\label{acknowledgements}

We would like to thank the dissertation students from the University of
Sussex for their help in data collection. DM would also like to thank
Anastasia for the motivation provided to write this paper.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-cabrera2018assessing}
Cabrera, A., Kolacz, J., Pailhez, G., Bulbena-Cabre, A., Bulbena, A., \&
Porges, S. W. (2018). Assessing body awareness and autonomic reactivity:
Factor structure and psychometric properties of the body perception
questionnaire-short form (BPQ-SF). \emph{International Journal of
Methods in Psychiatric Research}, \emph{27}(2), e1596.

\bibitem[\citeproctext]{ref-vcernis2024measuring}
Černis, E., Loe, B. S., Lofthouse, K., Waite, P., Molodynski, A.,
Ehlers, A., \& Freeman, D. (2024). Measuring dissociation across
adolescence and adulthood: Developing the short-form {Č}ernis felt sense
of anomaly scale ({Č}EFSA-14). \emph{Behavioural and Cognitive
Psychotherapy}, \emph{52}(2), 163--177.

\bibitem[\citeproctext]{ref-christensen2023unique}
Christensen, A. P., Garrido, L. E., \& Golino, H. (2023). Unique
variable analysis: A network psychometrics method to detect local
dependence. \emph{Multivariate Behavioral Research}, \emph{58}(6),
1165--1182.

\bibitem[\citeproctext]{ref-christensen2021equivalency}
Christensen, A. P., \& Golino, H. (2021). On the equivalency of factor
and network loadings. \emph{Behavior Research Methods}, \emph{53}(4),
1563--1580.

\bibitem[\citeproctext]{ref-clifton2021brief}
Clifton, J. D., \& Yaden, D. B. (2021). Brief measures of the four
highest-order primal world beliefs. \emph{Psychological Assessment},
\emph{33}(12), 1267.

\bibitem[\citeproctext]{ref-craig2002how}
Craig, A. D. (2002). How do you feel? Interoception: The sense of the
physiological condition of the body. \emph{Nature Reviews Neuroscience},
\emph{3}(8), 655--666. \url{https://doi.org/10.1038/nrn894}

\bibitem[\citeproctext]{ref-craig2003interoception}
Craig, A. D. (2003). Interoception: The sense of the physiological
condition of the body. \emph{Current Opinion in Neurobiology},
\emph{13}(4), 500--505.

\bibitem[\citeproctext]{ref-critchley2017interoception}
Critchley, H. D., \& Garfinkel, S. N. (2017). Interoception and emotion.
\emph{Current Opinion in Psychology}, \emph{17}, 7--14.
\url{https://doi.org/10.1016/j.copsyc.2017.04.020}

\bibitem[\citeproctext]{ref-crucianelli2023role}
Crucianelli, L., \& Ehrsson, H. H. (2023). The role of the skin in
interoception: A neglected organ? \emph{Perspectives on Psychological
Science}, \emph{18}(1), 224--238.

\bibitem[\citeproctext]{ref-de2015jspsych}
De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating
behavioral experiments in a web browser. \emph{Behavior Research
Methods}, \emph{47}, 1--12.

\bibitem[\citeproctext]{ref-desmedt2022measures}
Desmedt, O., Heeren, A., Corneille, O., \& Luminet, O. (2022). What do
measures of self-report interoception measure? Insights from a
systematic review, latent factor analysis, and network approach.
\emph{Biological Psychology}, \emph{169}, 108289.

\bibitem[\citeproctext]{ref-desmedt2025discrepancies}
Desmedt, O., Luminet, O., Maurage, P., \& Corneille, O. (2025).
Discrepancies in the definition and measurement of human interoception:
A comprehensive discussion and suggested ways forward.
\emph{Perspectives on Psychological Science}, \emph{20}(1), 76--98.

\bibitem[\citeproctext]{ref-di2016alexithymia}
Di Tella, M., \& Castelli, L. (2016). Alexithymia in chronic pain
disorders. \emph{Current Rheumatology Reports}, \emph{18}(7), 41.

\bibitem[\citeproctext]{ref-ferentzi2021examining}
Ferentzi, E., Olaru, G., Geiger, M., Vig, L., Köteles, F., \& Wilhelm,
O. (2021). Examining the factor structure and validity of the
multidimensional assessment of interoceptive awareness. \emph{Journal of
Personality Assessment}, \emph{103}(5), 675--684.

\bibitem[\citeproctext]{ref-friedman2008sparse}
Friedman, J., Hastie, T., \& Tibshirani, R. (2008). Sparse inverse
covariance estimation with the graphical lasso. \emph{Biostatistics},
\emph{9}(3), 432--441.

\bibitem[\citeproctext]{ref-garfinkel2015knowing}
Garfinkel, S. N., Seth, A. K., Barrett, A. B., Suzuki, K., \& Critchley,
H. D. (2015). Knowing your own heart: Distinguishing interoceptive
accuracy from interoceptive awareness. \emph{Biological Psychology},
\emph{104}, 65--74.
\url{https://doi.org/10.1016/j.biopsycho.2014.11.004}

\bibitem[\citeproctext]{ref-garnefski2006cognitive}
Garnefski, N., \& Kraaij, V. (2006). Cognitive emotion regulation
questionnaire--development of a short 18-item version (CERQ-short).
\emph{Personality and Individual Differences}, \emph{41}(6), 1045--1053.

\bibitem[\citeproctext]{ref-golino2017exploratory}
Golino, H. F., \& Epskamp, S. (2017). Exploratory graph analysis: A new
approach for estimating the number of dimensions in psychological
research. \emph{PloS One}, \emph{12}(6), e0174035.

\bibitem[\citeproctext]{ref-golino2020investigating}
Golino, H., Shi, D., Christensen, A. P., Garrido, L. E., Nieto, M. D.,
Sadana, R., Thiyagarajan, J. A., \& Martinez-Molina, A. (2020).
Investigating the performance of exploratory graph analysis and
traditional techniques to identify the number of latent factors: A
simulation and tutorial. \emph{Psychological Methods}, \emph{25}(3),
292.

\bibitem[\citeproctext]{ref-indahl2018similarity}
Indahl, U. G., Næs, T., \& Liland, K. H. (2018). A similarity index for
comparing coupled matrices. \emph{Journal of Chemometrics},
\emph{32}(10), e3049.

\bibitem[\citeproctext]{ref-jimenez2023dimensionality}
Jiménez, M., Abad, F. J., Garcia-Garzon, E., Golino, H., Christensen, A.
P., \& Garrido, L. E. (2023). Dimensionality assessment in bifactor
structures with multiple general factors: A network psychometrics
approach. \emph{Psychological Methods}.

\bibitem[\citeproctext]{ref-jovanovic2020longer}
Jovanović, V., \& Lazić, M. (2020). Is longer always better? A
comparison of the validity of single-item versus multiple-item measures
of life satisfaction. \emph{Applied Research in Quality of Life},
\emph{15}, 675--692.

\bibitem[\citeproctext]{ref-khalsa2018interoception}
Khalsa, S. S., Adolphs, R., Cameron, O. G., Critchley, H. D., Davenport,
P. W., Feinstein, J. S., \& Zucker, N. (2018). Interoception and mental
health: A roadmap. \emph{Biological Psychiatry: Cognitive Neuroscience
and Neuroimaging}, \emph{3}(6), 501--513.
\url{https://doi.org/10.1016/j.bpsc.2017.12.004}

\bibitem[\citeproctext]{ref-kolacz2018body}
Kolacz, J., Holmes, L., \& Porges, S. W. (2018). Body perception
questionnaire (BPQ) manual. \emph{Traumatic Stress Research Consortium}.

\bibitem[\citeproctext]{ref-kroenke2009ultra}
Kroenke, K., Spitzer, R. L., Williams, J. B., \& Löwe, B. (2009). An
ultra-brief screening scale for anxiety and depression: The PHQ--4.
\emph{Psychosomatics}, \emph{50}(6), 613--621.

\bibitem[\citeproctext]{ref-leising2009toronto}
Leising, D., Grande, T., \& Faber, R. (2009). The toronto alexithymia
scale (TAS-20): A measure of general psychological distress.
\emph{Journal of Research in Personality}, \emph{43}(4), 707--710.

\bibitem[\citeproctext]{ref-makowski2025measuring}
Makowski, D., Te, A., Neves, A., \& Chen, S. A. (2025). \emph{Measuring
depression and anxiety with 4 items? Adaptation of the PHQ-4 to increase
its sensitivity to subclinical variability}. PsyArXiv.
\url{https://doi.org/10.31234/osf.io/436np_v2}

\bibitem[\citeproctext]{ref-mayer2011exploratory}
Mayer, C.-D., Lorent, J., \& Horgan, G. W. (2011). Exploratory analysis
of multiple omics datasets using the adjusted RV coefficient.
\emph{Statistical Applications in Genetics \& Molecular Biology},
\emph{10}(1).

\bibitem[\citeproctext]{ref-mehling2018multidimensional}
Mehling, W. E., Acree, M., Stewart, A., Silas, J., \& Jones, A. (2018).
The multidimensional assessment of interoceptive awareness, version 2
(MAIA-2). \emph{PloS One}, \emph{13}(12), e0208034.

\bibitem[\citeproctext]{ref-murphy2020testing}
Murphy, J., Brewer, R., Plans, D., Khalsa, S. S., Catmur, C., \& Bird,
G. (2020). Testing the independence of self-reported interoceptive
accuracy and attention. \emph{Quarterly Journal of Experimental
Psychology}, \emph{73}(1), 115--133.

\bibitem[\citeproctext]{ref-murphy2018classifying}
Murphy, J., Catmur, C., \& Bird, G. (2018). Classifying individual
differences in interoception: Implications for the measurement of
interoceptive accuracy. \emph{Psychonomic Bulletin \& Review},
\emph{25}(4), 1424--1431.
\url{https://doi.org/10.3758/s13423-017-1339-z}

\bibitem[\citeproctext]{ref-paulus2010interoception}
Paulus, M. P., \& Stein, M. B. (2010). Interoception in anxiety and
depression. \emph{Brain Structure and Function}, \emph{214}(5),
451--463.

\bibitem[\citeproctext]{ref-sibson1978studies}
Sibson, R. (1978). Studies in the robustness of multidimensional
scaling: Procrustes statistics. \emph{Journal of the Royal Statistical
Society: Series B (Methodological)}, \emph{40}(2), 234--238.

\bibitem[\citeproctext]{ref-theriault2024check}
Thériault, R., Ben-Shachar, M. S., Patil, I., Lüdecke, D., Wiernik, B.
M., \& Makowski, D. (2024). Check your outliers! An introduction to
identifying statistical outliers in r with easystats. \emph{Behavior
Research Methods}, \emph{56}(4), 4162--4172.

\bibitem[\citeproctext]{ref-traag2019louvain}
Traag, V. A., Waltman, L., \& Van Eck, N. J. (2019). From louvain to
leiden: Guaranteeing well-connected communities. \emph{Scientific
Reports}, \emph{9}(1), 1--12.

\bibitem[\citeproctext]{ref-veilleux2024development}
Veilleux, J. C., Schreiber, R. E., Warner, E. A., \& Brott, K. H.
(2024). Development and validation of a brief version of the emotion
reactivity scale: The b-ERS. \emph{Current Psychology}, \emph{43}(14),
12586--12600.

\bibitem[\citeproctext]{ref-vig2022questionnaires}
Vig, L., Köteles, F., \& Ferentzi, E. (2022). Questionnaires of
interoception do not assess the same construct. \emph{PLoS One},
\emph{17}(8), e0273299.

\bibitem[\citeproctext]{ref-zamariola2018interoceptive}
Zamariola, G., Maurage, P., Luminet, O., \& Corneille, O. (2018).
Interoceptive accuracy scores from the heartbeat counting task are
problematic: Evidence from simple bivariate correlations.
\emph{Biological Psychology}, \emph{137}, 12--17.

\end{CSLReferences}

\end{document}